### 2021-12-03 doc/contrastive-learning.tex
adds first version of contrastive learning doc

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry} 
\usepackage{amsmath}
\usepackage{tcolorbox}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{lastpage}
\usepackage{fancyhdr}
\usepackage{accents}

\pagestyle{fancy}
\setlength{\headheight}{40pt}

\begin{document}

\lhead{Contrastive Learning \& Pre-training on 3D Data}
\rhead{Theory \& Proposal}
\cfoot{\thepage\ of \pageref{LastPage}}

\section{background}

\begin{itemize}
    \item Want to leverage unlabeled Data
    \item This is especially important in 3D
          \begin{itemize}
              \item where large 3D datasets do not exist on the web
              \item Easy to capture large amounts of 3D data with new sensors
              \item Extremely hard to label when compared to 2D
          \end{itemize}
    \item Semi-Supervised Contrastive pre-training has show to outperform supervised pre-training in some 2D vision downstream tasks
    \item Focus more on learning structure of the Data
    \item Learning from structure should be more general to downstream tasks than pre-training on a particular objective
    \item Contrastive Learning is a subset of Representational learning that aims to learn how to represent high dimensional data in a lower dimensional space.
          \begin{itemize}
              \item Generative (e.g. VAE)
              \item discriminative (e.g. Contrastive)
              \item Contrastive-Generative (Adversarial)
              \item Contrastive has shown biggest promise so far
          \end{itemize}
    \item Contrative learning learns by comparing similarities among data points.
\end{itemize}

\section{Theory}

Much of this theory is coming from the framework laid out in \cite{le-khac_contrastive_2020}

\subsection{Sampling}

To learn from comparing data samples, we first need to define positive and negative distributions. These allow us to sample points which we know should be like and which should be un-like. We define these distributions as:
%

$$\mathbf{x}^{+} \sim p^{+}( \cdot | \mathbf{x})$$
$$\mathbf{x}^{-} \sim p^{-}( \cdot | \mathbf{x})$$

Where x is an individual data point. At a high level, we want to sample points from both distributions and make those coming from the positive distribution close together in feature space and those from the negative distribution far in feature space.
%

In a supervised setting, these distributions can be defined by the labels themselves. But in self-supervised, we have to define these distributions using the raw data only.
%

Note that the definition of these two distributions are ultimately what will determine what kind of features and invariances the encoder will learn.
%

\subsection{Pre-text Tasks}

As mentioned in the previous section, the labels on which we are learning have to come solely from the data itself. This is done by defining some kind of pre-text task that automatically labels the data without human annotation.
%

For example, a common pre-text task in natural language processing is to mask a part of a sentence and try to predict the masked token. This would fall under the category of Predictive Coding \cite{oord_representation_2019} and is a generative method instead of contrastive method.
%

In vision, some common pre-text tasks are
\begin{itemize}
    \item \textbf{Colorization:} Predict the colorization of a grey-scale image
    \item \textbf{Rotation:} Predict the rotation applied to an image
    \item \textbf{Jigsaw:} Rebuild an image from randomly permuted rectangular patches
    \item \textbf{Instance Discrimination:} Treat each data point as its own instance
\end{itemize}
%

The most common contrastive pre-text task in vision, is instance discrimination. The idea is to treat each data point as its own instance. Then the goal is to discriminate this instance, an any of its augmentations, from other instances in the dataset coming from the negative distribution.
%

We define the positive distribution as different augmentations of the same data point. These augmentations come from a set of transforms $\mathcal{T}$ where we sample two (or more) transformations $ t, t^{'} \sim \mathcal{T} $ and apply them to the original data point.
%

$$x = t(x), \quad x^{+} = t^{'}(x)$$

Common transforms in 2D vision include random cropping, colour jitter, gaussian blurring, and elastic distortion.
%

Negative Samples are then defined as any other data point and its augmentations. This is obviously a very simple and naive sampling strategy. Its limitations will be addressed in a future section.
%

\subsection{Encoder \& Decoder}

The goal of representational learning is to learn a good mapping from the high-dimensional raw input $\mathbf{x} \in \mathbb{R}^{n_0 \times \dots \times n_N}$ to a smaller one-dimensional but more expressive latent representation in feature space denoted as $\mathbf{v} \in \mathbb{R}^{d}$ where $d << n_0 \times \dots \times n_N$. To get this representation we will use some function $e()$ which performs the mapping $\mathbf{v} = e(\mathbf{x})$. This function is usually implemented as some sort of neural network.
%

This representation is then fed into a decoder (otherwise known as a projection head) that projects these embeddings into a metric space that is used directly to contrast different instances from each other.

\subsection{Contrastive Loss}

\section{Contrastive in 3D}

\section{Limitations}

\section{Addressing Limitations}

\subsection{Supervised Contrastive Learning}

\subsection{Clustering}

\subsection{Memory Banks \& Momentum Encoders}

\subsection{Hard Negative \& Positive Mining}

\section{Proposal}

\newpage
\bibliographystyle{plain}
\bibliography{contrastive-learning}

\end{document}

### 2021-12-07 doc/contrastive-learning.tex
adds content on negative sampling

@@
 The most common contrastive pre-text task in vision, is instance discrimination.
+This is also the task that we will focus on through the remainder of the document.
 The idea is to treat each data point as its own instance.
-Then the goal is to discriminate this instance, an any of its augmentations, from other instances in the dataset coming from the negative distribution.
+Then the goal is to discriminate this instance, and any of its augmentations, from other instances in the dataset coming from the negative distribution.
+It's important to note that the augmentations that are selected will define what the network learns to be invariant to.
+It also has been shown that the proper selection of augmentations is absolutely vital to getting any contrastive system working.
 %
@@
 This representation is then fed into a decoder (otherwise known as a projection head) that projects these embeddings into a metric space that is used directly to contrast different instances from each other.
+We define this decoder as $z=h(\mathbf{v})$ with $\{z | z \in \mathbb{R}^{d'}, \lVert z \lVert^{2} = 1\}$ being another normalized latent representation but in a smaller dimension than the encoded vectors (i.e.
+$z \in \mathbb{R}^{d'}$ where $d^{'} < d$).
+The projection $z$ is what is ultimately being used to measure similarity between samples.
+It is also thrown away after pre-training and only the encoder is kept for downstream tasks.
 
@@
 
+The final component, after selecting positive and negatives samples along with an architecture, is to select what final loss function to optimize for.
+Whatever the function, it should have some common features.
+The first is that it makes positive samples close together in embedding space.
+Although this could potentially be sufficient, there exists a degenerate solution whereby all samples are given the same embedding and thus brings the loss to zero.
+To counter against this degenerate solution, we require some sort of repulsive force to balance out the attractive force.
+A simple solution is to use negative samples and make them far apart from their positive matches.
+%
+
+For notation we will define the concept of a query $q$ and key $k$.
+This is so that the problem can be formulated as a dictionary lookup where we want to make a query point similar to positive keys $k^+$ or dissimilar to negative keys $k^-$.
+
+This was first made popular in Noise Contrastive Estimation (NCE) where this was formulated with a single negative sample to compare against.
+
+$$\mathcal{L} = -\log \frac{\exp(q \cdot k^{+})}{\exp(q \cdot k^{+}) + \exp(q \cdot k^{-})}$$
+%
+
+However this has been shown to be hard to train as a single negative sample has a very weak signal and makes it hard to train.
+Instead, InfoNCE and NT-Xent were proposed to use many more negative samples with an added temperature parameter.
+This can be formulated as follows:
+
+$$\mathcal{L}_{i} = -\log \frac{\exp(q_{i} \cdot k^{+}_{i} / \tau)}{\exp(q_{i} \cdot k^{+}_{i} / \tau) + \sum^{K}_{j}\exp(q_{i} \cdot k^{-}_{j} / \tau)}$$
+
+Where $\tau \in (0,1]$ and represents the temperature parameter that controls the smoothness of the latent representations.
+It also scales the gradient by $1/\tau$ and so that has to be scaled along with the temperature.
+Generally it was found that smaller temperature values benefit training but having them too small creates a lot of numerical instability.
+
+It's been shown that NCE is an upperbound on the Mutual Information between samples...
+
+This is not the only loss formulation. There exist others using marginal losses, triplet losses, N-pair losses (which is shown to be a form of InfoNCE loss)
+
+We don't have to limit ourselves to a single positive either.
+That's what the formulation in supervised contrastive learning \cite{khosla_supervised_2021} used.
+There are two options, where the extra positives can be placed inside the log or outside but it was shown that placing it outside is preferred.
+This is formulated as:
+
+$$\mathcal{L}_{i} = \frac{-1}{|P(i)|} \sum_{p \in P(i)} -\log \frac{\exp(q_{i} \cdot k^{+}_{p} / \tau)}{\exp(q_{i} \cdot k^{+}_{i} / \tau) + \sum^{K}_{j}\exp(q_{i} \cdot k^{-}_{j} / \tau)}$$
+
 \section{Contrastive in 3D}
 
+Two options for instance discrimination in 3D, use entire scene or use points.
+
 \section{Limitations}
+
+The biggest limitation of contrastive learning is the need to select positive and negative samples.
+The positive samples have been shown to be really important in learning invariant properties, while a large negative sample size is critical to properly learn to separate samples.
+%
+
+Sampling large batches of negative samples assumes that there is a low probability of false negatives in the batch.
+Otherwise, many keys that should be close to the query will end up being pushed further away.
+
+Using really large negative batches requires a lot of computational resources and dramatically slows down training making research on the topic much slower and only possible for those researchers who have access to such computational resources.
+%
+
+Many of the approaches also do not condition their negative samples based off of their query.
 
@@
 
+This method assumes that even if we don't have complete label information, we have at least some ground truth labels.
+It tries to bridge self-supervised contrastive learning into supervised learning.
+With the few labels that we do have, we can train our model and use it's outputs to generate pseudo-labels.
+For outputs that the model is sufficiently certain of, it can use these to select positive samples as those in the same class as the query and negative samples as those from any other class.
+This has shown to work especially well, beating fully supervised pre-training on ImageNet in downstream vision tasks \cite{khosla_supervised_2021}.
+%
+
+The core of their contribution is re-formulation of the InfoNCE loss to use multiple positive samples. This is required as they use multiple positive samples from the same class.
+%
+
+It has also been shown to be useful in 3D \cite{jiang_guided_2021}
+
 \subsection{Clustering}
+
+Prototypical Contrastive Learning \cite{li_prototypical_2021}, BYOL \cite{grill_bootstrap_nodate} or SwAV \cite{caron_unsupervised_2021}
 
@@
 
+This method tries to deal with negative sampling by just having a lot of negative samples.
+This is obviously very brute force and requires a lot of compute resources.
+If we define $Q$ as a memory bank of fixed size that holds features from the previous iterations then we can reformulate the loss as:
+
+$$\mathcal{L}_{i} = -\log \frac{\exp(q_{i} \cdot k^{+}_{i} / \tau)}{\exp(q_{i} \cdot k^{+}_{i} / \tau) + \sum_{k^{-} \in Q} \exp(q_{i} \cdot k^{-} / \tau)}$$
+
 \subsection{Hard Negative \& Positive Mining}
 
+In the beginning when many negative samples are close to the query point, the gradient contribution from each negative sample is strong.
+But as training goes on, negative points become farther from the query and thus contribute much less to the gradient.
+Instead of tackling this using extremely large memory banks or batch sizes, this area of research tries to smartly select samples that best benefit encoder training and thus reduce the computational resources required and decrease the training time.
+%
+
+One approach is instead of using real data points, to generate synthetic data points from those negative samples that are closest to the query as was done in Mixing of Hard Negatives (MoChi) \cite{kalantidis_hard_nodate}.
+This creates more hard negative samples to train from and should increase the gradient contribution from the samples.
+%
+
+We can define these synthetic negatives as a mix of other negative samples that are most similar to the query point.
+
+$$\mathbf{h}_k = \frac{\mathbf{h}_k}{\lVert \mathbf{h}_k \lVert^{2}}, \quad \text{where} \: \mathbf{h}_k = \alpha \mathbf{k_i} + (1-\alpha) \mathbf{k_j}$$
+
+where $k_i$ and $k_j$ are randomly chosen from a set $\tilde{Q}$ of the logits that are most similar to the query point.
+Alpha is a mixing factor that is randomly selected from $(0,1)$.
+In theory, this shouldn't generate samples that are any harder than the existing negative samples but it does stretch the space around the query point more uniformly.
+
+One way to make the task even harder is to mix the negative with the query point.
+
+$$\mathbf{h}_k = \frac{\mathbf{h}_k}{\lVert \mathbf{h}_k \lVert^{2}}, \quad \text{where} \: \mathbf{h}_k = \beta \mathbf{q} + (1-\beta) \mathbf{k_i}$$
+
+This technique however does not address the case where there may exist a large number of false negatives that would and should cluster around the query point.
+
+Another approach that addresses this limitation is to remove these points from the dataset.
+Selecting false negatives is refereed to as a sampling bias in \cite{chuang_debiased_2020}.
+The key idea behind their approach is to approximate the underlying distribution of negative samples.
+They do this by accounting for the presence of positive samples in the denominator.
+
+$$\mathcal{L}_{debiased} = -\log \frac{\exp(q_{i} \cdot k^{+}_{i} / \tau)}{\exp(q_{i} \cdot k^{+}_{i} / \tau) + Ng(x, \{u_i\}^N_i, \{v_i\}^M_i)}$$
+
+$$g(x, \{u_i\}^N_i, \{v_i\}^M_i) = \frac{1}{\kappa^-}\max \{\frac{1}{N} \sum_i^N exp(q \cdot u_i)  - \tau^+\frac{1}{M} \sum_i^M exp(q \cdot v_i), \: \exp(\frac{1}{\tau})\}$$
+
+where $\kappa^+$ is the uniform class probability and $\kappa^- = 1 - \kappa^+$
+%
+
+A followup work was presented in \cite{robinson_contrastive_2021}.
+They define informative negative samples as those that have embeddings close to the query but in reality should be far apart.
+They ask the question what makes a good negative point?
+Their answer is that 1) whose labels are different from the query point and 2) those whose embedding is most similar to the query point.
+They have a method to get the first principal approximately.
+The challenge is that points near the query are the hardest but also have a higher likelihood of being of the same class.
+
+Their derivation is quite complicated but they end up with a similar formulation to \cite{chuang_debiased_2020} except with an additional parameter $\beta$ which specifies a concentration parameter.
+The formulation then becomes:
+
+$$g(x, \{u_i\}^N_i, \{v_i\}^M_i) = \frac{1}{\kappa^-}\max \{ \eta \frac{1}{N} \sum_i^N exp(q \cdot u_i)  - \tau^+\frac{1}{M} \sum_i^M exp(q \cdot v_i), \: \exp(\frac{1}{\tau})\}$$
+
+$$\eta = \frac{1}{N} \sum_i^N exp(\beta q \cdot u_i) $$
+
+Not sure if this is right
+
+Both of the above techniques \cite{chuang_debiased_2020} and \cite{robinson_contrastive_2021} have extremely simple implementations despite their complicated derivations.
+
+It can be hard to tune properly
+%
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.5]{images/synthetic-negatives.png}
+        \caption{Negative Hard Mixing \cite{kalantidis_hard_nodate}}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.45]{images/debiased.png}
+        \caption{t-SNE visualization of debiased objective \cite{chuang_debiased_2020}}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/hard-samplign.png}
+        \caption{Pytorch pseudo code of hard sampling objectives. \cite{robinson_contrastive_2021}}
+    \end{center}
+\end{figure}
+
 \section{Proposal}
+
+\pagebreak
 

### 2021-12-07 doc/contrastive-learning.tex
adds first version of proposal

@@
 \section{Addressing Limitations}
+\label{limitations}
 
 \subsection{Supervised Contrastive Learning}
+\label{limitations:supervised}
 
@@
 \subsection{Clustering}
+\label{limitations:clustering}
 
@@
 \subsection{Memory Banks \& Momentum Encoders}
+\label{limitations:memory}
 
@@
 \subsection{Hard Negative \& Positive Mining}
+\label{limitations:mining}
 
@@
     \begin{center}
-        \includegraphics[scale=0.5]{images/synthetic-negatives.png}
+        \includegraphics[scale=0.35]{images/synthetic-negatives.png}
         \caption{Negative Hard Mixing \cite{kalantidis_hard_nodate}}
@@
 
+\pagebreak
+
 \section{Proposal}
+
+My work is focused on improving self-supervised contrastive learning in 3D point clouds for semantic and instance segmentation.
+I will use the work of PointContrast as my starting point.
+My focus will be on improving the Positive and Negative sampling strategies so that the pretext task and downstream performance improves as well as requiring fewer computational resources.
+
+\subsection{Contrastive Loss}
+
+It's been shown that InfoNCE loss with a temperature parameter is the most effective contrastive loss and the main one used in all recent vision applications.
+For this reason I will stick with this loss function.
+It is reprinted below for completeness.
+
+$$\mathcal{L}_{i} = -\log \frac{\exp(q_{i} \cdot k^{+}_{i} / \tau)}{\exp(q_{i} \cdot k^{+}_{i} / \tau) + \sum^{K}_{j}\exp(q_{i} \cdot k^{-}_{j} / \tau)}$$
+
+\subsection{Negative Sampling}
+
+The main focus will be on improving the negative sampling techniques.
+Currently, negative samples are selected uniformly at random from the set of positive keys.
+However, this is a naive sampling strategy that is not conditioned on the sampled query point at all.
+It introduces the possibility for a large number of false negatives or semi-false negatives.
+Ideally we want to sample from a distribution like:
+
+$$\mathbf{x}^{-} \sim p^{-}( \cdot | \mathbf{x})$$
+
+To condition on the input point, I plan to use Fast Histogram Point Features (FHPF) as a way to select negative points that are sufficiently different from the input point.
+This operates on the assumption that if the point is different locally, then it for sure is different globally even if the surrounding area is the same.
+
+$$p^{-}( \cdot | \mathbf{x}) = \text{Uniform}(\{ \: \mathbf{x}^- \: | \: \text{sim}(\text{FPFH}(\mathbf{x}^-),\text{FPFH}(\mathbf{x})) > \delta \})$$
+
+where $\text{sim}(\cdot,\cdot)$ is some kind of distance function between the histograms such as earth movers distance.
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.3]{images/regions.png}
+        \caption{Visualization of local vs global region dissimilarity}
+    \end{center}
+\end{figure}
+
+Instead of uniform sampling, I can also try cluster sampling by clustering different local features into feature types (e.g.
+flat surfaces, round objects, corners) and sample equally from those distributions.
+
+$$p^{-}( \cdot | \mathbf{x}) = \text{Uniform}(\{ \textbf{c} | \textbf{c} \in \text{KNN}\{ \text{FPFH} ( \{ \textbf{x}^- | \textbf{x}^- \in \mathcal{X} \} ) \}, \textbf{c} \neq \textbf{x} \})$$
+
+Where $\text{KNN}$ represents the K-nearest-neigbours clustering algorithm and $\textbf{c}$ is a set of negative samples clustered together based on their similar properties.
+Thus this strategy constitutes sampling equally from different clusters of points.
+%
+
+Another option is to weight the negative samples based off of the local dissimilarity of the points and incorporate this into the loss function itself.
+%
+
+$$\mathcal{L}_{i} = -\log \frac{\exp(q_{i} \cdot k^{+}_{i} / \tau)}{\exp(q_{i} \cdot k^{+}_{i} / \tau) + \sum^{K}_{j}\exp(w_{i,j} (q_{i} \cdot k^{-}_{j}) / \tau)}$$
+
+$$\text{where} \: w_{i,j} = 1 - \frac{s_{i,j}}{\sum_k^{K}s_{i,k}} $$
+$$ s_{i,j} = \text{sim}(\text{FPFH}(q_{i}), \text{FPFH}(k^{-}_{j}))$$
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.5]{images/fpfh.png}
+        \caption{Visualization of Fast Point Feature Histogram features in a sample point cloud}
+    \end{center}
+\end{figure}
+
+The last possibility is to look at applying different loss functions at different parts of the network depending on the receptive field.
+So that the parts looking at local regions have different objects than those looking at more global regions.
+Perhaps there's no advantage to training the global region at all?
+
+\subsection{Positive Sampling}
+
+Using different augmentations or adding points from the same cluster as the query.
+
+\subsection{Informative Sampling}
+
+\subsection{Other Techniques}
+
+Implement all the negative sampling techniques from section \ref{limitations:mining} as well as compare using a momentum encoder and perhaps one of the clustering techniques from section \ref{limitations:clustering}
 

### 2021-12-08 doc/contrastive-learning.tex
finishes proposal

@@
 \end{figure}
-
+\newpage
 \begin{figure}[h!]
@@
 
+\newpage
 \pagebreak
+\newpage
 
@@
 
-$$p^{-}( \cdot | \mathbf{x}) = \text{Uniform}(\{ \textbf{c} | \textbf{c} \in \text{KNN}\{ \text{FPFH} ( \{ \textbf{x}^- | \textbf{x}^- \in \mathcal{X} \} ) \}, \textbf{c} \neq \textbf{x} \})$$
+$$p^{-}( \cdot | \mathbf{x}) = \text{Uniform}(\{ \textbf{c} \: | \: \textbf{c} \in \text{KNN}\{ \text{FPFH} ( \{ \: \textbf{x}^- \: | \: \textbf{x}^- \in \mathcal{X} \: \} ) \}, \textbf{c} \neq \textbf{x} \})$$
 
@@
 
-Using different augmentations or adding points from the same cluster as the query.
+The way you define positive sampling defines what you want the network to be invariant to.
+In the case where positive samples are different data augmentations, then the network will try to become invariant to these transforms.
+This may be desirable or undesirable for the task at hand.
+It has been shown in other works also that the positive sampling is what ultimately defines the hardness of the task.
+%
+
+Currently, the only data augmentations are random rotations.
+We can explore other kind of augmentations such as different elastic distortions but it may be too hard to learn as already it struggles to overfit.
+Another option is to train different networks to be invariant to different things such as one that is invariant to rotations and another that is not.
+%
+
+If we're using clusters, we can add points from the same cluster as other positive samples.
+%
 
@@
 
+Beyond positive and negative sampling techniques, it maybe be beneficial to sample points that have locally interesting features.
+Currently many of the points that are sampled are those on the wall which is not very informative or helpful to the algorithm.
+During supervised training its able to learn walls and floors the quickest and with high accuracy so any pre-training for flat surfaces may be a waste of time.
+
+Instead, we can to to select points according to the entropy of their feature vector.
+
+$$p^{-}( \cdot | \mathbf{x}) = \text{Weighted}(\{ \: (w_i, \mathbf{x}^-_i) \: | \: \mathbf{x}^-_i \in \mathcal{X} \: \})$$
+
+$$\text{where} \: w_i = \frac{\text{Entropy}(\text{FPFH}(\mathbf{x}^-_i))}{\sum_k^K \text{Entropy}(\text{FPFH}(\mathbf{x}^-_k))}$$
+
+With $w_i$ being the weighted probability of the sample being selected.
+
 \subsection{Other Techniques}
 
-Implement all the negative sampling techniques from section \ref{limitations:mining} as well as compare using a momentum encoder and perhaps one of the clustering techniques from section \ref{limitations:clustering}
+Although there does not exist and other kind of point sampling strategy in 3D (or even for 2D in fact), I would like to compare some of the other techniques that try to deal with the sampling problem.
+%
+
+I would like to compare my approach to one of each of the different options for addressing sampling issues. These being:
+
+\begin{itemize}
+    \item Momentum encoder from section \ref{limitations:memory}
+    \item Hard negative sampling from section \ref{limitations:mining}
+    \item Hard negative mixing from section \ref{limitations:mining}
+    \item Clustering using either SwAV or BYOL from section \ref{limitations:clustering}
+    \item Supervised contrastive from section \ref{limitations:supervised}
+\end{itemize}
 

### 2022-01-03 src/models/minkowski/dataset.py
adds entropy loss function

@@ -185,6 +185,19 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
             # Extract data
             xyz = np.ascontiguousarray(frame.points)
             features = torch.from_numpy(frame.point_colors)
+            fpfh = frame.fpfh
+            entropies = frame.entropies
+
+            # VISUALIZATION
+            # import copy
+            # inds = np.where(entropies < 0.4)[0]
+            # entropies_ = copy.deepcopy(entropies)
+            # entropies_[inds] = 0
+            # entropies_ = 1 - entropies_
+            # entropies_ = entropies_ / entropies_.max()
+            # visualize_mapping(xyz, entropies)
+            # colors = get_color_map(entropies_)
+            # features = torch.from_numpy(colors)
 
             # apply a random scalling
             xyz *= random_scale
@@ -212,6 +225,8 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
                     "discrete_coords": discrete_coords,
                     "unique_feats": unique_feats,
                     "mapping": mapping,
+                    "fpfh": fpfh,
+                    "entropies": entropies,
                 }
             )
 
@@ -222,9 +237,13 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
             {
                 "frame1": {
                     "voxel_inds": mapping1[k],
+                    "fpfh": quantized_frames[0]["fpfh"][k],
+                    "entropies": quantized_frames[0]["entropies"][k],
                 },
                 "frame2": {
                     "voxel_inds": mapping2[v],
+                    "fpfh": quantized_frames[1]["fpfh"][v],
+                    "entropies": quantized_frames[1]["entropies"][v],
                 },
             }
             for k, v in correspondences.items()
@@ -404,6 +423,30 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
             "correspondences": correspondences,
             "quantized_frames": quantized_frames,
         }
+
+
+import open3d as o3d
+from matplotlib import pyplot as plt
+
+
+def get_color_map(x):
+    colours = plt.cm.Spectral(x)
+    return colours[:, :3]
+
+
+def visualize_mapping(points1, entropies):
+    # points1 = points1.detach().cpu().numpy()
+
+    pcd1 = o3d.geometry.PointCloud()
+    pcd1.points = o3d.utility.Vector3dVector(points1)
+
[…]

### 2022-01-03 src/models/minkowski/trainer.py

@@ -87,6 +87,10 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         # which training loss to use
         if cfg.model.net.loss == "new":
             self.loss_fn = self.loss_fn_new
+        elif cfg.model.net.loss == "entropy":
+            self.loss_fn = self.loss_fn_entropy
+        elif cfg.model.net.loss == "delta":
+            self.loss_fn == self.loss_fn_delta
         else:
             self.loss_fn = self.loss_fn_original
 
@@ -186,8 +190,7 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
             q = q[inds]
             k = k[inds]
 
-        # TODO: Comment out for now but normalization should be done here
-        # # normalize to unit vectors
+        # normalize to unit vectors
         q = q / torch.norm(q, p=2, dim=1, keepdim=True)
         k = k / torch.norm(k, p=2, dim=1, keepdim=True)
 
@@ -289,47 +292,49 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         qs, ks = [], []
         es = []
         for i, matches in enumerate(batch.correspondences):
-            voxel_indices_1 = [v for v, _ in matches]
-            voxel_indices_2 = [v for _, v in matches]
-            # entropies = [v for _, _, v in matches]
+            voxel_indices_1 = [match["frame1"]["voxel_inds"] for match in matches]
+            voxel_indices_2 = [match["frame2"]["voxel_inds"] for match in matches]
+            entropies = [match["frame1"]["entropies"] for match in matches]
 
             output_batch_1 = output.features_at(2 * i)
             output_batch_2 = output.features_at(2 * i + 1)
             q = output_batch_1[voxel_indices_1]
             k = output_batch_2[voxel_indices_2]
 
+            # Visualize sampling
+            entropies = np.array(entropies)
+            inds = np.where(entropies < 0.4)[0]
+            entropies[inds] = 0
+
+            # VISUALIZATION
+            # entropies = entropies / entropies.sum()
+            # voxel_indices_1 = np.random.choice(
+            #     voxel_indices_1, 1000, p=entropies, replace=False
+            # )
             # model_input = ME.SparseTensor(batch.features, batch.points)
             # points1 = model_input.coordinates_at(2 * i)
-            # points2 = model_input.coordinates_at(2 * i + 1)
-
[…]

### 2022-01-03 src/models/trainers.py

@@ -21,26 +21,30 @@ import util.eval_semantic as eval_semantic
 
 log = logging.getLogger(__name__)
 
+
 class LambdaStepLR(LambdaLR):
+    def __init__(self, optimizer, lr_lambda, last_step=-1):
+        super(LambdaStepLR, self).__init__(
+            optimizer, lr_lambda, last_step, verbose=True
+        )
 
-  def __init__(self, optimizer, lr_lambda, last_step=-1):
-    super(LambdaStepLR, self).__init__(optimizer, lr_lambda, last_step)
+    @property
+    def last_step(self):
+        """Use last_epoch for the step counter"""
+        return self.last_epoch
 
-  @property
-  def last_step(self):
-    """Use last_epoch for the step counter"""
-    return self.last_epoch
-
-  @last_step.setter
-  def last_step(self, v):
-    self.last_epoch = v
+    @last_step.setter
+    def last_step(self, v):
+        self.last_epoch = v
 
 
 class PolyLR(LambdaStepLR):
-  """DeepLab learning rate policy"""
+    """DeepLab learning rate policy"""
 
-  def __init__(self, optimizer, max_iter, power=0.9, last_step=-1):
-    super(PolyLR, self).__init__(optimizer, lambda s: (1 - s / (max_iter + 1))**power, last_step)
+    def __init__(self, optimizer, max_iter, power=0.9, last_step=-1):
+        super(PolyLR, self).__init__(
+            optimizer, lambda s: (1 - s / (max_iter + 1)) ** power, last_step
+        )
 
 
 def configure_optimizers(parameters, optimizer_cfg, scheduler_cfg):
@@ -68,7 +72,9 @@ def configure_optimizers(parameters, optimizer_cfg, scheduler_cfg):
     elif scheduler_cfg.type == "ExpLR":
         scheduler = ExponentialLR(optimizer, scheduler_cfg.exp_gamma)
     elif scheduler_cfg.type == "PolyLR":
-        scheduler = PolyLR(optimizer, max_iter=scheduler_cfg.max_iter, power=scheduler_cfg.poly_power)
+        scheduler = PolyLR(
+            optimizer, max_iter=scheduler_cfg.max_iter, power=scheduler_cfg.poly_power
+        )
     else:
         log.error(f"Invalid scheduler type: {scheduler_cfg.type}")
         raise ValueError(f"Invalid scheduler type: {scheduler_cfg.type}")

### 2022-01-03 scripts/preprocess/scannet.sh
updates preprocess script

@@ -1,5 +1,5 @@
 export SCANNET_DIR=/media/starslab/datasets/scannet
-export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh
+export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh_v2
 export HYDRA_FULL_ERROR=1
 
 preprocess() {
@@ -15,4 +15,4 @@ preprocess() {
 export -f preprocess
 
 # preprocess
-parallel -j 4 --linebuffer time preprocess ::: `find $SCANNET_DIR/scans/ -mindepth 1 -maxdepth 1 -type d`
+parallel -j 16 --linebuffer time preprocess ::: `find $SCANNET_DIR/scans/ -mindepth 1 -maxdepth 1 -type d`

### 2022-01-03 scripts/pretrain.sh

@@ -1,5 +1,5 @@
 # Initalize environment
-dataset_dir=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh
+dataset_dir=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh_v2
 source .venv/bin/activate
 
 # Run training

### 2022-01-03 scripts/experiments/pretrain/entropy.sh
adds entropy training script

@@ -1,0 +1,15 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/entropy \
+    gpus=[0] \
+    max_time="01:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=entropy

### 2022-01-03 scripts/experiments/s3dis/minkowski-pretrained-new-loss.sh
updates pretrained training scripts

@@ -3,19 +3,20 @@ python src/main.py \
     dataset=s3dis \
     model=minkowski \
     tasks=["train","eval"] \
-    hydra.run.dir=outputs/pretrain/new-loss \
-    gpus=1 \
+    hydra.run.dir=outputs/pretrain/5cm-new-loss \
+    gpus=[1] \
     dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
     dataset.voxel_size=0.05 \
     model.optimizer.type=SGD \
-    model.optimizer.lr=0.1 \
+    model.optimizer.lr=0.3 \
     model.scheduler.type=PolyLR \
     model.scheduler.poly_power=0.9 \
     model.scheduler.max_iter=20000 \
     model.scheduler.interval=step \
     model.scheduler.frequency=100 \
-    max_time="00:12:00:00" \
-    max_epochs=400 \
-    check_val_every_n_epoch=10 \
+    max_time="02:00:00:00" \
+    max_epochs=4000 \
+    check_val_every_n_epoch=20 \
     model.train.train_workers=10 \
-    pretrain_checkpoint=\"epoch=248-step=9206-val_loss=6.46.ckpt\"
+    pretrain_checkpoint=\"new_loss.ckpt\"

### 2022-01-03 scripts/experiments/s3dis/minkowski-pretrained-old-loss.sh

@@ -4,9 +4,9 @@ python src/main.py \
     model=minkowski \
     tasks=["train","eval"] \
     hydra.run.dir=outputs/pretrain/5cm-extra-slow \
-    gpus=1 \
-    dataset.batch_size=16 \
-    dataset.accumulate_grad_batches=3 \
+    gpus=[0] \
+    dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
     dataset.voxel_size=0.05 \
     model.optimizer.type=SGD \
     model.optimizer.lr=0.2 \
@@ -17,6 +17,6 @@ python src/main.py \
     model.scheduler.frequency=10 \
     max_time="02:00:00:00" \
     max_epochs=2000 \
-    check_val_every_n_epoch=10 \
+    check_val_every_n_epoch=20 \
     model.train.train_workers=10 \
-    pretrain_checkpoint=\"epoch=270-step=19999-val_loss=6.47.ckpt\"
+    pretrain_checkpoint=\"old_loss.ckpt\"

### 2022-01-03 scripts/experiments/s3dis/pretrain-og-best.sh

@@ -5,7 +5,7 @@ python src/main.py \
     tasks=["train","eval"] \
     hydra.run.dir=outputs/pretrain/5cm-extra-slow \
     gpus=1 \
-    dataset.batch_size=6 \
+    dataset.batch_size=16 \
     dataset.accumulate_grad_batches=3 \
     dataset.voxel_size=0.05 \
     model.optimizer.type=SGD \

### 2022-01-03 doc/contrastive-learning.tex
adds some pictures for clustering
adds table with first set of pretraining results

@@
 Prototypical Contrastive Learning \cite{li_prototypical_2021}, BYOL \cite{grill_bootstrap_nodate} or SwAV \cite{caron_unsupervised_2021}
+
+Honestly still confused by some of these methods.
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.35]{images/swav.png}
+        \caption{Visualization of SwAV \cite{caron_unsupervised_2021}}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.35]{images/prototypes.png}
+        \caption{Visualization of PCL training framework \cite{li_prototypical_2021}}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.35]{images/byol.png}
+        \caption{Visualization of BYOL \cite{grill_bootstrap_nodate}}
+    \end{center}
+\end{figure}
 
@@
 
+\section{Results}
+
+\subsection{Experimental Setup}
+
+hyper-parameters:
+
+batch size: 16
+accumulate gradient: 3
+learning rate: 0.2
+iterations: 4000
+
+\subsection{Semantic Segmentation Negative Sampling}
+
+\begin{center}
+    \begin{table}[h!]
+        \resizebox{\textwidth}{!}{
+            \begin{tabular}{ c | c c c c c c c c c c c c c | c }
+                              & ceiling        & floor          & wall           & beam           & column         & window         & door           & table          & chair          & sofa           & bookcase       & board          & clutter        & mIOU           \\
+                \hline
+                baseline      & 0.869          & 0.95           & 0.798          & \textbf{0.001} & 0.253          & \textbf{0.585} & 0.510          & 0.731          & \textbf{0.841} & 0.391          & 0.715          & \textbf{0.725} & 0.534          & 0.608          \\
+                PointContrast & \textbf{0.895} & \textbf{0.961} & \textbf{0.817} & 0              & \textbf{0.350} & 0.561          & \textbf{0.671} & \textbf{0.737} & 0.834          & \textbf{0.572} & \textbf{0.727} & 0.699          & \textbf{0.554} & \textbf{0.644}
+            \end{tabular}
+        }
+        \caption{S3DIS pretraining results on Semantic Segmentation}
+    \end{table}
+\end{center}
+
+\pagebreak
+
 \newpage

### 2022-01-03 scripts/experiments/s3dis/minkowski.sh
updates scripts

@@ -5,8 +5,8 @@ python src/main.py \
     tasks=["train","eval"] \
     hydra.run.dir=outputs/s3dis/minkowski-5cm-extra-slow \
     gpus=1 \
-    dataset.batch_size=16 \
-    dataset.accumulate_grad_batches=3 \
+    dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
     dataset.voxel_size=0.05 \
     model.optimizer.type=SGD \
     model.optimizer.lr=0.2 \
@@ -16,6 +16,6 @@ python src/main.py \
     model.scheduler.interval=step \
     model.scheduler.frequency=100 \
     max_time="02:00:00:00" \
-    max_epochs=600 \
+    max_epochs=1200 \
     check_val_every_n_epoch=10 \
     model.train.train_workers=10 

### 2022-01-03 src/models/minkowski/trainer.py
updates delta loss

@@ -1,6 +1,7 @@
 import logging
 from omegaconf import DictConfig
 import functools
+import itertools
 
 import torch
 import torch.nn as nn
@@ -90,7 +91,7 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         elif cfg.model.net.loss == "entropy":
             self.loss_fn = self.loss_fn_entropy
         elif cfg.model.net.loss == "delta":
-            self.loss_fn == self.loss_fn_delta
+            self.loss_fn = self.loss_fn_delta
         else:
             self.loss_fn = self.loss_fn_original
 
@@ -138,8 +139,7 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         q = torch.cat(qs, 0)
         k = torch.cat(ks, 0)
 
-        # TODO: Comment out for now but normalization should be done here
-        # # normalize to unit vectors
+        # normalize to unit vectors
         q = q / torch.norm(q, p=2, dim=1, keepdim=True)
         k = k / torch.norm(k, p=2, dim=1, keepdim=True)
 
@@ -206,10 +206,8 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
 
     def loss_fn_delta(self, batch, output):
         tau = 0.4
-        max_pos = 50
+        max_pos = 200
         n = 4092
-
-        output = output.output
 
         # Get all positive and negative pairs
         qs, ks = [], []
@@ -242,42 +240,38 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         q = q / torch.norm(q, p=2, dim=1, keepdim=True)
         k = k / torch.norm(k, p=2, dim=1, keepdim=True)
 
-        batch_indices = np.zeros(q.shape[0])
-        old_pos = 0
-        for i in range(len(qs)):
-            batch_indices[old_pos : old_pos + qs[i].shape[0]] = i
-            old_pos = qs[i].shape[0]
-
         if q.shape[0] > max_pos:
             inds = np.random.choice(q.shape[0], max_pos, replace=False)
             q = q[inds]
             k = k[inds]
             q_fpfh = q_fpfh[inds]
             k_fpfh = k_fpfh[inds]
-            batch_indices = batch_indices[inds]
 
         pos = torch.exp(torch.sum(q * k, dim=-1) / tau)
         combined = torch.exp(torch.mm(q, k.t().contiguous()) / tau)
 
-        import itertools
-
[…]

### 2022-01-04 scripts/experiments/s3dis/minkowski-pretrained-projection-head.sh
deletes old scripts

@@ -1,12 +1,0 @@
-python src/main.py \
-    dataset_dir=$dataset_dir \
-    dataset=s3dis \
-    model=minkowski \
-    tasks=["train","eval"] \
-    hydra.run.dir=outputs/pretrain/projection-head \
-    gpus=1 \
-    max_time="00:12:00:00" \
-    max_epochs=400 \
-    check_val_every_n_epoch=10 \
-    model.train.train_workers=10 \
-    pretrain_checkpoint=\"epoch=389-step=14429-val_loss=6.58.ckpt\"

### 2022-01-04 scripts/experiments/s3dis/minkowski-pretrained-supervised.sh

@@ -1,11 +1,0 @@
-python src/main.py \
-    dataset_dir=$dataset_dir \
-    dataset=s3dis \
-    model=minkowski \
-    tasks=["train","eval"] \
-    hydra.run.dir=outputs/s3dis/minkowski-2cm-pretrained-supervised \
-    gpus=1 \
-    max_epochs=400 \
-    check_val_every_n_epoch=10 \
-    model.train.train_workers=10 \
-    supervised_pretrain_checkpoint=last.ckpt

### 2022-01-04 scripts/experiments/s3dis/minkowski-pretrained.sh

@@ -1,11 +1,0 @@
-python src/main.py \
-    dataset_dir=$dataset_dir \
-    dataset=s3dis \
-    model=minkowski \
-    tasks=["train","eval"] \
-    hydra.run.dir=outputs/s3dis/minkowski-2cm-pretrained \
-    gpus=1 \
-    max_epochs=400 \
-    check_val_every_n_epoch=10 \
-    model.train.train_workers=10 \
-    pretrain_checkpoint=\"epoch=389-step=14429-val_loss=6.58.ckpt\"

### 2022-01-04 scripts/experiments/s3dis/minkowski-pretrained-entropy.sh
adds entropy script

@@ -1,0 +1,23 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=s3dis \
+    model=minkowski \
+    tasks=["train","eval"] \
+    hydra.run.dir=outputs/pretrain/entropy \
+    gpus=[0] \
+    dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
+    dataset.voxel_size=0.05 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.2 \
+    model.scheduler.type=PolyLR \
+    model.scheduler.poly_power=0.9 \
+    model.scheduler.max_iter=20000 \
+    model.scheduler.interval=step \
+    model.scheduler.frequency=10 \
+    max_time="01:00:00:00" \
+    max_epochs=2000 \
+    max_steps=4000 \
+    check_val_every_n_epoch=20 \
+    model.train.train_workers=10 \
+    pretrain_checkpoint=\"entropy.ckpt\"

### 2022-01-04 scripts/experiments/s3dis/minkowski.sh
modifies base minkowski script

@@ -3,10 +3,10 @@ python src/main.py \
     dataset=s3dis \
     model=minkowski \
     tasks=["train","eval"] \
-    hydra.run.dir=outputs/s3dis/minkowski-5cm-extra-slow \
-    gpus=1 \
-    dataset.batch_size=16 \
-    dataset.accumulate_grad_batches=3 \
+    hydra.run.dir=outputs/s3dis/minkowski \
+    gpus=[0] \
+    dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
     dataset.voxel_size=0.05 \
     model.optimizer.type=SGD \
     model.optimizer.lr=0.2 \
@@ -16,6 +16,6 @@ python src/main.py \
     model.scheduler.interval=step \
     model.scheduler.frequency=100 \
     max_time="02:00:00:00" \
-    max_epochs=600 \
+    max_epochs=2000 \
     check_val_every_n_epoch=10 \
     model.train.train_workers=10 

### 2022-01-04 src/models/minkowski/dataset.py
removes old getitem methods from dataset

@@ -256,174 +256,6 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
             "quantized_frames": quantized_frames,
         }
 
-    def fpfh__getitem__(self, index):
-
-        scene = self.scenes[index].load_measurements()
-
-        # pick matching scenes at random
-        frame1 = random.choice(list(scene.matching_frames_map.keys()))
-        frame2 = random.choice(scene.matching_frames_map[frame1])
-
-        correspondences = scene.correspondance_map[frame1][frame2]
-
-        frame1 = scene.get_measurement(frame1)
-        frame2 = scene.get_measurement(frame2)
-
-        quantized_frames = []
-        random_scale = np.random.uniform(*self.scale_range)
-        for frame in [frame1, frame2]:
-
-            # Extract data
-            xyz = np.ascontiguousarray(frame.points)
-            features = torch.from_numpy(frame.point_colors)
-            fpfh = frame.fpfh
-            entropies = torch.from_numpy(frame.entropies)
-
-            # apply a random scalling
-            xyz *= random_scale
-
-            # Randomly rotate each frame
-            xyz = xyz - xyz.mean(0)
-            xyz, features, _ = self.augmentations(xyz, features, None)
-
-            # Voxelize input
-            discrete_coords, mapping = ME.utils.sparse_quantize(
-                coordinates=xyz, quantization_size=self.voxel_size, return_index=True
-            )
-
-            unique_feats = features[mapping]
-
-            # Get the point to voxel mapping
-            mapping = {
-                point_ind: voxel_ind
-                for voxel_ind, point_ind in enumerate(mapping.numpy())
-            }
-
-            # Append to quantized frames
-            quantized_frames.append(
-                {
-                    "discrete_coords": discrete_coords,
-                    "unique_feats": unique_feats,
-                    "mapping": mapping,
-                    "fpfh": fpfh,
-                    "entropies": entropies,
-                }
-            )
-
-        # Remap the correspondances into voxel world
-        mapping1 = quantized_frames[0]["mapping"]
-        mapping2 = quantized_frames[1]["mapping"]
-        correspondences = [
-            {
-                "frame1": {
-                    "voxel_inds": mapping1[k],
-                    "fpfh": quantized_frames[0]["fpfh"][k],
-                    "entropies": quantized_frames[0]["entropies"][k],
-                },
-                "frame2": {
-                    "voxel_inds": mapping2[v],
[…]

### 2022-01-04 scripts/experiments/pretrain/delta.sh
adds delta pre-training script

@@ -1,0 +1,15 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/delta \
+    gpus=[0] \
+    max_time="01:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=delta

### 2022-01-04 src/models/minkowski/trainer.py
adds clustering loss

@@ -2,6 +2,7 @@ import logging
 from omegaconf import DictConfig
 import functools
 import itertools
+import pickle
 
 import torch
 import torch.nn as nn
@@ -92,6 +93,9 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
             self.loss_fn = self.loss_fn_entropy
         elif cfg.model.net.loss == "delta":
             self.loss_fn = self.loss_fn_delta
+        elif cfg.model.net.loss == "cluster":
+            self.clusters = pickle.load(open("clustering.pkl", "rb"))
+            self.loss_fn = self.loss_fn_cluster
         else:
             self.loss_fn = self.loss_fn_original
 
@@ -268,6 +272,73 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
             # select the negative values
             select_indices = torch.tensor(
                 np.where(distances[ind] > 7.0)[0], device=q.device
+            )
+            neg = neg.index_select(1, select_indices)
+
+            # Compute the mean and normalize by n
+            Ng[ind] = neg.mean(dim=-1) * n
+
+        loss = (-torch.log(pos / (pos + Ng))).mean()
+
+        return loss
+
+    def loss_fn_cluster(self, batch, output):
+        tau = 0.4
+        max_pos = 2 * 4092
+        n = 4092
+
+        # Get all positive and negative pairs
+        qs, ks = [], []
+        qs_fpfh = []
+        for i, matches in enumerate(batch.correspondences):
+            voxel_indices_1 = [match["frame1"]["voxel_inds"] for match in matches]
+            voxel_indices_2 = [match["frame2"]["voxel_inds"] for match in matches]
+
+            fpfh_1 = [match["frame1"]["fpfh"] for match in matches]
+            # fpfh_2 = [match["frame2"]["fpfh"] for match in matches]
+
+            output_batch_1 = output.features_at(2 * i)
+            output_batch_2 = output.features_at(2 * i + 1)
+            q = output_batch_1[voxel_indices_1]
+            k = output_batch_2[voxel_indices_2]
+
+            qs.append(q)
+            ks.append(k)
+
+            qs_fpfh.append(np.array(fpfh_1))
+
+        q = torch.cat(qs, 0)
+        k = torch.cat(ks, 0)
+
+        q_fpfh = np.concatenate(qs_fpfh)
+
+        # normalize to unit vectors
+        q = q / torch.norm(q, p=2, dim=1, keepdim=True)
[…]

### 2022-01-05 src/visualize_features.py
updates visualize features script

@@ -160,98 +160,77 @@ def entropy(feature, base=None):
 import random
 
 
+def compute_fpfh(pcd):
+    pcd.estimate_normals(o3d.geometry.KDTreeSearchParamHybrid(radius=3, max_nn=100))
+
+    search_param = o3d.geometry.KDTreeSearchParamHybrid(radius=7, max_nn=200)
+    fpfh = o3d.pipelines.registration.compute_fpfh_feature(pcd, search_param)
+
+    return fpfh.data.T
+
+
+import pickle
+
+
 @hydra.main(config_path="config", config_name="config")
 def test_fpfh(cfg: DictConfig):
     """Test FPFH features."""
 
-    pl.seed_everything(62, workers=True)
+    pl.seed_everything(70, workers=True)
 
     # load a data interface
     data_interface_factory = DataInterfaceFactory(cfg)
     data_interface = data_interface_factory.get_interface()
 
+    # Load in a scene
+    model_factory = ModelFactory(cfg, data_interface)
+    dataset_type = model_factory.get_dataset_type()
+    dataset = dataset_type(data_interface.val_data[:100], cfg)
+    collate_fn = dataset.collate
+
     # # get a couple of values
-    # voxel_size = 0.02
     # fpfh_features = []
     # for i in range(50):
-
-    #     # Load in a scene
-    #     model_factory = ModelFactory(cfg, data_interface)
-    #     dataset_type = model_factory.get_dataset_type()
-    #     dataset = dataset_type(data_interface.pretrain_val_data, cfg)
-    #     collate_fn = dataset.collate
 
     #     scene_ind = random.randint(0, len(dataset) - 1)
     #     scene = collate_fn([dataset[scene_ind]])
 
-    #     # first_frame = np.where(scene.points[:, 0] == 0)[0]
-    #     # points = scene.points[first_frame, 1:4] * voxel_size
+    #     points = scene.points[:, 1:4]
 
-    #     points = scene.points[:, 1:4] * voxel_size
-
-    #     # points = points[:, 1:4] * voxel_size
     #     pcd = o3d.geometry.PointCloud()
     #     pcd.points = o3d.utility.Vector3dVector(points)
 
     #     # o3d.visualization.draw_geometries([pcd])
 
-    #     # compute the features
-    #     pcd = pcd.voxel_down_sample(voxel_size=voxel_size)
-    #     pcd.estimate_normals(
[…]

### 2022-01-05 scripts/experiments/s3dis/minkowski-pretrained-delta.sh
adds delta script

@@ -1,0 +1,23 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=s3dis \
+    model=minkowski \
+    tasks=["train","eval"] \
+    hydra.run.dir=outputs/pretrain/delta \
+    gpus=[0] \
+    dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
+    dataset.voxel_size=0.05 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.2 \
+    model.scheduler.type=PolyLR \
+    model.scheduler.poly_power=0.9 \
+    model.scheduler.max_iter=20000 \
+    model.scheduler.interval=step \
+    model.scheduler.frequency=10 \
+    max_time="01:00:00:00" \
+    max_epochs=2000 \
+    max_steps=4000 \
+    check_val_every_n_epoch=20 \
+    model.train.train_workers=10 \
+    pretrain_checkpoint=\"delta.ckpt\"

### 2022-01-05 src/models/minkowski/dataset.py
fixes data augmentation bug

@@ -334,10 +334,10 @@ class MinkowskiDataset(SegmentationDataset):
         features = torch.from_numpy(scene.features)
         labels = np.array([scene.semantic_labels, scene.instance_labels]).T
 
-        # if self.is_test:
-        #     xyz, features, labels = self.test_augmentations(xyz, features, labels)
-        # else:
-        #     xyz, features, labels = self.augmentations(xyz, features, labels)
+        if self.is_test:
+            xyz, features, labels = self.test_augmentations(xyz, features, labels)
+        else:
+            xyz, features, labels = self.augmentations(xyz, features, labels)
 
         coords, feats, labels = ME.utils.sparse_quantize(
             xyz,

### 2022-01-05 scripts/experiments/pretrain/cluster.sh
updates cluster loss and script

@@ -1,0 +1,15 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/clustering \
+    gpus=[0] \
+    max_time="01:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=cluster

### 2022-01-05 src/models/minkowski/trainer.py

@@ -284,7 +284,7 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
 
     def loss_fn_cluster(self, batch, output):
         tau = 0.4
-        max_pos = 2 * 4092
+        max_pos = int(4092 / 4)
         n = 4092
 
         # Get all positive and negative pairs
@@ -295,7 +295,6 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
             voxel_indices_2 = [match["frame2"]["voxel_inds"] for match in matches]
 
             fpfh_1 = [match["frame1"]["fpfh"] for match in matches]
-            # fpfh_2 = [match["frame2"]["fpfh"] for match in matches]
 
             output_batch_1 = output.features_at(2 * i)
             output_batch_2 = output.features_at(2 * i + 1)

### 2022-01-10 src/datasets/measurements.py
adds cluster calculation to preprocessing and loss

@@ -47,7 +47,14 @@ def get_point_cloud(depth, intrinsics):
 class SceneMeasurement:
     """Stores sensor measurements for a single frame"""
 
-    def __init__(self, directory: Path, output_dir: Path, info: dict, frame_id: str):
+    def __init__(
+        self,
+        directory: Path,
+        output_dir: Path,
+        info: dict,
+        frame_id: str,
+        cluster_model,
+    ):
         """
         Load all measurements of a single frame from a scene catpure
 
@@ -147,6 +154,9 @@ class SceneMeasurement:
         entropies = entropies - entropies.min()
         entropies = entropies / entropies.max()
         self.entropies = entropies
+
+        # Compute clusters
+        self.clusters = cluster_model.predict(self.fpfh)
 
         # save all processed data to file
         self.save_to_file(output_dir)
@@ -253,8 +263,13 @@ class SceneMeasurements:
         self.label_id_to_colour = {i: self.get_random_colour() for i in range(100)}
         self.instance_id_to_colour = {i: self.get_random_colour() for i in range(100)}
 
+        # Load cluster model
+        cluster_model = pickle.load(open("clustering.pkl", "rb"))
+
         measurements = [
-            SceneMeasurement(directory, output_dir, self.info, f"{frame:06d}")
+            SceneMeasurement(
+                directory, output_dir, self.info, f"{frame:06d}", cluster_model
+            )
             for frame in range(0, int(self.info["m_frames.size"]), frame_skip)
         ]
         self.num_measurements = len(measurements)

### 2022-01-10 src/models/minkowski/dataset.py

@@ -187,6 +187,7 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
             features = torch.from_numpy(frame.point_colors)
             fpfh = frame.fpfh
             entropies = frame.entropies
+            clusters = frame.clusters
 
             # VISUALIZATION
             # import copy
@@ -227,6 +228,7 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
                     "mapping": mapping,
                     "fpfh": fpfh,
                     "entropies": entropies,
+                    "clusters": clusters,
                 }
             )
 
@@ -239,11 +241,13 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
                     "voxel_inds": mapping1[k],
                     "fpfh": quantized_frames[0]["fpfh"][k],
                     "entropies": quantized_frames[0]["entropies"][k],
+                    "clusters": quantized_frames[0]["clusters"][k],
                 },
                 "frame2": {
                     "voxel_inds": mapping2[v],
                     "fpfh": quantized_frames[1]["fpfh"][v],
                     "entropies": quantized_frames[1]["entropies"][v],
+                    "clusters": quantized_frames[1]["clusters"][v],
                 },
             }
             for k, v in correspondences.items()

### 2022-01-10 src/models/minkowski/trainer.py

@@ -94,7 +94,6 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         elif cfg.model.net.loss == "delta":
             self.loss_fn = self.loss_fn_delta
         elif cfg.model.net.loss == "cluster":
-            self.clusters = pickle.load(open("clustering.pkl", "rb"))
             self.loss_fn = self.loss_fn_cluster
         else:
             self.loss_fn = self.loss_fn_original
@@ -284,17 +283,17 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
 
     def loss_fn_cluster(self, batch, output):
         tau = 0.4
-        max_pos = int(4092 / 4)
+        max_pos = int(4092 / 1)
         n = 4092
 
         # Get all positive and negative pairs
         qs, ks = [], []
-        qs_fpfh = []
+        qs_clusters = []
         for i, matches in enumerate(batch.correspondences):
             voxel_indices_1 = [match["frame1"]["voxel_inds"] for match in matches]
             voxel_indices_2 = [match["frame2"]["voxel_inds"] for match in matches]
 
-            fpfh_1 = [match["frame1"]["fpfh"] for match in matches]
+            clusters = [match["frame1"]["clusters"] for match in matches]
 
             output_batch_1 = output.features_at(2 * i)
             output_batch_2 = output.features_at(2 * i + 1)
@@ -304,12 +303,12 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
             qs.append(q)
             ks.append(k)
 
-            qs_fpfh.append(np.array(fpfh_1))
+            qs_clusters.append(np.array(clusters))
 
         q = torch.cat(qs, 0)
         k = torch.cat(ks, 0)
 
-        q_fpfh = np.concatenate(qs_fpfh)
+        q_clusters = np.concatenate(qs_clusters)
 
         # normalize to unit vectors
         q = q / torch.norm(q, p=2, dim=1, keepdim=True)
@@ -319,9 +318,9 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
             inds = np.random.choice(q.shape[0], max_pos, replace=False)
             q = q[inds]
             k = k[inds]
-            q_fpfh = q_fpfh[inds]
+            q_clusters = q_clusters[inds]
 
-        cluster_pred = self.clusters.predict(q_fpfh)
+        # cluster_pred = self.clusters.predict(q_fpfh)
 
         pos = torch.exp(torch.sum(q * k, dim=-1) / tau)
[…]

### 2022-01-10 scripts/preprocess/scannet.sh
updates preprocessing scripts

@@ -1,10 +1,11 @@
 export SCANNET_DIR=/media/starslab/datasets/scannet
-export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh_v2
+export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh_cluster
 export HYDRA_FULL_ERROR=1
 
 preprocess() {
     filename=$1
     python src/preprocess.py \
+        hydra.run.dir=outputs/preprocess \
         dataset_dir=$SCANNET_DIR \
         output_dir=$OUTPUT_DIR \
         dataset=scannet \
@@ -15,4 +16,4 @@ preprocess() {
 export -f preprocess
 
 # preprocess
-parallel -j 16 --linebuffer time preprocess ::: `find $SCANNET_DIR/scans/ -mindepth 1 -maxdepth 1 -type d`
+parallel -j 4 --linebuffer time preprocess ::: `find $SCANNET_DIR/scans/ -mindepth 1 -maxdepth 1 -type d`

### 2022-01-10 scripts/pretrain.sh

@@ -1,5 +1,5 @@
 # Initalize environment
-dataset_dir=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh_v2
+dataset_dir=/media/starslab/datasets/scannet_preprocessed_2cm_fpfh_cluster
 source .venv/bin/activate
 
 # Run training

### 2022-01-10 src/models/minkowski/trainer.py
adds debiased and hard negative sampling

@@ -95,6 +95,10 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
             self.loss_fn = self.loss_fn_delta
         elif cfg.model.net.loss == "cluster":
             self.loss_fn = self.loss_fn_cluster
+        elif cfg.model.net.loss == "debiased":
+            self.loss_fn = self.loss_fn_debiased
+        elif cfg.model.net.loss == "hard":
+            self.loss_fn = self.loss_fn_hard
         else:
             self.loss_fn = self.loss_fn_original
 
@@ -408,6 +412,124 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         out = out.squeeze().contiguous()
 
         return self.criterion(out, labels)
+
+    def loss_fn_debiased(self, batch, output):
+        tau = 0.4
+        max_pos = 4092
+        n = 4092
+        tau_plus = 0.1  # class probability
+
+        # Get all positive and negative pairs
+        qs, ks = [], []
+        for i, matches in enumerate(batch.correspondences):
+            voxel_indices_1 = [match["frame1"]["voxel_inds"] for match in matches]
+            voxel_indices_2 = [match["frame2"]["voxel_inds"] for match in matches]
+
+            output_batch_1 = output.features_at(2 * i)
+            output_batch_2 = output.features_at(2 * i + 1)
+            q = output_batch_1[voxel_indices_1]
+            k = output_batch_2[voxel_indices_2]
+
+            qs.append(q)
+            ks.append(k)
+
+        q = torch.cat(qs, 0)
+        k = torch.cat(ks, 0)
+
+        # normalize to unit vectors
+        q = q / torch.norm(q, p=2, dim=1, keepdim=True)
+        k = k / torch.norm(k, p=2, dim=1, keepdim=True)
+
+        if q.shape[0] > max_pos:
+            inds = np.random.choice(q.shape[0], max_pos, replace=False)
+            q = q[inds]
+            k = k[inds]
+
+        pos = torch.exp(torch.sum(q * k, dim=-1) / tau)
+        combined = torch.exp(torch.mm(q, k.t().contiguous()) / tau)
+
+        Ng = torch.zeros(q.shape[0], device=q.device)
+        for ind in range(q.shape[0]):
+
+            # select the negative values
+            neg = combined.index_select(0, torch.tensor([ind], device=q.device))
+
+            # Select all but the same index
+            select_indices = torch.tensor(
[…]

### 2022-01-10 scripts/experiments/pretrain/hard.sh
adds hard and debiased sampling scripts

@@ -1,0 +1,15 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/hard \
+    gpus=[1] \
+    max_time="01:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=hard

### 2022-01-17 scripts/experiments/pretrain/moco-easy.sh
adds hard, easy, medium negative sampling

@@ -1,0 +1,16 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/moco-easy \
+    gpus=[0] \
+    max_time="03:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=mixing \
+    model.pretrain.loss.difficulty=easy

### 2022-01-17 scripts/experiments/pretrain/moco-hard.sh

@@ -1,0 +1,16 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/moco-hard \
+    gpus=[0] \
+    max_time="03:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=mixing \
+    model.pretrain.loss.difficulty=hard

### 2022-01-17 scripts/experiments/pretrain/moco-medium.sh

@@ -1,0 +1,16 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/moco-medium \
+    gpus=[0] \
+    max_time="03:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=mixing \
+    model.pretrain.loss.difficulty=medium

### 2022-01-17 scripts/experiments/pretrain/moco.sh

@@ -1,0 +1,15 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    model=minkowski \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/moco \
+    gpus=[1] \
+    max_time="03:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.loss=mixing

### 2022-01-17 src/util/utils.py

@@ -32,6 +32,7 @@ class NCELossMoco(nn.Module):
         )
         self.dim = config.net.model_n_out
         self.T = config.pretrain.loss.temperature
+        self.difficulty = config.pretrain.loss.difficulty
 
         self.register_buffer("queue", torch.randn(self.dim, self.K))
         self.queue = nn.functional.normalize(self.queue, dim=0)
@@ -71,6 +72,20 @@ class NCELossMoco(nn.Module):
             "nc,ck->nk", [normalized_output1, self.queue.clone().detach()]
         )
 
+        # sort based off of hardness to each positive sample
+        # select based off of hardest indices
+        if self.difficulty == "hard":
+            values, indices = torch.sort(l_neg, dim=1, descending=True)
+            l_neg = values[:, : 3 * 4092]
+        elif self.difficulty == "medium":
+            values, indices = torch.sort(l_neg, dim=1, descending=True)
+            l_neg = values[:, 3 * 4092 : 6 * 4092]
+        elif self.difficulty == "easy":
+            values, indices = torch.sort(l_neg, dim=1, descending=False)
+            l_neg = values[:, : 3 * 4092]
+        else:  # mixing
+            values, indices = torch.sort(l_neg, dim=1, descending=True)
+
         # logits: Nx(1+K)
         logits = torch.cat([l_pos, l_neg], dim=1)
 
@@ -83,6 +98,34 @@ class NCELossMoco(nn.Module):
         labels = torch.zeros(logits.shape[0], device=logits.device, dtype=torch.int64)
 
         return self.xe_criterion(torch.squeeze(logits), labels)
+
+    # def forward(self, output):
+
+    #     normalized_output1 = nn.functional.normalize(output[0], dim=1, p=2)
+    #     normalized_output2 = nn.functional.normalize(output[1], dim=1, p=2)
+
+    #     # positive logits: Nx1
+    #     l_pos = torch.einsum(
+    #         "nc,nc->n", [normalized_output1, normalized_output2]
+    #     ).unsqueeze(-1)
+
+    #     # negative logits: NxK
+    #     l_neg = torch.einsum(
+    #         "nc,ck->nk", [normalized_output1, self.queue.clone().detach()]
+    #     )
+
+    #     # logits: Nx(1+K)
+    #     logits = torch.cat([l_pos, l_neg], dim=1)
+
+    #     # apply temperature
+    #     logits /= self.T
+
+    #     # Why do this at the end?
[…]

### 2022-01-17 src/util/utils.py
first iteration of mixing (need to fix issue)

@@ -84,7 +84,29 @@ class NCELossMoco(nn.Module):
             values, indices = torch.sort(l_neg, dim=1, descending=False)
             l_neg = values[:, : 3 * 4092]
         else:  # mixing
-            values, indices = torch.sort(l_neg, dim=1, descending=True)
+
+            # TODO: Don't do it on logits!!
+
+            # Select hard examples to mix
+            values, indicies = torch.sort(l_neg, dim=1, descending=True)
+            hard_neg = values[:, : 2 * 4092]
+
+            # Select mixing coefficients and random i,j values
+            # note that these are same across all positive samples
+            alpha_s = torch.rand((hard_neg.shape[1],), device=l_neg.device)
+            i_s = torch.randint(hard_neg.shape[1], (hard_neg.shape[1],))
+            j_s = torch.randint(hard_neg.shape[1], (hard_neg.shape[1],))
+
+            # Perform mixing
+            mixed_neg = torch.mul(alpha_s, hard_neg[:, i_s]) + torch.mul(
+                1 - alpha_s, hard_neg[:, j_s]
+            )
+
+            # Normalize new samples
+            mixed_neg = nn.functional.normalize(mixed_neg, dim=1, p=2)
+
+            # Add mixed samples to negatives samples
+            l_neg = torch.cat((l_neg, mixed_neg), 1)
 
         # logits: Nx(1+K)
         logits = torch.cat([l_pos, l_neg], dim=1)

### 2022-01-18 scripts/experiments/s3dis/minkowski-pretrained-moco.sh
adds moco training script

@@ -1,0 +1,23 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=s3dis \
+    model=minkowski \
+    tasks=["train","eval"] \
+    hydra.run.dir=outputs/pretrain/moco \
+    gpus=[0] \
+    dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
+    dataset.voxel_size=0.05 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.2 \
+    model.scheduler.type=PolyLR \
+    model.scheduler.poly_power=0.9 \
+    model.scheduler.max_iter=20000 \
+    model.scheduler.interval=step \
+    model.scheduler.frequency=10 \
+    max_time="01:00:00:00" \
+    max_epochs=2000 \
+    max_steps=4000 \
+    check_val_every_n_epoch=20 \
+    model.train.train_workers=10 \
+    pretrain_checkpoint=\"moco.ckpt\"

### 2022-01-18 scripts/experiments/s3dis/minkowski-pretrained-moco.sh
updates moco script

@@ -20,4 +20,5 @@ python src/main.py \
     max_steps=4000 \
     check_val_every_n_epoch=20 \
     model.train.train_workers=10 \
+    model.net.loss=mixing \
     pretrain_checkpoint=\"moco.ckpt\"

### 2022-01-18 src/models/minkowski/dataset.py
adds preprocessing with just frames

@@ -185,6 +185,78 @@ class MinkowskiEntropyPretrainDataset(PretrainDataset):
             # Extract data
             xyz = np.ascontiguousarray(frame.points)
             features = torch.from_numpy(frame.point_colors)
+
+            # apply a random scalling
+            xyz *= random_scale
+
+            # Randomly rotate each frame
+            xyz = xyz - xyz.mean(0)
+            xyz, features, _ = self.augmentations(xyz, features, None)
+
+            # Voxelize input
+            discrete_coords, mapping = ME.utils.sparse_quantize(
+                coordinates=xyz, quantization_size=self.voxel_size, return_index=True
+            )
+
+            unique_feats = features[mapping]
+
+            # Get the point to voxel mapping
+            mapping = {
+                point_ind: voxel_ind
+                for voxel_ind, point_ind in enumerate(mapping.numpy())
+            }
+
+            # Append to quantized frames
+            quantized_frames.append(
+                {
+                    "discrete_coords": discrete_coords,
+                    "unique_feats": unique_feats,
+                    "mapping": mapping,
+                }
+            )
+
+        # Remap the correspondances into voxel world
+        mapping1 = quantized_frames[0]["mapping"]
+        mapping2 = quantized_frames[1]["mapping"]
+        correspondences = [
+            {
+                "frame1": {
+                    "voxel_inds": mapping1[k],
+                },
+                "frame2": {
+                    "voxel_inds": mapping2[v],
+                },
+            }
+            for k, v in correspondences.items()
+            if k in mapping1.keys() and v in mapping2.keys()
+        ]
+
+        # visualize_correspondances(quantized_frames, correspondences)
+        return {
+            "correspondences": correspondences,
+            "quantized_frames": quantized_frames,
+        }
+
+    def __getitem__fpfh(self, index):
+
+        scene = self.scenes[index].load_measurements()
+
+        # pick matching scenes at random
+        frame1 = random.choice(list(scene.matching_frames_map.keys()))
+        frame2 = random.choice(scene.matching_frames_map[frame1])
+
+        correspondences = scene.correspondance_map[frame1][frame2]
+
+        frame1 = scene.get_measurement(frame1)
+        frame2 = scene.get_measurement(frame2)
+
+        quantized_frames = []
+        random_scale = np.random.uniform(*self.scale_range)
+        for frame in [frame1, frame2]:
+
+            # Extract data
+            xyz = np.ascontiguousarray(frame.points)
[…]

### 2022-01-18 src/util/utils.py
adds first matrix version of mixing

@@ -68,9 +68,8 @@ class NCELossMoco(nn.Module):
         ).unsqueeze(-1)
 
         # negative logits: NxK
-        l_neg = torch.einsum(
-            "nc,ck->nk", [normalized_output1, self.queue.clone().detach()]
-        )
+        neg_features = self.queue.clone().detach()
+        l_neg = torch.einsum("nc,ck->nk", [normalized_output1, neg_features])
 
         # sort based off of hardness to each positive sample
         # select based off of hardest indices
@@ -89,24 +88,46 @@ class NCELossMoco(nn.Module):
 
             # Select hard examples to mix
             values, indicies = torch.sort(l_neg, dim=1, descending=True)
-            hard_neg = values[:, : 2 * 4092]
+            # cut_off = 2 * 4092
+            cut_off = 2
+            n_dim = neg_features.shape[0]
+            hard_neg = values[:, :cut_off]
+            hard_indices = indicies[:, :cut_off]
 
             # Select mixing coefficients and random i,j values
             # note that these are same across all positive samples
-            alpha_s = torch.rand((hard_neg.shape[1],), device=l_neg.device)
-            i_s = torch.randint(hard_neg.shape[1], (hard_neg.shape[1],))
-            j_s = torch.randint(hard_neg.shape[1], (hard_neg.shape[1],))
+            n_pos = hard_neg.shape[0]
+            n_neg = hard_neg.shape[1]
+            alpha_s = torch.rand((n_neg,), device=l_neg.device)
+            i_s = torch.randint(n_neg, (n_neg,))
+            j_s = torch.randint(n_neg, (n_neg,))
+
+            # indices
+            hard_indices_is = hard_indices[:, i_s].reshape(-1)
+            hard_indices_js = hard_indices[:, j_s].reshape(-1)
+
+            hard_neg_is = neg_features[:, hard_indices_is].T.reshape(n_pos, cut_off, -1)
+            hard_neg_js = neg_features[:, hard_indices_js].T.reshape(n_pos, cut_off, -1)
+
+            hard_neg_is = hard_neg_is.transpose(1, 2)
+            hard_neg_js = hard_neg_js.transpose(1, 2)
 
             # Perform mixing
-            mixed_neg = torch.mul(alpha_s, hard_neg[:, i_s]) + torch.mul(
-                1 - alpha_s, hard_neg[:, j_s]
+            # TODO: check that this is correct
[…]

### 2022-01-18 src/util/utils.py
adds checks to hard negative mixing

@@ -88,16 +88,14 @@ class NCELossMoco(nn.Module):
 
             # Select hard examples to mix
             values, indicies = torch.sort(l_neg, dim=1, descending=True)
-            # cut_off = 2 * 4092
-            cut_off = 2
+            cut_off = 4092
             n_dim = neg_features.shape[0]
-            hard_neg = values[:, :cut_off]
             hard_indices = indicies[:, :cut_off]
 
             # Select mixing coefficients and random i,j values
             # note that these are same across all positive samples
-            n_pos = hard_neg.shape[0]
-            n_neg = hard_neg.shape[1]
+            n_pos = hard_indices.shape[0]
+            n_neg = hard_indices.shape[1]
             alpha_s = torch.rand((n_neg,), device=l_neg.device)
             i_s = torch.randint(n_neg, (n_neg,))
             j_s = torch.randint(n_neg, (n_neg,))
@@ -119,12 +117,27 @@ class NCELossMoco(nn.Module):
                 1 - alpha_s, hard_neg_js
             )
 
+            # check
+            # mixed_neg_check = torch.zeros_like(hard_neg_is)
+            # for pos_ind in range(n_pos):
+            #     for neg_ind in range(n_neg):
+            #         mixed_neg_check[pos_ind, :, neg_ind] = (
+            #             alpha_s[neg_ind] * hard_neg_is[pos_ind, :, neg_ind]
+            #             + (1 - alpha_s[neg_ind]) * hard_neg_js[pos_ind, :, neg_ind]
+            #         )
+
             # Normalize new samples
             # TODO: check that its normalizing along correct axis
             mixed_neg = nn.functional.normalize(mixed_neg, dim=1, p=2)
 
+            # check
+            # check = torch.norm(mixed_neg, dim=1, p=2)
+
             # compute new logits
             l_new = torch.einsum("nc,nck->nk", [normalized_output1, mixed_neg])
+
+            # check
+            # test = normalized_output1[0].dot(mixed_neg[0, :, 0])
 
             # Add mixed samples to negatives samples
             l_neg = torch.cat((l_neg, l_new), 1)
@@ -141,34 +154,6 @@ class NCELossMoco(nn.Module):
         labels = torch.zeros(logits.shape[0], device=logits.device, dtype=torch.int64)
 
         return self.xe_criterion(torch.squeeze(logits), labels)
-
[…]

### 2022-01-24 scripts/experiments/pretrain/moco.sh
update moco scripts

@@ -6,7 +6,7 @@ python src/main.py \
     model.name=minkowski_moco \
     tasks=["pretrain"] \
     hydra.run.dir=outputs/pretrain/moco \
-    gpus=[1] \
+    gpus=[0] \
     max_time="01:00:00:00" \
     dataset.pretrain.batch_size=4 \
     dataset.pretrain.accumulate_grad_batches=4 \

### 2022-01-24 scripts/experiments/s3dis/minkowski-pretrained-moco.sh

@@ -2,6 +2,7 @@ python src/main.py \
     dataset_dir=$dataset_dir \
     dataset=s3dis \
     model=minkowski \
+    model.name=minkowski_moco \
     tasks=["train","eval"] \
     hydra.run.dir=outputs/pretrain/moco \
     gpus=[0] \
@@ -20,5 +21,4 @@ python src/main.py \
     max_steps=4000 \
     check_val_every_n_epoch=20 \
     model.train.train_workers=10 \
-    model.net.loss=mixing \
-    pretrain_checkpoint=\"moco.ckpt\"
+    pretrain_checkpoint=\"last.ckpt\"

### 2022-01-24 doc/contrastive-learning.tex
updates results with new batch sizes
adds entropy semantic segmentation eval results
adds delta results
adds cluster results
updates delta results
update cluster results
updates results

@@
 
-batch size: 16
-accumulate gradient: 3
-learning rate: 0.2
-iterations: 4000
+\begin{itemize}
+    \item batch size: 6
+    \item accumulate gradient: 8
+    \item learning rate: 0.2
+    \item iterations: 4000
+\end{itemize}
 
@@
                 \hline
-                baseline      & 0.869          & 0.95           & 0.798          & \textbf{0.001} & 0.253          & \textbf{0.585} & 0.510          & 0.731          & \textbf{0.841} & 0.391          & 0.715          & \textbf{0.725} & 0.534          & 0.608          \\
-                PointContrast & \textbf{0.895} & \textbf{0.961} & \textbf{0.817} & 0              & \textbf{0.350} & 0.561          & \textbf{0.671} & \textbf{0.737} & 0.834          & \textbf{0.572} & \textbf{0.727} & 0.699          & \textbf{0.554} & \textbf{0.644}
+                Baseline          & 0.862          & 0.947          & 0.811          & \textbf{0.002} & \textbf{0.356} & 0.551          & 0.590          & 0.745          & 0.751          & 0.452          & 0.719          & 0.606          & 0.514          & 0.608          \\
+                PointContrast     & \textbf{0.905} & \textbf{0.962} & \textbf{0.818} & 0              & 0.345          & \textbf{0.581} & \textbf{0.633} & \textbf{0.750} & \textbf{0.861} & \textbf{0.648} & \textbf{0.729} & \textbf{0.799} & \textbf{0.566} & \textbf{0.661} \\
+                \hline
+                Entropy           & \textbf{0.912} & 0.961          & \textbf{0.828} & 0              & 0.307          & 0.513          & 0.626          & 0.746          & 0.853          & 0.634          & \textbf{0.742} & 0.761          & \textbf{0.576} & 0.651          \\
+                Delta             & \textbf{0.924} & 0.959          & 0.817          & 0              & \textbf{0.388} & \textbf{0.611} & 0.563          & \textbf{0.770} & \textbf{0.867} & \textbf{0.667} & \textbf{0.742} & 0.635          & \textbf{0.587} & 0.656          \\
+                Custer            & \textbf{0.924} & 0.957          & 0.814          & 0              & \textbf{0.373} & 0.562          & \textbf{0.688} & \textbf{0.776} & 0.844          & 0.524          & 0.729          & 0.717          & 0.587          & 0.653          \\
+                \hline
+                Debiased          & 0.910          & 0.958          & 0.811          & \textbf{0.001} & 0.260          & 0.537          & 0.655          & 0.745          & 0.859          & 0.555          & 0.699          & 0.770          & 0.562          & 0.640          \\
+                Debiased++        & 0.901          & 0.961          & 0.815          & \textbf{0.008} & 0.286          & 0.586          & 0.683          & 0.745          & 0.855          & 0.599          & 0.716          & 0.779          & 0.554          & 0.653          \\
+                \hline
+                Sampling - Hard   & 0.904          & 0.959          & 0.824          & 0              & 0.361          & 0.579          & 0.659          & 0.731          & 0.837          & 0.505          & 0.697          & 0.711          & 0.557          & 0.640          \\
+                Sampling - Easy   & 0.666          & 0.952          & 0.779          & 0              & 0.201          & 0              & 0              & 0.740          & 0              & 0              & 0.690          & 0.784          & 0.526          & 0.411          \\
+                Sampling - Mixing & 0.895          & 0.962          & \textbf{0.820} & 0              & 0.277          & 0.576          & \textbf{0.706} & 0.747          & 0.849          & 0.567          & 0.724          & 0.769          & 0.562          & 0.650          \\
             \end{tabular}

### 2022-01-24 scripts/experiments/pretrain/byol.sh
adds BYOL

@@ -1,0 +1,17 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    dataset.name=scannetv2_pretrain \
+    model=minkowski \
+    model.name=minkowski_byol \
+    tasks=["pretrain"] \
+    hydra.run.dir=outputs/pretrain/byol \
+    gpus=[0] \
+    max_time="01:00:00:00" \
+    dataset.pretrain.batch_size=4 \
+    dataset.pretrain.accumulate_grad_batches=4 \
+    model.train.train_workers=8 \
+    check_val_every_n_epoch=10 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.1 \
+    model.net.model_n_out=64

### 2022-01-24 src/models/factory.py

@@ -9,6 +9,7 @@ from models.minkowski.trainer import (
     MinkowskiTrainer,
     MinkowskiBackboneTrainer,
     MinkowskiMocoBackboneTrainer,
+    MinkowskiBOYLBackboneTrainer,
 )
 from models.minkowski.dataset import (
     MinkowskiDataset,
@@ -20,7 +21,8 @@ from util.types import DataInterface
 pointgroup_name = "pointgroup"
 minkowski_name = "minkowski"
 moco_name = "minkowski_moco"
-supported_models = [pointgroup_name, minkowski_name, moco_name]
+byol_name = "minkowski_byol"
+supported_models = [pointgroup_name, minkowski_name, moco_name, byol_name]
 
 
 class ModelFactory:
@@ -51,8 +53,10 @@ class ModelFactory:
     def get_backbone_wrapper_type(self):
         if self.model_name == minkowski_name:
             return MinkowskiBackboneTrainer
-        if self.model_name == moco_name:
+        elif self.model_name == moco_name:
             return MinkowskiMocoBackboneTrainer
+        elif self.model_name == byol_name:
+            return MinkowskiBOYLBackboneTrainer
         else:
             raise RuntimeError(self.error_msg)
 

### 2022-01-24 src/models/factory.py
fix get_dataset_type

@@ -91,7 +91,7 @@ class ModelFactory:
         if self.model_name == pointgroup_name:
             raise NotImplementedError("No pointgroup")
             # return SpconvDataset
-        elif self.model_name == minkowski_name:
+        elif minkowski_name in self.model_name:
             return MinkowskiDataset
         else:
             raise RuntimeError(self.error_msg)

### 2022-01-25 scripts/experiments/s3dis/minkowski-pretrained-byol.sh
adds byol script

@@ -1,0 +1,25 @@
+python src/main.py \
+    dataset_dir=$dataset_dir \
+    dataset=s3dis \
+    model=minkowski \
+    model.name=minkowski_byol \
+    tasks=["train","eval"] \
+    hydra.run.dir=outputs/pretrain/byol \
+    gpus=[0] \
+    dataset.batch_size=6 \
+    dataset.accumulate_grad_batches=8 \
+    dataset.voxel_size=0.05 \
+    model.optimizer.type=SGD \
+    model.optimizer.lr=0.2 \
+    model.scheduler.type=PolyLR \
+    model.scheduler.poly_power=0.9 \
+    model.scheduler.max_iter=10000 \
+    model.scheduler.interval=step \
+    model.scheduler.frequency=10 \
+    max_time="01:00:00:00" \
+    max_epochs=2000 \
+    max_steps=4000 \
+    check_val_every_n_epoch=20 \
+    model.train.train_workers=6 \
+    model.net.model_n_out=64 \
+    pretrain_checkpoint=\"last.ckpt\"

### 2022-01-25 scripts/experiments/s3dis/minkowski-pretrained-moco-hard.sh
updates moco script

@@ -1,24 +1,0 @@
-python src/main.py \
-    dataset_dir=$dataset_dir \
-    dataset=s3dis \
-    model=minkowski \
-    tasks=["train","eval"] \
-    hydra.run.dir=outputs/pretrain/moco-hard \
-    gpus=[0] \
-    dataset.batch_size=6 \
-    dataset.accumulate_grad_batches=8 \
-    dataset.voxel_size=0.05 \
-    model.optimizer.type=SGD \
-    model.optimizer.lr=0.2 \
-    model.scheduler.type=PolyLR \
-    model.scheduler.poly_power=0.9 \
-    model.scheduler.max_iter=20000 \
-    model.scheduler.interval=step \
-    model.scheduler.frequency=10 \
-    max_time="01:00:00:00" \
-    max_epochs=2000 \
-    max_steps=4000 \
-    check_val_every_n_epoch=20 \
-    model.train.train_workers=10 \
-    model.net.loss=mixing \
-    pretrain_checkpoint=\"moco-hard.ckpt\"

### 2022-02-01 src/models/minkowski/trainer.py
BYOL tunning

@@ -222,16 +222,6 @@ class MinkowskiBOYLBackboneTrainer(BackboneTrainer):
         target_view_1 = torch.cat(target_view_1, 0)
         target_view_2 = torch.cat(target_view_2, 0)
 
-        # Apply stop gradient to target network outputs
-        target_view_1 = target_view_1.detach()
-        target_view_2 = target_view_2.detach()
-
-        # normalize to unit vectors
-        online_view_1 = nn.functional.normalize(online_view_1, dim=1, p=2)
-        online_view_2 = nn.functional.normalize(online_view_2, dim=1, p=2)
-        target_view_1 = nn.functional.normalize(target_view_1, dim=1, p=2)
-        target_view_2 = nn.functional.normalize(target_view_2, dim=1, p=2)
-
         # limit max number of query points
         if online_view_1.shape[0] > max_pos:
             inds = np.random.choice(online_view_1.shape[0], max_pos, replace=False)
@@ -240,11 +230,21 @@ class MinkowskiBOYLBackboneTrainer(BackboneTrainer):
             target_view_1 = target_view_1[inds]
             target_view_2 = target_view_2[inds]
 
+        # normalize to unit vectors
+        online_view_1 = nn.functional.normalize(online_view_1, dim=1, p=2)
+        online_view_2 = nn.functional.normalize(online_view_2, dim=1, p=2)
+        target_view_1 = nn.functional.normalize(target_view_1, dim=1, p=2)
+        target_view_2 = nn.functional.normalize(target_view_2, dim=1, p=2)
+
+        # Apply stop gradient to target network outputs
+        target_view_1 = target_view_1.detach()
+        target_view_2 = target_view_2.detach()
+
         # Compute regression loss (symmetrically)
         loss = self.criterion(online_view_1, target_view_2)
         loss += self.criterion(online_view_2, target_view_1)
 
-        return loss / 2.0
+        return loss * 100
 
 
 class MinkowskiMocoBackboneTrainer(BackboneTrainer):

### 2022-02-01 doc/contrastive-learning.tex
updates with results from moco

--- 2021-12-08 doc/contrastive-learning.tex
@@
 
+\section{Results}
+
+\subsection{Experimental Setup}
+
+hyper-parameters:
+
+\begin{itemize}
+    \item batch size: 6
+    \item accumulate gradient: 8
+    \item learning rate: 0.2
+    \item iterations: 4000
+\end{itemize}
+
+\subsection{Semantic Segmentation Negative Sampling}
+
+\begin{center}
+    \begin{table}[h!]
+        \resizebox{\textwidth}{!}{
+            \begin{tabular}{ c | c c c c c c c c c c c c c | c }
+                                                                & ceiling        & floor          & wall           & beam           & column         & window         & door           & table          & chair          & sofa           & bookcase       & board          & clutter        & mIOU           \\
+                \hline
+                Baseline \cite{choy20194d}                      & 0.862          & 0.947          & 0.811          & \textbf{0.002} & \textbf{0.356} & 0.551          & 0.590          & 0.745          & 0.751          & 0.452          & 0.719          & 0.606          & 0.514          & 0.608          \\
+                PointContrast \cite{xie2020pointcontrast}       & \textbf{0.905} & \textbf{0.962} & \textbf{0.818} & 0              & 0.345          & \textbf{0.581} & \textbf{0.633} & \textbf{0.750} & \textbf{0.861} & \textbf{0.648} & \textbf{0.729} & \textbf{0.799} & \textbf{0.566} & \textbf{0.661} \\
+                \hline
+                Entropy                                         & \textbf{0.912} & 0.961          & \textbf{0.828} & 0              & 0.307          & 0.513          & 0.626          & 0.746          & 0.853          & 0.634          & \textbf{0.742} & 0.761          & \textbf{0.576} & 0.651          \\
+                Delta                                           & \textbf{0.924} & 0.959          & 0.817          & 0              & \textbf{0.388} & \textbf{0.611} & 0.563          & \textbf{0.770} & \textbf{0.867} & \textbf{0.667} & \textbf{0.742} & 0.635          & \textbf{0.587} & 0.656          \\
+                Custer                                          & \textbf{0.924} & 0.957          & 0.814          & 0              & \textbf{0.373} & 0.562          & \textbf{0.688} & \textbf{0.776} & 0.844          & 0.524          & 0.729          & 0.717          & 0.587          & 0.653          \\
+                \hline
+                Debiased \cite{chuang_debiased_2020}            & 0.910          & 0.958          & 0.811          & \textbf{0.001} & 0.260          & 0.537          & 0.655          & 0.745          & 0.859          & 0.555          & 0.699          & 0.770          & 0.562          & 0.640          \\
+                Debiased++ \cite{robinson_contrastive_2021}     & 0.901          & 0.961          & 0.815          & \textbf{0.008} & 0.286          & 0.586          & 0.683          & 0.745          & 0.855          & 0.599          & 0.716          & 0.779          & 0.554          & 0.653          \\
+                \hline
+                Sampling - Hard                                 & 0.904          & 0.959          & 0.824          & 0              & 0.361          & 0.579          & 0.659          & 0.731          & 0.837          & 0.505          & 0.697          & 0.711          & 0.557          & 0.640          \\
+                Sampling - Easy                                 & 0.666          & 0.952          & 0.779          & 0              & 0.201          & 0              & 0              & 0.740          & 0              & 0              & 0.690          & 0.784          & 0.526          & 0.411          \\
+                Sampling - Mixing \cite{kalantidis_hard_nodate} & 0.895          & 0.962          & \textbf{0.820} & 0              & 0.277          & 0.576          & \textbf{0.706} & 0.747          & 0.849          & 0.567          & 0.724          & 0.769          & 0.562          & 0.650          \\
+                \hline
+                Moco \cite{he_momentum_2020}                    & 0.888          & 0.941          & \textbf{0.831} & \textbf{0.001} & \textbf{0.369} & \textbf{0.582} & 0.610          & \textbf{0.752} & \textbf{0.863} & 0.578          & \textbf{0.749} & 0.758          & 0.565          & 0.653          \\
+            \end{tabular}
+        }
+        \caption{S3DIS pretraining results on Semantic Segmentation}
+    \end{table}
+\end{center}
+
+\pagebreak
+
 \newpage

### 2022-02-22 src/models/factory.py
implements first version of image to point pretraining

@@ -7,6 +7,7 @@ from models.minkowski.trainer import (
     MinkowskiMocoBackboneTrainer,
     MinkowskiBOYLBackboneTrainer,
     CMEBackboneTrainer,
+    CMEBackboneTrainerFull,
     ImageTrainer,
 )
 from models.minkowski.dataset import (
@@ -54,7 +55,7 @@ class ModelFactory:
         elif self.model_name == byol_name:
             return MinkowskiBOYLBackboneTrainer
         elif self.model_name == cme_name:
-            return CMEBackboneTrainer
+            return CMEBackboneTrainerFull
         elif self.model_name == image_name:
             return ImageTrainer
         else:

### 2022-02-22 src/main.py
adds pretrained 2D checkpoint option

@@ -83,6 +83,13 @@ class Trainer:
                 f"Resuming pretraining checkpoint: {Path(self.pretrain_checkpoint).name}"
             )
 
+        self.checkpoint_2d_path = None
+        if cfg.pretrain_checkpoint_2d:
+            self.checkpoint_2d_path = str(
+                Path.cwd() / "pretrain_checkpoints_2d" / cfg.pretrain_checkpoint_2d
+            )
+            log.info(f"Resuming 2D checkpoint: {Path(self.checkpoint_2d_path).name}")
+
         self.supervised_pretrain_checkpoint = None
         if cfg.supervised_pretrain_checkpoint:
             self.supervised_pretrain_checkpoint = str(
@@ -147,7 +154,7 @@ class Trainer:
         return pl.Trainer(
             logger=tb_logger,
             gpus=self.cfg.gpus,
-            accelerator=self.cfg.accelerator,
+            strategy=self.cfg.accelerator,
             max_epochs=self.cfg.max_epochs,
             check_val_every_n_epoch=int(self.cfg.check_val_every_n_epoch),
             callbacks=[get_checkpoint_callback(), lr_monitor],
@@ -168,7 +175,7 @@ class Trainer:
         return pl.Trainer(
             logger=tb_logger,
             gpus=self.cfg.gpus,
-            accelerator=self.cfg.accelerator,
+            strategy=self.cfg.accelerator,
             resume_from_checkpoint=self.pretrain_checkpoint,
             max_steps=self.cfg.dataset.pretrain.max_steps,
             check_val_every_n_epoch=self.cfg.check_val_every_n_epoch,
@@ -211,6 +218,11 @@ class Trainer:
             backbonewraper = backbone_wrapper_type.load_from_checkpoint(
                 cfg=self.cfg,
                 checkpoint_path=self.pretrain_checkpoint,
+            )
+        elif self.checkpoint_2d_path:
+            log.info("Loading pretrained 2D network")
+            backbonewraper = backbone_wrapper_type(
+                self.cfg, checkpoint_2d_path=self.checkpoint_2d_path
             )
         else:
             backbonewraper = backbone_wrapper_type(self.cfg)

### 2022-02-22 src/models/minkowski/trainer.py

@@ -482,10 +482,7 @@ class CMEBackboneTrainer(BackboneTrainer):
 
 # Cross modal expert trainer
 class CMEBackboneTrainerFull(BackboneTrainer):
-    def __init__(
-        self,
-        cfg: DictConfig,
-    ):
+    def __init__(self, cfg: DictConfig, checkpoint_2d_path=None):
         super(CMEBackboneTrainerFull, self).__init__(cfg)
 
         # config
@@ -496,9 +493,12 @@ class CMEBackboneTrainerFull(BackboneTrainer):
         self.model = MinkovskiSemantic(cfg)
 
         # 2D feature extraction
-        image_trainer = ImageTrainer.load_from_checkpoint(
-            cfg=cfg, checkpoint_path="pretrain_checkpoints_2d/last.ckpt"
-        )
+        if checkpoint_2d_path:
+            image_trainer = ImageTrainer.load_from_checkpoint(
+                cfg=cfg, checkpoint_path=checkpoint_2d_path
+            )
+        else:
+            image_trainer = ImageTrainer(cfg)
         self.image_feature_extractor = image_trainer.model
 
     def training_step(self, batch: MinkowskiPretrainInput, batch_idx: int):

### 2022-02-24 doc/contrastive-learning.tex
adds BYOL results
updates results table
adds CME encoder only results
update cme results
adds CME full simsiam experiment

@@
 
-\subsection{Semantic Segmentation Negative Sampling}
+\subsection{Semantic Segmentation Pretraining}
 
@@
                 Moco \cite{he_momentum_2020}                    & 0.888          & 0.941          & \textbf{0.831} & \textbf{0.001} & \textbf{0.369} & \textbf{0.582} & 0.610          & \textbf{0.752} & \textbf{0.863} & 0.578          & \textbf{0.749} & 0.758          & 0.565          & 0.653          \\
+                BYOL \cite{grill_bootstrap_nodate}              & \textbf{0.907} & 0.958          & \textbf{8.25}  & 0              & 0.291          & 0.523          & \textbf{0.644} & 0.719          & 0.809          & 0.464          & \textbf{0.740} & 0.733          & 0.525          & 0.626          \\
+                \hline
+                CME - Encoder only                              & \textbf{0.911} & \textbf{0.963} & \textbf{0.818} & \textbf{0.002} & 0.278          & 0.541          & 0.521          & 0.733          & \textbf{0.883} & 0.524          & 0.705          & 0.756          & 0.559          & 0.630          \\
+                CME - Full, SimSiam                             & 0.882          & 0.958          & 0.812          & 0              & 0.355          & 0.525          & 0.474          & \textbf{0.759} & \textbf{0.869} & 0.458          & 0.723          & 0.767          & 0.533          & 0.624          \\
             \end{tabular}
@@
 \end{center}
+
+\subsection{Experiments}
+
+\subsubsection{24.02.2022 - Image to 3D full pretraining}
+
+\textbf{Details:} Ran 2D pre-training with SimSiam loss (which went to zero very quickly), then ran the same loss but between 2D to 3D using bi-linear interpolation to get the target feature vectors.
+Then finally ran full training on that.
+
+\textbf{Results:} The error for 2D and 2D-3D went down way too quickly and became super small.
+However, on the final output of the pre-training it seems to have helped somewhat (0.608 to 0.624) but could certainly be better.
+Seeing as the embedding only got to 0.63 I think there's more performance to be had and it also validates the approach somewhat.
 

### 2022-02-25 src/models/minkowski/dataset.py
Changes loss to contrastive for image and image to point

@@ -19,16 +19,18 @@ from scipy.spatial import KDTree
 from models.minkowski.types import (
     MinkowskiInput,
     MinkowskiPretrainInput,
+    MinkowskiPretrainInputNew,
     ImagePretrainInput,
 )
 from dataloaders.datasets import PretrainDataset, SegmentationDataset
 
 
-class ImagePretrainDataset(Dataset):
+class ImagePretrainDataset(PretrainDataset):
     def __init__(self, scenes, cfg):
-        super(ImagePretrainDataset, self).__init__()
+        super(ImagePretrainDataset, self).__init__(scenes, cfg)
 
         self.scenes = scenes
+
         image_size = 224
         crop = 0.2
         self.image_augmentations = transform_coord.Compose(
@@ -51,7 +53,6 @@ class ImagePretrainDataset(Dataset):
         coords1 = torch.vstack([datapoint["coords1"] for datapoint in batch])
         coords2 = torch.vstack([datapoint["coords2"] for datapoint in batch])
 
-        # correspondences = torch.cat([datapoint["correspondences"] for datapoint in batch], 0)
         correspondences = [
             datapoint["correspondences"] for datapoint in batch if datapoint
         ]
@@ -77,15 +78,23 @@ class ImagePretrainDataset(Dataset):
         with scene.open("rb") as scene_pickle:
             scene = pickle.load(scene_pickle)
 
+        if scene.points.shape[0] == 0:
+            new_ind = random.randint(0, len(self.scenes) - 1)
+            return self[new_ind]
+
         # Normalize image
         image = PIL.Image.fromarray(scene.color_image)
 
-        # Do two sets of augmentations
-        # self.image_augmentations(image).unsqueeze(dim=0).to(torch.float32)
+        # Do two sets of augmentations for 2D training
         image1, coords1 = self.image_augmentations(image)
         image2, coords2 = self.image_augmentations(image)
 
-        # visualize_image(image1)
+        # # Visualize image
+        # image1 = image1.transpose(2, 0) * 255.0
+        # image1 = image1.transpose(1, 0)
+        # image1 = image1.to(torch.uint8)
+        # vis = PIL.Image.fromarray(image1.detach().cpu().numpy())
+        # vis.show()
 
         # Generate a mapping
         coords1 = coords1.view(2, -1).T.detach().numpy()
@@ -204,89 +213,67 @@ class MinkowskiPretrainDataset(PretrainDataset):
     def collate(self, batch):
 
[…]

### 2022-02-25 src/models/minkowski/types.py

@@ -33,11 +33,60 @@ class MinkowskiInput:
 class ImagePretrainInput:
     """Pretrain input type of Minkowki Networks."""
 
+    # 2D Input
     images1: torch.Tensor
     images2: torch.Tensor
     coords1: torch.Tensor
     coords2: torch.Tensor
     correspondences: list
+
+    batch_size: int
+
+    # # 3D input
+    # points1: torch.Tensor
+    # features1: torch.Tensor
+    # points2: torch.Tensor
+    # features2: torch.Tensor
+
+    # # 2D to 3D
+    # point_to_pixel_maps1: torch.Tensor
+    # point_to_pixel_maps2: torch.Tensor
+    # image_coordinates: list
+
+    def to(self, device):
+        """Cast all tensor-type attributes to device"""
+        self.device = device
+        for fieldname, data in self.__dict__.items():
+            if type(data) == torch.Tensor:
+                setattr(self, fieldname, data.to(device))
+
+        return self
+
+    def __len__(self):
+        return self.batch_size
+
+
+@dataclass
+class MinkowskiPretrainInputNew:
+    """Pretrain input type of Minkowki Networks."""
+
+    # 3D input
+    points1: torch.Tensor
+    features1: torch.Tensor
+    points2: torch.Tensor
+    features2: torch.Tensor
+
+    # 2D
+    images: torch.Tensor
+    image_coordinates: list
+
+    # 3D to 3D
+    point_to_point_map: list
+
+    # 2D to 3D
+    point_to_pixel_maps1: torch.Tensor
+    point_to_pixel_maps2: torch.Tensor
+
     batch_size: int
 
     def to(self, device):

### 2022-02-25 doc/contrastive-learning.tex

@@
 
-\subsection{Experiments}
+\subsection{Experiment log}
 
@@
 
-\pagebreak
+\subsubsection{25.02.2022 - Image Contrastive learning}
+
+\textbf{Details:} Instead of using a BYOL loss with only positive samples, I went back to the old contrastive loss and was able to see a much more normal loss curve.
+Visualizing the results shows that it has learned key features in the images.
+
+\textbf{Results:} Here are some sample visualizations:
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/experiments/25.02.2022-image-pretrain-vis1.png}
+        % \caption{Visualization of pre-trained image features}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.43]{images/experiments/25.02.2022-image-pretrain-vis2.png}
+        % \caption{Visualization of pre-trained image features}
+    \end{center}
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+        \includegraphics[scale=0.35]{images/experiments/25.02.2022-image-pretrain-vis3.png}
+        % \caption{Visualization of pre-trained image features}
+    \end{center}
+\end{figure}
+
+\newpage
 

### 2022-03-01 doc/contrastive-learning.tex
adds results for contrastive 2D-3D experiment results

@@
                 CME - Full, SimSiam                             & 0.882          & 0.958          & 0.812          & 0              & 0.355          & 0.525          & 0.474          & \textbf{0.759} & \textbf{0.869} & 0.458          & 0.723          & 0.767          & 0.533          & 0.624          \\
+                CME - Full, Contrast                            & 0.899          & 0.960          & 0.820          & 0              & 0.369          & 0.527          & 0.584          & 0.757          & 0.849          & 0.551          & 0.713          & 0.697          & 0.529          & 0.635          \\
             \end{tabular}
@@
 
+\subsubsection{01.03.2022 - Image to 3D full pretraining}
+
+\textbf{Details:} Using the 2D Features learned from the previous experiment with a 2D pre-training objective, the goal was to transfer these 2D features over to the 3D points.
+This formed one side of the loss function, while the other side came from a 3D contrastive loss which compared points in the two transformed scans.
+
+\textbf{Results:} The results seem to show an improvement over baseline, and slightly over previous attempts. So far the best effort from 2D transfer learning. But still much bellow PointGroup.
+
 \newpage

### 2022-03-17 src/main.py
visualization updates to main

@@ -110,6 +110,9 @@ class Trainer:
             backbone_wrapper_type = self.model_factory.get_backbone_wrapper_type()
 
             # Load backbone parameters only
+            # self.pretrain_checkpoint = self.pretrain_checkpoint.replace(
+            #     "pretrain_checkpoints", "pretrain_checkpoints_2d"
+            # )
             checkpoint = torch.load(self.pretrain_checkpoint)
             head_params = [
                 key
@@ -304,7 +307,7 @@ class Trainer:
         dataset_type = self.model_factory.get_backbone_dataset_type()
         dataset = dataset_type(self.data_interface.pretrain_val_data, self.cfg)
         collate_fn = dataset.collate
-        scene = collate_fn([dataset[2175]])
+        scene = collate_fn([dataset[9759]])
 
         # Generate feature embeddings
         z1 = backbonewraper.model(scene.images1).detach().cpu().numpy()
@@ -318,6 +321,8 @@ class Trainer:
         plt.imshow(embeddings, cmap="autumn", interpolation="nearest")
         plt.title("2-D Heat Map")
         plt.show()
+
+        waithere = 1
 
     def pretrain_vis_3d(self):
         """Visualize the pretrained feature embeddings."""
@@ -338,7 +343,7 @@ class Trainer:
         dataset_type = self.model_factory.get_backbone_dataset_type()
         dataset = dataset_type(self.data_interface.pretrain_val_data, self.cfg)
         collate_fn = dataset.collate
-        scene = collate_fn([dataset[2175]])
+        scene = collate_fn([dataset[9759]])
 
         # Generate feature embeddings
         input_3d_1 = ME.SparseTensor(scene.features1, scene.points1)
@@ -346,27 +351,17 @@ class Trainer:
 
         pcd = o3d.geometry.PointCloud()
         pcd.points = o3d.utility.Vector3dVector(
-            scene.points1[:, 0:3].detach().cpu().numpy()
+            scene.points1[:, 1:4].detach().cpu().numpy()
         )
         pcd.colors = o3d.utility.Vector3dVector(scene.features1.detach().cpu().numpy())
 
-        # z1 = backbonewraper.model(scene.images1).detach().cpu().numpy()
-
-        # TSNE
-        # test = z1.reshape(16, -1).T
-        # embeddings = embed_tsne(output_3d_1.F.detach().cpu().numpy())
-        # # embeddings = embeddings.reshape(z1.shape[2], z1.shape[3])
-
[…]

### 2022-03-17 src/models/factory.py
updates transoforms to be like PointContrast

@@ -15,6 +15,7 @@ from models.minkowski.dataset import (
     MinkowskiDataset,
     MinkowskiPretrainDataset,
     MinkowskiFrameDataset,
+    MinkowskiS3DISDataset,
 )
 from util.types import DataInterface
 
@@ -29,6 +30,7 @@ supported_models = [minkowski_name, moco_name, byol_name, cme_name, image_name]
 class ModelFactory:
     def __init__(self, cfg: DictConfig, data_interface: DataInterface, backbone=None):
         self.model_name = cfg.model.name
+        self.dataset_name = cfg.dataset.name
         self.cfg = cfg
         self.data_interface = data_interface
         self.backbone = backbone
@@ -73,13 +75,10 @@ class ModelFactory:
             raise RuntimeError(self.error_msg)
 
     def get_dataset_type(self):
-        return MinkowskiDataset
-        # if cme_name in self.model_name:
-        #     return MinkowskiFrameDataset
-        # if minkowski_name in self.model_name:
-        #     return MinkowskiDataset
-        # else:
-        #     raise RuntimeError(self.error_msg)
+        if self.dataset_name == "S3DIS":
+            return MinkowskiS3DISDataset
+        else:
+            return MinkowskiDataset
 
     def get_backbone_dataset_type(self):
         if image_name == self.model_name:

### 2022-03-17 src/models/minkowski/dataset.py

@@ -507,9 +507,9 @@ class MinkowskiDataset(SegmentationDataset):
     def __init__(self, scenes, cfg, is_test=False):
         super(MinkowskiDataset, self).__init__(scenes, cfg, is_test=is_test)
 
-        color_jitter_std = 0.05
-        color_trans_ratio = 0.1
-        scale_range = (0.8, 1.2)
+        color_jitter_std = 0.005
+        color_trans_ratio = 0.05
+        scale_range = (0.9, 1.1)
         elastic_distortion_params = ((0.2, 0.4), (0.8, 1.6))
         self.augmentations = transforms.Compose(
             [
@@ -521,12 +521,6 @@ class MinkowskiDataset(SegmentationDataset):
                 transforms.RandomScale(scale_range),
                 transforms.RandomRotate(),
                 transforms.ElasticDistortion(elastic_distortion_params),
-            ]
-        )
-
-        self.test_augmentations = transforms.Compose(
-            [
-                transforms.Crop(self.max_pointcloud_size, self.ignore_label),
             ]
         )
 
@@ -556,9 +550,7 @@ class MinkowskiDataset(SegmentationDataset):
         features = torch.from_numpy(scene.features)
         labels = np.array([scene.semantic_labels, scene.instance_labels]).T
 
-        if self.is_test:
-            xyz, features, labels = self.test_augmentations(xyz, features, labels)
-        else:
+        if not self.is_test:
             xyz, features, labels = self.augmentations(xyz, features, labels)
 
         coords, feats, labels = ME.utils.sparse_quantize(
@@ -574,4 +566,41 @@ class MinkowskiDataset(SegmentationDataset):
             labels=labels,
             batch_size=1,
             test_filename=scene.name,
+        )
+
+
+class MinkowskiS3DISDataset(MinkowskiDataset):
+    def __init__(self, scenes, cfg, is_test=False):
+        super(MinkowskiS3DISDataset, self).__init__(scenes, cfg, is_test=is_test)
+
+        color_jitter_std = 0.005
+        color_trans_ratio = 0.05
+        scale_range = (0.9, 1.1)
+        elastic_distortion_params = None
+
+        CLIP_BOUND = 4  # [-N, N]
+
+        # Augmentation arguments
+        ROTATION_AUGMENTATION_BOUND = (
+            (-np.pi / 32, np.pi / 32),
+            (-np.pi / 32, np.pi / 32),
+            (-np.pi, np.pi),
+        )
[…]

### 2022-03-17 src/util/utils.py
updates TSNE parameters

@@ -268,7 +268,7 @@ def embed_tsne(data):
     """
     N x D np.array data
     """
-    tsne = TSNE(n_components=1, verbose=1, perplexity=40, n_iter=300, random_state=0)
+    tsne = TSNE(n_components=1, verbose=1, perplexity=40, n_iter=350, random_state=0)
     tsne_results = tsne.fit_transform(data)
     tsne_results = np.squeeze(tsne_results)
     tsne_min = np.min(tsne_results)

### 2022-03-18 doc/contrastive-learning.tex
updates CME results

@@
                 CME - Full, SimSiam                             & 0.882          & 0.958          & 0.812          & 0              & 0.355          & 0.525          & 0.474          & \textbf{0.759} & \textbf{0.869} & 0.458          & 0.723          & 0.767          & 0.533          & 0.624          \\
-                CME - Full, Contrast                            & 0.899          & 0.960          & 0.820          & 0              & 0.369          & 0.527          & 0.584          & 0.757          & 0.849          & 0.551          & 0.713          & 0.697          & 0.529          & 0.635          \\
+                CME - Full, Contrast                            & 0.892          & 0.960          & 0.828          & 0              & 0.307          & 0.558          & 0.554          & 0.767          & 0.847          & 0.505          & 0.725          & 0.793          & 0.542          & 0.637          \\
             \end{tabular}
@@
 
+\subsubsection{18.03.2022 - Image to 3D full pretraining}
+
+\textbf{Details:} Using the 3D features learned previously, the goal was to transfer these features to the 3D points. Here I used a 3D to 2D contrastive loss with no additional 3D-3D contrastive loss.
+
+\textbf{Results:} It seems to improve performance slightly.
+The biggest difference is that it starts at a significantly lower loss (2.5 vs 2.8) and thus goes down slightly quicker than without pre-training.
+The visualization also shows a big difference in terms of making sense of the points.
+
 \newpage

### 2022-03-18 src/main.py
adds back checkpoint loading

@@ -163,6 +163,7 @@ class Trainer:
             gpus=self.cfg.gpus,
             strategy=self.cfg.accelerator,
             max_epochs=self.cfg.max_epochs,
+            resume_from_checkpoint=self.checkpoint_path,
             check_val_every_n_epoch=int(self.cfg.check_val_every_n_epoch),
             callbacks=[get_checkpoint_callback(), lr_monitor],
             limit_train_batches=self.cfg.limit_train_batches,

### 2022-03-18 scripts/experiments/s3dis/minkowski-pretrained-cme-full.sh
updates scripts

@@ -2,6 +2,7 @@ python src/main.py \
     dataset_dir=$dataset_dir \
     dataset=s3dis \
     model=minkowski \
+    model.name=minkowski_cme \
     tasks=["train","eval"] \
     hydra.run.dir=$output_dir/pretrain/images \
     gpus=[0] \

### 2022-03-18 src/main.py

@@ -161,7 +161,7 @@ class Trainer:
         return pl.Trainer(
             logger=tb_logger,
             gpus=self.cfg.gpus,
-            strategy=self.cfg.accelerator,
+            accelerator=self.cfg.accelerator,
             max_epochs=self.cfg.max_epochs,
             resume_from_checkpoint=self.checkpoint_path,
             check_val_every_n_epoch=int(self.cfg.check_val_every_n_epoch),
@@ -183,7 +183,7 @@ class Trainer:
         return pl.Trainer(
             logger=tb_logger,
             gpus=self.cfg.gpus,
-            strategy=self.cfg.accelerator,
+            accelerator=self.cfg.accelerator,
             resume_from_checkpoint=self.pretrain_checkpoint,
             max_steps=self.cfg.dataset.pretrain.max_steps,
             check_val_every_n_epoch=self.cfg.check_val_every_n_epoch,

### 2022-03-18 scripts/experiments/s3dis/minkowski.sh
updates scripts

@@ -8,14 +8,8 @@ python src/main.py \
     dataset.batch_size=6 \
     dataset.accumulate_grad_batches=8 \
     dataset.voxel_size=0.05 \
-    model.optimizer.type=SGD \
-    model.optimizer.lr=0.2 \
-    model.scheduler.type=PolyLR \
-    model.scheduler.poly_power=0.9 \
-    model.scheduler.max_iter=20000 \
-    model.scheduler.interval=step \
-    model.scheduler.frequency=100 \
     max_time="02:00:00:00" \
     max_epochs=2000 \
-    check_val_every_n_epoch=10 \
-    model.train.train_workers=10 
+    check_val_every_n_epoch=20 \
+    model.train.train_workers=10 \
+    checkpoint=last.ckpt

### 2022-03-18 src/dataloaders/transforms.py
updates training to follow PointContrast

@@ -37,9 +37,8 @@ class ChromaticTranslation(object):
 
     def __call__(self, coords, feats, labels):
         if random.random() < 0.95:
-            tr = (np.random.rand(1, 3) - 0.5) * 2 * self.trans_range_ratio
-            # tr = (np.random.rand(1, 3) - 0.5) * 255 * 2 * self.trans_range_ratio
-            feats[:, :3] = np.clip(feats[:, :3] + tr, -0.5, 0.5)
+            tr = (np.random.rand(1, 3) - 0.5) * 255 * 2 * self.trans_range_ratio
+            feats[:, :3] = np.clip(feats[:, :3] + tr, 0, 255)
         return coords, feats, labels
 
 
@@ -54,11 +53,11 @@ class ChromaticAutoContrast(object):
             # std = np.std(feats, 0, keepdims=True)
             # lo = mean - std
             # hi = mean + std
-            lo = feats[:, :3].min(0, keepdims=True)
-            hi = feats[:, :3].max(0, keepdims=True)
+            lo = feats[:, :3].min(0, keepdims=True).values
+            hi = feats[:, :3].max(0, keepdims=True).values
             assert hi.max() > 1, f"invalid color value. Color is supposed to be [0-255]"
 
-            scale = 255 / (hi - lo)
+            scale = 255.0 / (hi - lo)
 
             contrast_feats = (feats[:, :3] - lo) * scale
 
@@ -76,8 +75,8 @@ class ChromaticJitter(object):
     def __call__(self, coords, feats, labels):
         if random.random() < 0.95:
             noise = np.random.randn(feats.shape[0], 3)
-            noise *= self.std
-            feats[:, :3] = np.clip(feats[:, :3] + noise, -1, 1)
+            noise *= self.std * 255
+            feats[:, :3] = np.clip(feats[:, :3] + noise, 0, 255)
         return coords, feats, labels
 
 
@@ -95,8 +94,8 @@ class RandomDropout(object):
     def __call__(self, coords, feats, labels):
         if random.random() < self.dropout_ratio:
             N = len(coords)
-            if N < 10:
-                return coords, feats, labels
+            # if N < 10:
+            #     return coords, feats, labels
             inds = np.random.choice(N, int(N * (1 - self.dropout_ratio)), replace=False)
             return coords[inds], feats[inds], labels[inds]
         return coords, feats, labels

### 2022-03-18 src/models/minkowski/dataset.py

@@ -550,8 +550,14 @@ class MinkowskiDataset(SegmentationDataset):
         features = torch.from_numpy(scene.features)
         labels = np.array([scene.semantic_labels, scene.instance_labels]).T
 
+        # Unnormalize features for a sec
+        features = (features + 1.0) * 127.5
+
         if not self.is_test:
             xyz, features, labels = self.augmentations(xyz, features, labels)
+
+        # Renormalize
+        features = features / 255.0 - 0.5
 
         coords, feats, labels = ME.utils.sparse_quantize(
             xyz,
@@ -559,6 +565,8 @@ class MinkowskiDataset(SegmentationDataset):
             labels=labels[:, 0],
             quantization_size=self.voxel_size,
         )
+
+        coords[:, :3] += (np.random.random(3) * 100).astype(np.int)
 
         return MinkowskiInput(
             points=coords,
@@ -595,6 +603,7 @@ class MinkowskiS3DISDataset(MinkowskiDataset):
                 ),
                 transforms.RandomDropout(0.2),
                 transforms.RandomHorizontalFlip("z", False),
+                transforms.ChromaticAutoContrast(),
                 transforms.ChromaticTranslation(color_trans_ratio),
                 transforms.ChromaticJitter(color_jitter_std),
                 transforms.RandomScale(scale_range),

### 2022-03-18 src/models/minkowski/trainer.py

@@ -23,7 +23,7 @@ from models.minkowski.types import (
     MinkowskiPretrainInputNew,
     ImagePretrainInput,
 )
-from util.utils import NCESoftmaxLoss
+from util.utils import NCESoftmaxLoss, set_seed
 import matplotlib.pyplot as plt
 
 # Visualization
@@ -1143,6 +1143,9 @@ class MinkowskiTrainer(SegmentationTrainer):
     @property
     def return_instances(self):
         return False
+
+    def on_before_zero_grad(self, optimizer) -> None:
+        set_seed(self.trainer.global_step)
 
     def training_step(self, batch: MinkowskiInput, batch_idx: int):
         model_input = ME.SparseTensor(batch.features, batch.points)

### 2022-03-18 src/util/utils.py

@@ -2,6 +2,7 @@ import sys
 import shutil
 import copy
 import math
+import random
 import pprint
 from pathlib import Path
 import concurrent.futures
@@ -12,6 +13,16 @@ import torch
 import torch.nn as nn
 import open3d as o3d
 import numpy as np
+import pytorch_lightning as pl
+
+
+def set_seed(curr_iteration):
+    # Set random seeds for reproductability
+    seed = 123 + curr_iteration
+    torch.manual_seed(seed)
+    random.seed(0)
+    np.random.seed(0)
+    pl.seed_everything(seed, workers=True)
 
 
 class NCESoftmaxLoss(nn.Module):

### 2022-03-31 doc/paper.tex
adds initial paper setup

% CVPR 2022 Paper Template
% based on the CVPR template provided by Ming-Ming Cheng (https://github.com/MCG-NKU/CVPR_Template)
% modified and extended by Stefan Roth (stefan.roth@NOSPAMtu-darmstadt.de)

\documentclass[10pt,twocolumn,letterpaper]{article}

%%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
\usepackage[review]{cvpr}      % To produce the REVIEW version
%\usepackage{cvpr}              % To produce the CAMERA-READY version
%\usepackage[pagenumbers]{cvpr} % To force page numbers, e.g. for an arXiv version

% Include other packages here, before hyperref.
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}

% It is strongly recommended to use hyperref, especially for the review version.
% hyperref with option pagebackref eases the reviewers' job.
% Please disable hyperref *only* if you encounter grave issues, e.g. with the
% file validation for the camera-ready version.
%
% If you comment hyperref and then uncomment it, you should delete
% ReviewTempalte.aux before re-running LaTeX.
% (Or just hit 'q' on the first LaTeX run, let it finish, and you
%  should be clear).
\usepackage[pagebackref,breaklinks,colorlinks]{hyperref}

% Support for easy cross-referencing
\usepackage[capitalize]{cleveref}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\crefname{table}{Tab.}{Tabs.}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{*****} % *** Enter the CVPR Paper ID here
\def\confName{CVPR}
\def\confYear{2022}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{\LaTeX\ Author Guidelines for \confName~Proceedings}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}
\maketitle

%%%%%%%%% ABSTRACT
\begin{abstract}
    The ABSTRACT is to be in fully justified italicized text, at the top of the left-hand column, below the author and affiliation information.
    Use the word ``Abstract'' as the title, in 12-point Times, boldface type, centered relative to the column, initially capitalized.
    The abstract is to be in 10-point, single-spaced type.
    Leave two blank lines after the Abstract, then begin the main text.
    Look at previous CVPR abstracts to get a feel for style and length.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:intro}

Please follow the steps outlined below when submitting your manuscript to the IEEE Computer Society Press.
This style guide now has several important modifications (for example, you are no longer warned against the use of sticky tape to attach your artwork to the paper), so all authors should read this new version.

%-------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

\subsection{Dual submission}

Please refer to the author guidelines on the \confName\ \confYear\ web page for a
discussion of the policy on dual submissions.

\subsection{Paper length}
Papers, excluding the references section, must be no longer than eight pages in length.
The references section will not be included in the page count, and there is no limit on the length of the references section.
For example, a paper of eight pages with two pages of references would have a total length of 10 pages.
    {\bf There will be no extra page charges for \confName\ \confYear.}

Overlength papers will simply not be reviewed.
This includes papers where the margins and formatting are deemed to have been significantly altered from those laid down by this style guide.
Note that this \LaTeX\ guide already sets figure captions and references in a smaller font.
The reason such papers will not be reviewed is that there is no provision for supervised revisions of manuscripts.
The reviewing process cannot determine the suitability of the paper for presentation in eight pages if it is reviewed in eleven.

%-------------------------------------------------------------------------
\subsection{The ruler}
The \LaTeX\ style defines a printed ruler which should be present in the version submitted for review.
The ruler is provided in order that reviewers may comment on particular lines in the paper without circumlocution.
If you are preparing a document using a non-\LaTeX\ document preparation system, please arrange for an equivalent ruler to appear on the final output pages.
The presence or absence of the ruler should not change the appearance of any other content on the page.
The camera-ready copy should not contain a ruler.
(\LaTeX\ users may use options of cvpr.sty to switch between different versions.)

Reviewers:
note that the ruler measurements do not align well with lines in the paper --- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
Just use fractional references (\eg, this line is $087.5$), although in most cases one would expect that the approximate location will be adequate.

\subsection{Paper ID}
Make sure that the Paper ID from the submission system is visible in the version submitted for review (replacing the ``*****'' you see in this document).
If you are using the \LaTeX\ template, \textbf{make sure to update paper ID in the appropriate place in the tex file}.

\subsection{Mathematics}

Please number all of your sections and displayed equations as in these examples:
\begin{equation}
    E = m\cdot c^2
    \label{eq:important}
\end{equation}
and
\begin{equation}
    v = a\cdot t.
    \label{eq:also-important}
\end{equation}
It is important for readers to be able to refer to any particular equation.
Just because you did not refer to it in the text does not mean some future reader might not need to refer to it.
It is cumbersome to have to use circumlocutions like ``the equation second from the top of page 3 column 1''.
(Note that the ruler will not be present in the final copy, so is not an alternative to equation numbers).
All authors will benefit from reading Mermin's description of how to write mathematics:
\url{http://www.pamitc.org/documents/mermin.pdf}.

\subsection{Blind review}

Many authors misunderstand the concept of anonymizing for blind review.
Blind review does not mean that one must remove citations to one's own work---in fact it is often impossible to review a paper unless the previous citations are known and available.

Blind review means that you do not use the words ``my'' or ``our'' when citing previous work.
That is all.
(But see below for tech reports.)

Saying ``this builds on the work of Lucy Smith [1]'' does not say that you are Lucy Smith;
it says that you are building on her work.
If you are Smith and Jones, do not say ``as we show in [7]'', say ``as Smith and Jones show in [7]'' and at the end of the paper, include reference 7 as you would any other cited work.

An example of a bad paper just asking to be rejected:
\begin{quote}
    \begin{center}
        An analysis of the frobnicatable foo filter.
    \end{center}

    In this paper we present a performance analysis of our previous paper [1], and show it to be inferior to all previously known methods.
    Why the previous paper was accepted without this analysis is beyond me.

        [1] Removed for blind review
\end{quote}

An example of an acceptable paper:
\begin{quote}
    \begin{center}
        An analysis of the frobnicatable foo filter.
    \end{center}

    In this paper we present a performance analysis of the  paper of Smith \etal [1], and show it to be inferior to all previously known methods.
    Why the previous paper was accepted without this analysis is beyond me.

        [1] Smith, L and Jones, C. ``The frobnicatable foo filter, a fundamental contribution to human knowledge''. Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time, which covers similar or overlapping material, you may need to refer to that submission in order to explain the differences, just as you would if you had previously published related work.
In such cases, include the anonymized parallel submission~\cite{Authors14} as supplemental material and cite it as
\begin{quote}
    [1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324, Supplied as supplemental material {\tt fg324.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be found elsewhere, and refer them to a technical report.
For conference submissions, the paper must stand on its own, and not {\em require} the reviewer to go to a tech report for further details.
Thus, you may say in the body of the paper ``further details may be found in~\cite{Authors14b}''.
Then submit the tech report as supplemental material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool that is widely known to be restricted to a single institution.
For example, let's say it's 1969, you have solved a key problem on the Apollo lander, and you believe that the CVPR70 audience would like to hear about your
solution.
The work is a development of your celebrated 1968 paper entitled ``Zero-g frobnication: How being the only people in the world with access to the Apollo lander source code makes us a wow at parties'', by Zeus \etal.

You can handle this paper like any other.
Do not write ``We show how to improve our previous work [Anonymous, 1968].
This time we tested the algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors.
Instead write the following:
\begin{quotation}
    \noindent
    We describe a system for zero-g frobnication.
    This system is new because it handles the following cases:
    A, B.  Previous systems [Zeus et al. 1968] did not  handle case B properly.
    Ours handles it by including a foo term in the bar integral.

    ...

    The proposed system was integrated with the Apollo lunar lander, and went all the way to the moon, don't you know.
    It displayed the following behaviours, which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention, reads better than the first version, and does not explicitly name you as the authors.
A reviewer might think it likely that the new paper was written by Zeus \etal, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been contracted to solve problem B.
\medskip

\noindent
FAQ\medskip\\
{\bf Q:} Are acknowledgements OK?\\
{\bf A:} No.  Leave them for the final copy.\medskip\\
{\bf Q:} How do I cite my results reported in open challenges?
{\bf A:} To conform with the double-blind review policy, you can report results of other challenge participants together with your results in your paper.
For your results, however, you should not identify yourself and should not mention your participation in the challenge.
Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\

\begin{figure}[t]
    \centering
    \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
    %\includegraphics[width=0.8\linewidth]{egfigure.eps}

    \caption{Example of caption.
        It is set in Roman so that mathematics (always set in Roman: $B \sin A = A \sin B$) may be included without an ugly clash.}
    \label{fig:onecol}
\end{figure}

\subsection{Miscellaneous}

\noindent
Compare the following:\\
\begin{tabular}{ll}
    \verb'$conf_a$' & $conf_a$          \\
    \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
\end{tabular}\\
See The \TeX book, p165.

The space after \eg, meaning ``for example'', should not be a sentence-ending space.
So \eg is correct, {\em e.g.} is not.
The provided \verb'\eg' macro takes care of this.

When citing a multi-author paper, you may save space by using ``et alia'', shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word).
If you use the \verb'\etal' macro provided, then you need not worry about double periods when used at the end of a sentence as in Alpher \etal.
However, use it only when there are three or more authors.
Thus, the following is correct:
``Frobnication has been trendy lately.
It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''

This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...'' because reference~\cite{Alpher03} has just two authors.

% Update the cvpr.cls to do the following automatically.
% For this citation style, keep multiple citations in numerical (not
% chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
% \cite{Alpher02,Alpher03,Authors14}.

\begin{figure*}
    \centering
    \begin{subfigure}{0.68\linewidth}
        \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
        \caption{An example of a subfigure.}
        \label{fig:short-a}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.28\linewidth}
        \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
        \caption{Another example of a subfigure.}
        \label{fig:short-b}
    \end{subfigure}
    \caption{Example of a short caption, which should be centered.}
    \label{fig:short}
\end{figure*}

%------------------------------------------------------------------------
\section{Formatting your paper}
\label{sec:formatting}

All text must be in a two-column format.
The total allowable size of the text area is $6\frac78$ inches (17.46 cm) wide by $8\frac78$ inches (22.54 cm) high.
Columns are to be $3\frac14$ inches (8.25 cm) wide, with a $\frac{5}{16}$ inch (0.8 cm) space between them.
The main title (on the first page) should begin 1 inch (2.54 cm) from the top edge of the page.
The second and following pages should begin 1 inch (2.54 cm) from the top edge.
On all pages, the bottom margin should be $1\frac{1}{8}$ inches (2.86 cm) from the bottom edge of the page for $8.5 \times 11$-inch paper;
for A4 paper, approximately $1\frac{5}{8}$ inches (4.13 cm) from the bottom edge of the
page.

%-------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must be kept
within a print area $6\frac{7}{8}$ inches (17.46 cm) wide by $8\frac{7}{8}$ inches (22.54 cm)
high.
%
Page numbers should be in the footer, centered and $\frac{3}{4}$ inches from the bottom of the page.
The review version should have page numbers, yet the final version submitted as camera ready should not show any page numbers.
The \LaTeX\ template takes care of this when used properly.

%-------------------------------------------------------------------------
\subsection{Type style and fonts}

Wherever Times is specified, Times Roman may also be used.
If neither is available on your word processor, please use the font closest in
appearance to Times to which you have access.

MAIN TITLE.
Center the title $1\frac{3}{8}$ inches (3.49 cm) from the top edge of the first page.
The title should be in Times 14-point, boldface type.
Capitalize the first letter of nouns, pronouns, verbs, adjectives, and adverbs;
do not capitalize articles, coordinate conjunctions, or prepositions (unless the title begins with such a word).
Leave two blank lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
and printed in Times 12-point, non-boldface type.
This information is to be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT.
Type main text in 10-point Times, single-spaced.
Do NOT use double-spacing.
All paragraphs should be indented 1 pica (approx.~$\frac{1}{6}$ inch or 0.422 cm).
Make sure your text is fully justified---that is, flush left and flush right.
Please do not place any additional blank lines between paragraphs.

Figure and table captions should be 9-point Roman type as in \cref{fig:onecol,fig:short}.
Short captions should be centred.

\noindent Callouts should be 9-point Helvetica, non-boldface type.
Initially capitalize only the first word of section titles and first-, second-, and third-order headings.

FIRST-ORDER HEADINGS.
(For example, {\large \bf 1. Introduction}) should be Times 12-point boldface, initially capitalized, flush left, with one blank line before, and one blank line after.

SECOND-ORDER HEADINGS.
(For example, { \bf 1.1. Database elements}) should be Times 11-point boldface, initially capitalized, flush left, with one blank line before, and one after.
If you require a third-order heading (we discourage it), use 10-point Times, boldface, initially capitalized, flush left, preceded by one blank line, followed by a period and your text on the same line.

%-------------------------------------------------------------------------
\subsection{Footnotes}

Please use footnotes\footnote{This is what a footnote looks like.
    It often distracts the reader from the main flow of the argument.} sparingly.
Indeed, try to avoid footnotes altogether and include necessary peripheral observations in the text (within parentheses, if you prefer, as in this sentence).
If you wish to use a footnote, place it at the bottom of the column on the page on which it is referenced.
Use Times 8-point type, single-spaced.

%-------------------------------------------------------------------------
\subsection{Cross-references}

For the benefit of author(s) and readers, please use the
    {\small\begin{verbatim}
  \cref{...}
\end{verbatim}}  command for cross-referencing to figures, tables, equations, or sections.
This will automatically insert the appropriate label alongside the cross-reference as in this example:
\begin{quotation}
    To see how our method outperforms previous work, please see \cref{fig:onecol} and \cref{tab:example}.
    It is also possible to refer to multiple targets as once, \eg~to \cref{fig:onecol,fig:short-a}.
    You may also return to \cref{sec:formatting} or look at \cref{eq:also-important}.
\end{quotation}
If you do not wish to abbreviate the label, for example at the beginning of the sentence, you can use the
    {\small\begin{verbatim}
  \Cref{...}
\end{verbatim}}
command. Here is an example:
\begin{quotation}
    \Cref{fig:onecol} is also quite important.
\end{quotation}

%-------------------------------------------------------------------------
\subsection{References}

List and number all bibliographical references in 9-point Times, single-spaced, at the end of your paper.
When referenced in the text, enclose the citation number in square brackets, for
example~\cite{Authors14}.
Where appropriate, include page numbers and the name(s) of editors of referenced books.
When you cite multiple papers at once, please make sure that you cite them in numerical order like this \cite{Alpher02,Alpher03,Alpher05,Authors14b,Authors14}.
If you use the template as advised, this will be taken care of automatically.

\begin{table}
    \centering
    \begin{tabular}{@{}lc@{}}
        \toprule
        Method & Frobnability           \\
        \midrule
        Theirs & Frumpy                 \\
        Yours  & Frobbly                \\
        Ours   & Makes one's heart Frob \\
        \bottomrule
    \end{tabular}
    \caption{Results.   Ours is better.}
    \label{tab:example}
\end{table}

%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.
In \LaTeX, avoid using the \texttt{center} environment for this purpose, as this adds potentially unwanted whitespace.
Instead use
    {\small\begin{verbatim}
  \centering
\end{verbatim}}
at the beginning of your figure.
Please ensure that any point you wish to make is resolvable in a printed copy of the paper.
Resize fonts in figures to match the font in the body text, and choose line widths that render effectively in print.
Readers (and reviewers), even of an electronic copy, may choose to print your paper in order to read it.
You cannot insist that they do otherwise, and therefore must not assume that they can zoom in to see tiny details on a graphic.

When placing figures in \LaTeX, it's almost always best to use \verb+\includegraphics+, and to specify the figure width as a multiple of the line width as in the example below
    {\small\begin{verbatim}
   \usepackage{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.pdf}
\end{verbatim}
    }

%-------------------------------------------------------------------------
\subsection{Color}

Please refer to the author guidelines on the \confName\ \confYear\ web page for a discussion of the use of color in your document.

If you use color in your plots, please keep in mind that a significant subset of reviewers and readers may have a color vision deficiency; red-green blindness is the most frequent kind.
Hence avoid relying only on color as the discriminative feature in plots (such as red \vs green lines), but add a second discriminative feature to ease disambiguation.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit your finished paper.
We MUST have this form before your paper can be published in the proceedings.

Please direct any questions to the production editor in charge of these proceedings at the IEEE Computer Society Press:
\url{https://www.computer.org/about/contact}.

%%%%%%%%% REFERENCES
{\small
\bibliographystyle{ieee_fullname}
\bibliography{egbib}
}

\end{document}

### 2022-03-31 scripts/preprocess/scannet.sh
adds 2D semantic labels to measurements

@@ -1,5 +1,5 @@
 export SCANNET_DIR=/media/starslab/datasets/scannet
-export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_2cm_frames_and_scans
+export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_2cm_frames_with_image_labels
 export HYDRA_FULL_ERROR=1
 
 preprocess() {

### 2022-03-31 src/dataloaders/transforms.py

@@ -97,10 +97,7 @@ class RandomDropout(object):
             # if N < 10:
             #     return coords, feats, labels
             inds = np.random.choice(N, int(N * (1 - self.dropout_ratio)), replace=False)
-            if labels:
-                return coords[inds], feats[inds], labels[inds]
-            else:
-                return coords[inds], feats[inds], labels
+            return coords[inds], feats[inds], labels[inds]
         return coords, feats, labels
 
 

### 2022-03-31 src/datasets/measurements_new.py

@@ -2,11 +2,13 @@
 
 import pickle
 from pathlib import Path
+from re import L
 import numpy as np
 import open3d as o3d
 import cv2
 
 from util.types import Scene, SceneWithLabels
+from util.utils import get_random_colour
 
 measurements_dir_name = "measurements"
 colour_image_extension = ".color.jpg"
@@ -95,6 +97,10 @@ class SceneMeasurement:
         )
         pose_info_name = measurements_dir / (f"frame-{frame_id}" + pose_extension)
 
+        frame_number = int(self.frame_id)
+        instance_image_name = directory / (f"instance-filt/{frame_number}.png")
+        label_image_name = directory / (f"label-filt/{frame_number}.png")
+
         # load calibrations
         self.colour_intrinsics = np.fromstring(
             info["m_calibrationColorIntrinsic"], sep=" "
@@ -109,9 +115,23 @@ class SceneMeasurement:
         depth_image = depth_image.astype(np.float32) / float(info["m_depthShift"])
         self.pose = np.loadtxt(str(pose_info_name), delimiter=" ").astype(np.float32)
 
+        # load annotations
+        self.label_image = cv2.imread(str(directory / label_image_name))
+        self.instance_image = cv2.imread(str(directory / instance_image_name))
+
         # Resize colour image to be same as depth image
         depth_resolution = (int(info["m_depthWidth"]), int(info["m_depthHeight"]))
         self.color_image = cv2.resize(self.color_image, depth_resolution)
+        self.instance_image = cv2.resize(
+            self.instance_image,
+            depth_resolution,
+            interpolation=cv2.INTER_NEAREST,
+        )
+        # self.label_image = cv2.resize(
+        #     self.label_image,
+        #     depth_resolution,
+        #     interpolation=cv2.INTER_NEAREST,
+        # )
 
         # Convert depth map into point cloud
         self.scan_points, self.point_to_pixel_map = get_point_cloud(
@@ -128,7 +148,7 @@ class SceneMeasurement:
         self.scan_point_colors = self.scan_point_colors / 127.5 - 1
 
         # Get points visible in camera frame
-        frame = self.get_projected_scene(info, scene)
+        frame, pixels = self.get_projected_scene(info, scene)
 
         # # Get point cloud for camera view
         # pcd = o3d.geometry.PointCloud()
@@ -147,6 +167,47 @@ class SceneMeasurement:
         self.point_colors = frame.features
[…]

### 2022-03-31 doc/paper.tex

@@
 %%%%%%%%% TITLE - PLEASE UPDATE
-\title{\LaTeX\ Author Guidelines for \confName~Proceedings}
+\title{Exploring Self-Supervised Learning for Indoor 3D scenes}
 
-\author{First Author\\
-Institution1\\
+\author{Andrej Janda\\
+University of Toronto Institute for Aerospace Studies\\
 Institution1 address\\
-{\tt\small firstauthor@i1.org}
+{\tt\small andrej.janda@robotics.utias.utoronto.ca}
 % For a paper whose authors are all at the same institution,
@@
 \and
-Second Author\\
-Institution2\\
+Jonathan Kelly\\
+University of Toronto Institute for Aerospace Studies\\
 First line of institution2 address\\
-{\tt\small secondauthor@i2.org}
+{\tt\small jonathan.kelly@robotics.utias.utoronto.ca}
 }
@@
 \begin{abstract}
-    The ABSTRACT is to be in fully justified italicized text, at the top of the left-hand column, below the author and affiliation information.
-    Use the word ``Abstract'' as the title, in 12-point Times, boldface type, centered relative to the column, initially capitalized.
-    The abstract is to be in 10-point, single-spaced type.
-    Leave two blank lines after the Abstract, then begin the main text.
-    Look at previous CVPR abstracts to get a feel for style and length.
+    We show that previous work \cite{xie2020pointcontrast} on self-supervised contrastive learning of 3D point clouds relies on extrapolating points in the scene and not from learning noise as was claimed in the original paper.
+We then investigate if using 2D pre-trained features transfered to 3D can do as well as fully supervised learning.
 \end{abstract}
@@
 
-Please follow the steps outlined below when submitting your manuscript to the IEEE Computer Society Press.
-This style guide now has several important modifications (for example, you are no longer warned against the use of sticky tape to attach your artwork to the paper), so all authors should read this new version.
+% Write topic sentences and go from there
+Segmentation of indoor 3D scenes has the potential to allow robots to navigate and interact with complicated industrial environments, such as warehouses and production facilities.
+Although robots already play a limited role in these facilities, understanding their surroundings would give them much greater autonomy.
+For instance, through segmentation, an accurate up-to-date digital twin of a physical area can be maintained for the purposes of inventory management, item retrieval, and item routing.
+Advances in segmentation of 3D point clouds \cite{choy20194d} has made this closer to reality.
 
-%-------------------------------------------------------------------------
-\subsection{Language}
+Alternatively, segmentation of 2D images has come a long way.
+But it does not give us information about the 3D world and so we are left with projecting these predictions into 3D.
+This was the approach taken by \cite{} but has recently been overtaken by methods operating purely in 3D.
 
-All manuscripts must be in English.
+In contrast to 2D images which are bountiful, easy to collect and annotate, 3D scenes are hard to collect, not very abundant and notoriously hard to annotate.
+Collecting 3D scenes requires a user to physically scan different environments with complicated sensors (\eg stereo cameras, Lidar) and estimate the mapping between measurements to generate a completed scene.
+Since not everyone has access to these sensors, they cannot be easily scrapped off of the web.
+Still, the hardest component is annotation which requires users to manipulate 3D scenes and select only those points that belong to a particular object.
+This motivates us to look for ways of avoiding the annotation process and leverage un-labeled 3D data instead.
 
-\subsection{Dual submission}
+It is common practice in image tasks to use a model that was pre-trained in a supervised way on a larger but different dataset.
+However, recent works have been exploring the use of self-supervised objectives without labels in 2D image classification and found that they can greatly improve downstream accuracy, even matching fully supervised training \cite{}.
 
-Please refer to the author guidelines on the \confName\ \confYear\ web page for a
-discussion of the policy on dual submissions.
+Recently, works such as \cite{xie2020pointcontrast} have shown that these methods also work with 3D data.
+Similar to 2D methods they use a contrastive loss objective where the goal is to learn distinguishing features between different datapoints and their respective augmentations.
 
-\subsection{Paper length}
-Papers, excluding the references section, must be no longer than eight pages in length.
-The references section will not be included in the page count, and there is no limit on the length of the references section.
-For example, a paper of eight pages with two pages of references would have a total length of 10 pages.
-    {\bf There will be no extra page charges for \confName\ \confYear.}
+These algorithms work exclusively with point clouds but throw away the corresponding images of the scene which provide a rich and dense source in extra information that can be used in downstream training.
+This paper investigates how this extra data modality can be leveraged.
+Additionally, it investigates the properties of an existing 3D pre-training algorithm.
 
-Overlength papers will simply not be reviewed.
-This includes papers where the margins and formatting are deemed to have been significantly altered from those laid down by this style guide.
-Note that this \LaTeX\ guide already sets figure captions and references in a smaller font.
-The reason such papers will not be reviewed is that there is no provision for supervised revisions of manuscripts.
-The reviewing process cannot determine the suitability of the paper for presentation in eight pages if it is reviewed in eleven.
+This paper has the following contributions.
 
-%-------------------------------------------------------------------------
-\subsection{The ruler}
-The \LaTeX\ style defines a printed ruler which should be present in the version submitted for review.
-The ruler is provided in order that reviewers may comment on particular lines in the paper without circumlocution.
-If you are preparing a document using a non-\LaTeX\ document preparation system, please arrange for an equivalent ruler to appear on the final output pages.
-The presence or absence of the ruler should not change the appearance of any other content on the page.
-The camera-ready copy should not contain a ruler.
-(\LaTeX\ users may use options of cvpr.sty to switch between different versions.)
+\begin{itemize}
+    \item Demonstrates that previous work in 3D \cite{xie2020pointcontrast} works by learning to extrapolate 3D scenes.
+When purely learning improved features on fully-overlapping regions the performance gain disapears.
+    \item Provide a 2D feature extractor and pre-training algorithm. Demonstrate that these features can improve 2D segmentation.
+    \item Transfer learned 2D features onto a 3D model
+\end{itemize}
 
-Reviewers:
-note that the ruler measurements do not align well with lines in the paper --- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
-Just use fractional references (\eg, this line is $087.5$), although in most cases one would expect that the approximate location will be adequate.
+\section{Related Work}
+\label{sec:relatedWork}
 
-\subsection{Paper ID}
-Make sure that the Paper ID from the submission system is visible in the version submitted for review (replacing the ``*****'' you see in this document).
-If you are using the \LaTeX\ template, \textbf{make sure to update paper ID in the appropriate place in the tex file}.
+\section{Method}
+\label{sec:method}
 
-\subsection{Mathematics}
+\section{Feature Extractors}
+\label{sec:method:featuresExtractors}
 
-Please number all of your sections and displayed equations as in these examples:
-\begin{equation}
-    E = m\cdot c^2
-    \label{eq:important}
-\end{equation}
-and
-\begin{equation}
-    v = a\cdot t.
-    \label{eq:also-important}
-\end{equation}
-It is important for readers to be able to refer to any particular equation.
-Just because you did not refer to it in the text does not mean some future reader might not need to refer to it.
-It is cumbersome to have to use circumlocutions like ``the equation second from the top of page 3 column 1''.
-(Note that the ruler will not be present in the final copy, so is not an alternative to equation numbers).
-All authors will benefit from reading Mermin's description of how to write mathematics:
-\url{http://www.pamitc.org/documents/mermin.pdf}.
+\section{Extracting 2D features}
+\label{sec:method:features2d}
 
-\subsection{Blind review}
+\section{Transfering 2D features}
+\label{sec:method:features2D3D}
 
-Many authors misunderstand the concept of anonymizing for blind review.
-Blind review does not mean that one must remove citations to one's own work---in fact it is often impossible to review a paper unless the previous citations are known and available.
+Talk about projection and Bi-linear interpolation
 
-Blind review means that you do not use the words ``my'' or ``our'' when citing previous work.
-That is all.
-(But see below for tech reports.)
+\section{Experiment Results}
+\label{sec:results}
 
-Saying ``this builds on the work of Lucy Smith [1]'' does not say that you are Lucy Smith;
-it says that you are building on her work.
-If you are Smith and Jones, do not say ``as we show in [7]'', say ``as Smith and Jones show in [7]'' and at the end of the paper, include reference 7 as you would any other cited work.
+\begin{table*}[t!]
+    \centering
+    \resizebox{\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c c c c c c c c | c }
+                                        & ceiling        & floor          & wall           & beam & column & window & door           & table          & chair          & sofa           & bookcase       & board & clutter        & mIOU           \\
+            \hline
+            Scratch                     & 91.53          & 96.79          & 81.25          & 0.0  & 29.81  & 56.92  & 64.77          & 74.43          & 86.35          & 55.24          & 68.38          & 79.7  & 55.65          & 64.68          \\
+            Supervised                  &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
+            \hline
+            PointContrast               &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
+            PointContrast - No Overhang &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
+            CME                         & \textbf{91.07} & \textbf{96.74} & \textbf{82.84} & 0    & 28.38  & 55.98  & \textbf{70.34} & \textbf{75.01} & \textbf{87.34} & \textbf{63.24} & \textbf{71.25} & 75.01 & \textbf{56.42} & \textbf{65.67} \\
+        \end{tabular}
+    }
+    \caption{S3DIS results using ContrastiveSceneContexts Code}
+    \label{table:s3disResults}
+\end{table*}
 
-An example of a bad paper just asking to be rejected:
-\begin{quote}
-    \begin{center}
-        An analysis of the frobnicatable foo filter.
-    \end{center}
+\subsection{Datasets}
+\label{sec:results:datasets}
 
-    In this paper we present a performance analysis of our previous paper [1], and show it to be inferior to all previously known methods.
-    Why the previous paper was accepted without this analysis is beyond me.
+\subsection{Implementation}
+\label{sec:results:implementation}
 
-        [1] Removed for blind review
-\end{quote}
+\subsection{Contrastive Loss in 3D}
+\label{sec:results:3d}
 
-An example of an acceptable paper:
-\begin{quote}
-    \begin{center}
-        An analysis of the frobnicatable foo filter.
-    \end{center}
+\subsection{Contrastive Loss in 2D}
+\label{sec:results:2d}
 
-    In this paper we present a performance analysis of the  paper of Smith \etal [1], and show it to be inferior to all previously known methods.
-    Why the previous paper was accepted without this analysis is beyond me.
+\subsection{Transfering 2D features to 3D}
+\label{sec:results:2d3d}
 
-        [1] Smith, L and Jones, C. ``The frobnicatable foo filter, a fundamental contribution to human knowledge''. Nature 381(12), 1-213.
-\end{quote}
-
-If you are making a submission to another conference at the same time, which covers similar or overlapping material, you may need to refer to that submission in order to explain the differences, just as you would if you had previously published related work.
-In such cases, include the anonymized parallel submission~\cite{Authors14} as supplemental material and cite it as
-\begin{quote}
-    [1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324, Supplied as supplemental material {\tt fg324.pdf}.
-\end{quote}
-
-Finally, you may feel you need to tell the reader that more details can be found elsewhere, and refer them to a technical report.
-For conference submissions, the paper must stand on its own, and not {\em require} the reviewer to go to a tech report for further details.
-Thus, you may say in the body of the paper ``further details may be found in~\cite{Authors14b}''.
-Then submit the tech report as supplemental material.
-Again, you may not assume the reviewers will read this material.
-
-Sometimes your paper is about a problem which you tested using a tool that is widely known to be restricted to a single institution.
-For example, let's say it's 1969, you have solved a key problem on the Apollo lander, and you believe that the CVPR70 audience would like to hear about your
-solution.
-The work is a development of your celebrated 1968 paper entitled ``Zero-g frobnication: How being the only people in the world with access to the Apollo lander source code makes us a wow at parties'', by Zeus \etal.
-
-You can handle this paper like any other.
-Do not write ``We show how to improve our previous work [Anonymous, 1968].
-This time we tested the algorithm on a lunar lander [name of lander removed for blind review]''.
-That would be silly, and would immediately identify the authors.
-Instead write the following:
-\begin{quotation}
-    \noindent
-    We describe a system for zero-g frobnication.
-    This system is new because it handles the following cases:
-    A, B.  Previous systems [Zeus et al. 1968] did not  handle case B properly.
-    Ours handles it by including a foo term in the bar integral.
-
-    ...
-
-    The proposed system was integrated with the Apollo lunar lander, and went all the way to the moon, don't you know.
-    It displayed the following behaviours, which show how well we solved cases A and B: ...
-\end{quotation}
-As you can see, the above text follows standard scientific convention, reads better than the first version, and does not explicitly name you as the authors.
-A reviewer might think it likely that the new paper was written by Zeus \etal, but cannot make any decision based on that guess.
-He or she would have to be sure that no other authors could have been contracted to solve problem B.
-\medskip
-
-\noindent
-FAQ\medskip\\
-{\bf Q:} Are acknowledgements OK?\\
-{\bf A:} No.  Leave them for the final copy.\medskip\\
-{\bf Q:} How do I cite my results reported in open challenges?
-{\bf A:} To conform with the double-blind review policy, you can report results of other challenge participants together with your results in your paper.
-For your results, however, you should not identify yourself and should not mention your participation in the challenge.
-Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\
-
-\begin{figure}[t]
-    \centering
-    \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
-    %\includegraphics[width=0.8\linewidth]{egfigure.eps}
-
-    \caption{Example of caption.
-        It is set in Roman so that mathematics (always set in Roman: $B \sin A = A \sin B$) may be included without an ugly clash.}
-    \label{fig:onecol}
-\end{figure}
-
-\subsection{Miscellaneous}
-
-\noindent
-Compare the following:\\
-\begin{tabular}{ll}
-    \verb'$conf_a$' & $conf_a$          \\
-    \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
-\end{tabular}\\
-See The \TeX book, p165.
-
-The space after \eg, meaning ``for example'', should not be a sentence-ending space.
-So \eg is correct, {\em e.g.} is not.
-The provided \verb'\eg' macro takes care of this.
-
-When citing a multi-author paper, you may save space by using ``et alia'', shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word).
-If you use the \verb'\etal' macro provided, then you need not worry about double periods when used at the end of a sentence as in Alpher \etal.
-However, use it only when there are three or more authors.
-Thus, the following is correct:
-``Frobnication has been trendy lately.
-It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
-Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''
-
-This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...'' because reference~\cite{Alpher03} has just two authors.
-
-% Update the cvpr.cls to do the following automatically.
-% For this citation style, keep multiple citations in numerical (not
-% chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
-% \cite{Alpher02,Alpher03,Authors14}.
-
-\begin{figure*}
-    \centering
-    \begin{subfigure}{0.68\linewidth}
-        \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
-        \caption{An example of a subfigure.}
-        \label{fig:short-a}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.28\linewidth}
-        \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
-        \caption{Another example of a subfigure.}
-        \label{fig:short-b}
-    \end{subfigure}
-    \caption{Example of a short caption, which should be centered.}
-    \label{fig:short}
-\end{figure*}
-
-%------------------------------------------------------------------------
-\section{Formatting your paper}
-\label{sec:formatting}
-
-All text must be in a two-column format.
-The total allowable size of the text area is $6\frac78$ inches (17.46 cm) wide by $8\frac78$ inches (22.54 cm) high.
-Columns are to be $3\frac14$ inches (8.25 cm) wide, with a $\frac{5}{16}$ inch (0.8 cm) space between them.
-The main title (on the first page) should begin 1 inch (2.54 cm) from the top edge of the page.
-The second and following pages should begin 1 inch (2.54 cm) from the top edge.
-On all pages, the bottom margin should be $1\frac{1}{8}$ inches (2.86 cm) from the bottom edge of the page for $8.5 \times 11$-inch paper;
-for A4 paper, approximately $1\frac{5}{8}$ inches (4.13 cm) from the bottom edge of the
-page.
-
-%-------------------------------------------------------------------------
-\subsection{Margins and page numbering}
-
-All printed material, including text, illustrations, and charts, must be kept
-within a print area $6\frac{7}{8}$ inches (17.46 cm) wide by $8\frac{7}{8}$ inches (22.54 cm)
-high.
-%
-Page numbers should be in the footer, centered and $\frac{3}{4}$ inches from the bottom of the page.
-The review version should have page numbers, yet the final version submitted as camera ready should not show any page numbers.
-The \LaTeX\ template takes care of this when used properly.
-
-%-------------------------------------------------------------------------
-\subsection{Type style and fonts}
-
-Wherever Times is specified, Times Roman may also be used.
-If neither is available on your word processor, please use the font closest in
-appearance to Times to which you have access.
-
-MAIN TITLE.
-Center the title $1\frac{3}{8}$ inches (3.49 cm) from the top edge of the first page.
-The title should be in Times 14-point, boldface type.
-Capitalize the first letter of nouns, pronouns, verbs, adjectives, and adverbs;
-do not capitalize articles, coordinate conjunctions, or prepositions (unless the title begins with such a word).
-Leave two blank lines after the title.
-
-AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
-and printed in Times 12-point, non-boldface type.
-This information is to be followed by two blank lines.
-
-The ABSTRACT and MAIN TEXT are to be in a two-column format.
-
-MAIN TEXT.
-Type main text in 10-point Times, single-spaced.
-Do NOT use double-spacing.
-All paragraphs should be indented 1 pica (approx.~$\frac{1}{6}$ inch or 0.422 cm).
-Make sure your text is fully justified---that is, flush left and flush right.
-Please do not place any additional blank lines between paragraphs.
-
-Figure and table captions should be 9-point Roman type as in \cref{fig:onecol,fig:short}.
-Short captions should be centred.
-
-\noindent Callouts should be 9-point Helvetica, non-boldface type.
-Initially capitalize only the first word of section titles and first-, second-, and third-order headings.
-
-FIRST-ORDER HEADINGS.
-(For example, {\large \bf 1. Introduction}) should be Times 12-point boldface, initially capitalized, flush left, with one blank line before, and one blank line after.
-
-SECOND-ORDER HEADINGS.
-(For example, { \bf 1.1. Database elements}) should be Times 11-point boldface, initially capitalized, flush left, with one blank line before, and one after.
-If you require a third-order heading (we discourage it), use 10-point Times, boldface, initially capitalized, flush left, preceded by one blank line, followed by a period and your text on the same line.
-
-%-------------------------------------------------------------------------
-\subsection{Footnotes}
-
-Please use footnotes\footnote{This is what a footnote looks like.
-    It often distracts the reader from the main flow of the argument.} sparingly.
-Indeed, try to avoid footnotes altogether and include necessary peripheral observations in the text (within parentheses, if you prefer, as in this sentence).
-If you wish to use a footnote, place it at the bottom of the column on the page on which it is referenced.
-Use Times 8-point type, single-spaced.
-
-%-------------------------------------------------------------------------
-\subsection{Cross-references}
-
-For the benefit of author(s) and readers, please use the
-    {\small\begin{verbatim}
-  \cref{...}
-\end{verbatim}}  command for cross-referencing to figures, tables, equations, or sections.
-This will automatically insert the appropriate label alongside the cross-reference as in this example:
-\begin{quotation}
-    To see how our method outperforms previous work, please see \cref{fig:onecol} and \cref{tab:example}.
-    It is also possible to refer to multiple targets as once, \eg~to \cref{fig:onecol,fig:short-a}.
-    You may also return to \cref{sec:formatting} or look at \cref{eq:also-important}.
-\end{quotation}
-If you do not wish to abbreviate the label, for example at the beginning of the sentence, you can use the
-    {\small\begin{verbatim}
-  \Cref{...}
-\end{verbatim}}
-command. Here is an example:
-\begin{quotation}
-    \Cref{fig:onecol} is also quite important.
-\end{quotation}
-
-%-------------------------------------------------------------------------
-\subsection{References}
-
-List and number all bibliographical references in 9-point Times, single-spaced, at the end of your paper.
-When referenced in the text, enclose the citation number in square brackets, for
-example~\cite{Authors14}.
-Where appropriate, include page numbers and the name(s) of editors of referenced books.
-When you cite multiple papers at once, please make sure that you cite them in numerical order like this \cite{Alpher02,Alpher03,Alpher05,Authors14b,Authors14}.
-If you use the template as advised, this will be taken care of automatically.
-
-\begin{table}
-    \centering
-    \begin{tabular}{@{}lc@{}}
-        \toprule
-        Method & Frobnability           \\
-        \midrule
-        Theirs & Frumpy                 \\
-        Yours  & Frobbly                \\
-        Ours   & Makes one's heart Frob \\
-        \bottomrule
-    \end{tabular}
-    \caption{Results.   Ours is better.}
-    \label{tab:example}
-\end{table}
-
-%-------------------------------------------------------------------------
-\subsection{Illustrations, graphs, and photographs}
-
-All graphics should be centered.
-In \LaTeX, avoid using the \texttt{center} environment for this purpose, as this adds potentially unwanted whitespace.
-Instead use
-    {\small\begin{verbatim}
-  \centering
-\end{verbatim}}
-at the beginning of your figure.
-Please ensure that any point you wish to make is resolvable in a printed copy of the paper.
-Resize fonts in figures to match the font in the body text, and choose line widths that render effectively in print.
-Readers (and reviewers), even of an electronic copy, may choose to print your paper in order to read it.
-You cannot insist that they do otherwise, and therefore must not assume that they can zoom in to see tiny details on a graphic.
-
-When placing figures in \LaTeX, it's almost always best to use \verb+\includegraphics+, and to specify the figure width as a multiple of the line width as in the example below
-    {\small\begin{verbatim}
-   \usepackage{graphicx} ...
-   \includegraphics[width=0.8\linewidth]
-                   {myfile.pdf}
-\end{verbatim}
-    }
-
-%-------------------------------------------------------------------------
-\subsection{Color}
-
-Please refer to the author guidelines on the \confName\ \confYear\ web page for a discussion of the use of color in your document.
-
-If you use color in your plots, please keep in mind that a significant subset of reviewers and readers may have a color vision deficiency; red-green blindness is the most frequent kind.
-Hence avoid relying only on color as the discriminative feature in plots (such as red \vs green lines), but add a second discriminative feature to ease disambiguation.
-
-%------------------------------------------------------------------------
-\section{Final copy}
-
-You must include your signed IEEE copyright release form when you submit your finished paper.
-We MUST have this form before your paper can be published in the proceedings.
-
-Please direct any questions to the production editor in charge of these proceedings at the IEEE Computer Society Press:
-\url{https://www.computer.org/about/contact}.
+\section{Conclusion}
+\label{sec:conclusion}
 
@@
     \bibliographystyle{ieee_fullname}
-\bibliography{egbib}
+    \bibliography{contrastive-learning}
 }

### 2022-03-31 src/models/factory.py
adds back old school point contrast

@@ -16,15 +16,24 @@ from models.minkowski.dataset import (
     MinkowskiPretrainDataset,
     MinkowskiFrameDataset,
     MinkowskiS3DISDataset,
+    PointContrastPretrainDataset,
 )
 from util.types import DataInterface
 
 minkowski_name = "minkowski"
+pointcontrast_name = "pointcontrast"
 moco_name = "minkowski_moco"
 byol_name = "minkowski_byol"
 cme_name = "minkowski_cme"
 image_name = "image_pretrain"
-supported_models = [minkowski_name, moco_name, byol_name, cme_name, image_name]
+supported_models = [
+    minkowski_name,
+    moco_name,
+    byol_name,
+    cme_name,
+    image_name,
+    pointcontrast_name,
+]
 
 
 class ModelFactory:
@@ -51,6 +60,8 @@ class ModelFactory:
 
     def get_backbone_wrapper_type(self):
         if self.model_name == minkowski_name:
+            return MinkowskiBackboneTrainer
+        elif self.model_name == pointcontrast_name:
             return MinkowskiBackboneTrainer
         elif self.model_name == moco_name:
             return MinkowskiMocoBackboneTrainer
@@ -83,7 +94,10 @@ class ModelFactory:
     def get_backbone_dataset_type(self):
         if image_name == self.model_name:
             return ImagePretrainDataset
+        if self.model_name == pointcontrast_name:
+            return PointContrastPretrainDataset
         if minkowski_name in self.model_name:
             return MinkowskiPretrainDataset
+
         else:
             raise RuntimeError(self.error_msg)

### 2022-03-31 src/models/minkowski/dataset.py

@@ -166,6 +166,155 @@ def visualize_image(image):
     vis.show()
 
 
+class PointContrastPretrainDataset(PretrainDataset):
+    def __init__(self, scenes, cfg):
+        super(PointContrastPretrainDataset, self).__init__(scenes, cfg)
+
+        color_trans_ratio = 0.05
+        color_jitter_std = 0.005
+        self.scale_range = (0.9, 1.1)
+
+        ROTATION_AUGMENTATION_BOUND = (
+            (-np.pi / 32, np.pi / 32),
+            (-np.pi / 32, np.pi / 32),
+            (-np.pi, np.pi),
+        )
+
+        self.augmentations = transforms.Compose(
+            [
+                transforms.RandomHorizontalFlip("z", False),
+                transforms.ChromaticAutoContrast(),
+                transforms.ChromaticTranslation(color_trans_ratio),
+                transforms.ChromaticJitter(color_jitter_std),
+                transforms.RandomRotateZ(
+                    ROTATION_AUGMENTATION_BOUND=ROTATION_AUGMENTATION_BOUND
+                ),
+            ]
+        )
+
+    def collate(self, batch):
+
+        # Points stuff
+        coords1_list = [
+            frame["discrete_coords"]
+            for datapoint in batch
+            for frame in datapoint["quantized_frames"]
+            if datapoint
+        ]
+        features1_list = [
+            frame["unique_feats"]
+            for datapoint in batch
+            for frame in datapoint["quantized_frames"]
+            if datapoint
+        ]
+        # coords2_list = [
+        #     datapoint["quantized_frames"][1]["discrete_coords"]
+        #     for datapoint in batch
+        #     if datapoint
+        # ]
+        # features2_list = [
+        #     datapoint["quantized_frames"][1]["unique_feats"]
+        #     for datapoint in batch
+        #     if datapoint
+        # ]
+
+        points1, features1 = ME.utils.sparse_collate(coords1_list, features1_list)
+        # points2, features2 = ME.utils.sparse_collate(coords2_list, features2_list)
+
+        # Point to Point map
+        point_to_point_map = [
+            datapoint["point_to_point_map"] for datapoint in batch if datapoint
+        ]
+
+        pretrain_input = MinkowskiPretrainInputNew(
+            points1=points1,
+            features1=features1.float(),
+            points2=None,
+            features2=None,
+            point_to_pixel_maps1=None,
+            point_to_pixel_maps2=None,
+            images=None,
+            image_coordinates=None,
[…]

### 2022-03-31 src/models/minkowski/trainer.py

@@ -1022,9 +1022,9 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         return self._model.backbone
 
     def training_step(self, batch: MinkowskiPretrainInput, batch_idx: int):
-        model_input = ME.SparseTensor(batch.features.float(), batch.points)
-        output = self._model(model_input).output
-        loss = self.loss_fn(batch, output)
+        model_input1 = ME.SparseTensor(batch.features1.float(), batch.points1)
+        output1 = self._model(model_input1).output
+        loss = self.loss_fn(batch, output1)
 
         # Log losses
         log = functools.partial(self.log, on_step=True, on_epoch=True)
@@ -1036,9 +1036,9 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
         torch.cuda.empty_cache()
 
     def validation_step(self, batch: MinkowskiPretrainInput, batch_idx: int):
-        model_input = ME.SparseTensor(batch.features, batch.points)
-        output = self._model(model_input).output
-        loss = self.loss_fn(batch, output)
+        model_input1 = ME.SparseTensor(batch.features1.float(), batch.points1)
+        output1 = self._model(model_input1).output
+        loss = self.loss_fn(batch, output1)
         self.log("val_loss", loss, sync_dist=True)
 
     def loss_fn_new(self, batch, output):
@@ -1092,14 +1092,12 @@ class MinkowskiBackboneTrainer(BackboneTrainer):
 
         # Get all positive and negative pairs
         qs, ks = [], []
-        for i, matches in enumerate(batch.correspondences):
-            voxel_indices_1 = [match["frame1"]["voxel_inds"] for match in matches]
-            voxel_indices_2 = [match["frame2"]["voxel_inds"] for match in matches]
+        for i in range(batch.batch_size):
+            # voxel_indices_1 = [match["frame1"]["voxel_inds"] for match in matches]
+            # voxel_indices_2 = [match["frame2"]["voxel_inds"] for match in matches]
 
-            output_batch_1 = output.features_at(2 * i)
-            output_batch_2 = output.features_at(2 * i + 1)
-            q = output_batch_1[voxel_indices_1]
-            k = output_batch_2[voxel_indices_2]
+            q = output.features_at(2 * i)[batch.point_to_point_map[i][:, 0]]
+            k = output.features_at(2 * i + 1)[batch.point_to_point_map[i][:, 1]]
 
             # visualize_mapping(points1, points2, voxel_indices_1)
 

### 2022-04-01 doc/paper.tex
Adds pretrained and supervised results
adds first part of methods section

@@
 %%%%%%%%% TITLE - PLEASE UPDATE
-\title{Exploring Self-Supervised Learning for Indoor 3D scenes}
+\title{Self-Supervised Transfer Learning From 2D Images to 3D Point Clouds}
 
@@
 
-\section{Method}
-\label{sec:method}
+\section{Self-Supervised Contrastive learning}
+\label{sec:contrastiveLearning}
 
-\section{Feature Extractors}
+The goal of self-supervised transfer learning is to learn from unlabeled data using a self-supervised objective and use the initialized backbone parameters to improve performance of downstream tasks.
+Specifically, contrastive learning uses an objective function that encourages the outputs of a model on similar datapoints to be close together while keeping those of dissimilar datapoints to be far apart.
+
+Defining what datapoints are similar and dissimilar depends on the pretext task.
+Image colorization for example defines color pixels and their grayscale equivalent as similar and all other pixels as dissimilar.
+In this work, we adopt the instance discrimination task as in \cite{} whereby each input individually is taken to belong to its own class.
+
+Given a sampled point $\mathbf{x}$, positive points are sampled according to a positive distribution $\mathbf{x}^{+} \sim p^{+}( \cdot | \mathbf{x})$ which is commonly defined as a set of transformations $t$ of the original point.
+Often, positive pairs are derived by applying two different augmentations to the input point, $x = t(x), \quad x^{+} = t^{'}(x)$.
+Conversely, negative points are sampled according to $\mathbf{x}^{-} \sim p^{-}( \cdot | \mathbf{x})$ which in the context of an instance discriminative task means that they are taken as any other point not derived from the input point, regardless of augmentation.
+
+As for the final loss, we adopt InfoNCE (Info Noise Contrastive Estimation) loss which is a negative log probability of a point's similarity to its positive sample conditioned by the similarity of that point to all the negative samples.
+It is defined as:
+\begin{equation}
+    \mathcal{L}_{i} = -\log \frac{\exp(x_{i} \cdot x^{+}_{i} / \tau)}{\exp(x_{i} \cdot x^{+}_{i} / \tau) + \sum^{K}_{j}\exp(x_{i} \cdot x^{-}_{j} / \tau)}
+    \label{eq:contrastive_loss}
+\end{equation}
+
+Where $\tau \in (0,1]$ and represents the temperature parameter that controls the smoothness of the latent representations. Similarity between features is computed as the dot product.
+
+\subsection{Feature Extraction}
 \label{sec:method:featuresExtractors}
 
-\section{Extracting 2D features}
+Then we need an encoder and decoder.
+
+Throw away the end of the
+
+Use a ResNet and U-net following the same structure as was defined in...
+
+\subsection{Extracting 2D features}
 \label{sec:method:features2d}
 
-\section{Transfering 2D features}
+How the loss was formulated in 2D.
+
+What augmentations, cropping the image in two places, only using the points on the overlap regions.
+
+Add figure of what your feature extractor looks like
+
+\subsection{Transfering 2D features}
 \label{sec:method:features2D3D}
+
+How the loss was formulated in the 2D to 3D transfer.
 
@@
             \hline
-            Scratch                     & 91.53          & 96.79          & 81.25          & 0.0  & 29.81  & 56.92  & 64.77          & 74.43          & 86.35          & 55.24          & 68.38          & 79.7  & 55.65          & 64.68          \\
-            Supervised                  &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
+            Scratch                     & 91.53           & \textbf{96.79} & 81.25          & 0             & 29.81          & 56.92          & 64.77          & 74.43          & 86.35           & 55.24          & 68.38          & 79.7           & 55.65          & 64.68                   \\
+            Supervised                  & \textbf{91.81}  & 96.74          & \textbf{84.11} & \textbf{0.21} & \textbf{34.68} & \textbf{57.64} & \textbf{81.53} & \textbf{77.63} & \textbf{89.82}  & \textbf{85.76} & \textbf{76.03} & \textbf{79.27} & \textbf{57.67} & \textbf{70.22} (+5.54)  \\
             \hline
-            PointContrast               &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
+            PointContrast               & \textbf{92.118} & \textbf{96.93} & 82.2           & 0             & 24.94          & 53.78          & \textbf{76.91} & \textbf{75.94} & \textbf{87.929} & \textbf{70.27} & 71.196         & \textbf{77.76} & 56.36          & \textbf{66.641} (+1.73) \\
             PointContrast - No Overhang &                 &                &                &               &                &                &                &                &                 &                &                &                &                &                         \\
-            CME                         & \textbf{91.07} & \textbf{96.74} & \textbf{82.84} & 0    & 28.38  & 55.98  & \textbf{70.34} & \textbf{75.01} & \textbf{87.34} & \textbf{63.24} & \textbf{71.25} & 75.01 & \textbf{56.42} & \textbf{65.67} \\
+            CME                         & 91.07           & 96.74          & \textbf{82.84} & 0             & \textbf{28.38} & \textbf{55.98} & 70.34          & 75.01          & 87.34           & 63.24          & \textbf{71.25} & 75.01          & \textbf{56.42} & 65.67 (+0.99)           \\
         \end{tabular}
@@
 \label{sec:results:3d}
+
+Things to include in results
+\begin{itemize}
+    \item S3DIS semantic segmentation
+    \item S3DIS instance segmentation
+    \item Scannet semantic segmentation
+    \item Visualization of features in 2d
+    \item Visualization of features in 3D
+    \item Training and validation curves
+    \item Ablation studies
+\end{itemize}
 

### 2022-04-04 doc/paper.tex
adds latest results
updates scratch results
Updates methods section

--- 2022-03-31 doc/paper.tex
@@
 \usepackage[review]{cvpr}      % To produce the REVIEW version
-%\usepackage{cvpr}              % To produce the CAMERA-READY version
+% \usepackage{cvpr}              % To produce the CAMERA-READY version
 %\usepackage[pagenumbers]{cvpr} % To force page numbers, e.g. for an arXiv version
@@
 
+\newcommand{\AJ}[1]{{\color{red}{[Andrej: #1]}}}
+
 \begin{document}
@@
 
-\section{Method}
-\label{sec:method}
+\textbf{Sparse 3D convolutions.} Talk about submanifold networks, Minkowski Engine, etc
 
-\section{Feature Extractors}
-\label{sec:method:featuresExtractors}
+\textbf{Contrastive Learning.} Talk about SimCLR, BOYL, etc. Then talk about PointContrast.
 
-\section{Extracting 2D features}
+\textbf{Knowledge Distilation} Need to read up on this
+
+\section{Self-Supervised Contrastive learning}
+\label{sec:contrastiveLearning}
+
+The goal of self-supervised transfer learning is to learn from unlabeled data using a self-supervised objective which in turn initializes a model's backbone.
+The initialization is meant to improve performance of downstream tasks.
+Specifically, contrastive learning uses an objective function that encourages the outputs of a model on similar datapoints to be close together while keeping those of dissimilar datapoints to be far apart, in some metric space.
+
+Defining what datapoints are similar and dissimilar depends on the pretext task.
+Image colorization for example defines color pixels and their grayscale equivalent as similar and all other pixels as dissimilar.
+In this work, we adopt the instance discrimination task as in \cite{} whereby each input individually is taken to belong to its own class.
+
+We follow the formulation from \cite{le-khac_contrastive_2020}.
+Given a sampled point $\mathbf{x}$, positive points are sampled according to a positive distribution $\mathbf{x}^{+} \sim p^{+}( \cdot | \mathbf{x})$ which is commonly defined as a set of transformations $t$ of the original point.
+Often, positive pairs are derived by applying two different augmentations to the input point, $\mathbf{x}^{+} = t_{1}(\mathbf{x}), \, \mathbf{x} = t_{2}(\mathbf{x})$.
+Conversely, negative points are sampled according to $\mathbf{x}^{-} \sim p^{-}( \cdot | \mathbf{x})$ which in the context of an instance discriminative task means that they are taken as any other point not derived from the input point, regardless of its augmentation.
+
+This point is fed through an encoder $e()$ to get feature $\mathbf{v} = e(\mathbf{x}), \mathbf{v} \in \mathbb{R}^d$.
+This encoder is the backbone of the model we are trying to initialize.
+The embedding is subsequently passed through a decoder $h()$ to generate a normalized latent representation $\mathbf{z}=h(\mathbf{v})$ with $\{\mathbf{z} \, | \, \mathbf{z} \in \mathbb{R}^{m}, \lVert \mathbf{z} \lVert^{2} = 1\}$  and where $m \leq d$.
+Once pretrained, only the encoder parameters are retained as the initialized backbone.
+
+As for the objective function, we adopt an InfoNCE (Info Noise Contrastive Estimation) loss.
+It computes a negative log probability of an embedding's similarity to the embedding of its corresponding positive sample.
+This is conditioned by the similarity of that embedding to all the negative embeddings.
+It is defined as:
+\begin{equation}
+    \mathcal{L}_{i} = -\log \frac{\exp(\mathbf{z}_{i} \cdot \mathbf{z}^{+}_{i} / \tau)}{\exp(\mathbf{z}_{i} \cdot \mathbf{z}^{+}_{i} / \tau) + \sum^{K}_{j}\exp(\mathbf{z}_{i} \cdot \mathbf{z}^{-}_{j} / \tau)}
+    \label{eq:contrastive_loss}
+\end{equation}
+
+Where $\tau \in (0,1]$ and represents the temperature parameter that controls the smoothness of the latent representations.
+$K$ is a hyperparameter that represents the number of negative embeddings to sample.
+Similarity between features is computed as the dot product.
+
+\section{Feature Extraction}
+\label{sec:featureExtraction}
+
+This section goes through the types of backbones used as well as how features are extracted from 2D images and transferred to a 3D model.
+
+\AJ{Need a picture depicting the pipeline}
+
+\subsection{Extracting 2D features}
 \label{sec:method:features2d}
 
-\section{Transfering 2D features}
-\label{sec:method:features2D3D}
+We follow the ResUNet architecture from Godard \etal \cite{godard2019Digging}, which proved hugely successful in self-supervised monocular depth estimation.
+Its backbone consists of a ResNet34 with weights that were pretrained on ImageNet.
+Instead of estimating depth however, our decoder computes a 16 dimensional feature vector for each pixel in the input image.
 
-Talk about projection and Bi-linear interpolation
+Images are selected from the desired pre-training dataset.
+Positive samples are generated by randomly cropping the image, and then applying the following transformations: random horizontal flip, random gray-scale, gaussian blur and color jitter.
 
+Only pixels found in the overlap of the two crops are used in the objective function.
+Corresponding pixels are taken as positive samples while all others, including from other images in a batch are taken as negatives.
+
+This is then passed through the loss function described in \ref{eq:contrastive_loss}.
+
+\AJ{Add figure of what your feature extractor looks like}
+
+\subsection{Feature projection}
+\label{sec:featureProjection}
+
+Here we leverage that images are a projection of 3D space.
+We assume that from a given viewpoint we have both an image as well as a depth or range scan, which is often the case when capturing raw data.
+Unlike in \cite{xie2020pointcontrast}, we don't require a pose linking adjacent frames making our method applicable to more use cases.
+
+The two modalities can be mapped by projecting the point cloud into the color image using a standard pinhole camera model as defined below \cite{?}:
+
+\begin{equation}
+    \begin{bmatrix}
+        u \\
+        v \\
+        1
+    \end{bmatrix} = \frac{1}{z}
+    \begin{bmatrix}
+        f_u & 0   & c_u & 0 \\
+        0   & f_v & c_v & 0 \\
+        0   & 0   & 0   & 1 \\
+    \end{bmatrix}
+    \begin{bmatrix}
+        x \\
+        y \\
+        z \\
+        1 \\
+    \end{bmatrix}
+    \label{eq:pinhole}
+\end{equation}
+
+\AJ{Is this equation necessary or is this considered common knowledge}
+
+Where $f$ represents focal length, $c$ the center offset, $u,v$ the pixels of an image and $x,y,z$ the 3D coordinates of the input point.
+
+In the case of stereo cameras, this step is unnecessary as the mapping comes directly from the depth image.
+Note that its important to have the viewpoints of the 2D and 3D sensors be as close together as possible.
+This is exactly true for stereo cameras, but not guaranteed when using different sensors for each modality.
+Its important because occluded points can falsely be mapped to a pixel even though they are not directly visible in the image.
+Note also that using scans as opposed to any point cloud is also a limitation of \cite{xie2020pointcontrast}.
+
+If the viewpoints are slightly off or the point cloud is a reconstruction from multiple scans, hidden points can be removed using the algorithm from Katz \etal \cite{katz2007Direct}.
+
+After passing in the image through the model, the output is a scaled version of the original image.
+To complete the mapping, we can either take the feature with the closest feature pixel or interpolate the features using bi-linear interpolation.
+We will show an analysis of these two methods in section \ref{sec:results}
+
+\AJ{Do I need to give the equation for Bi-linear interpolation or is this also considered basic knowledge}
+
+\subsection{Feature Transfer}
+\label{sec:featureTransfer}
+
+With the mapping, we can now
+
+Transformations applied to 3D points.
+
+How the loss was formulated in the 2D to 3D transfer.
+
+What does the loss function look like?
+
+\section{Downstream Tasks}
+\label{sec:downstream}
+
+\newpage
 \section{Experiment Results}
@@
             \hline
-            Scratch                     & 91.53          & 96.79          & 81.25          & 0.0  & 29.81  & 56.92  & 64.77          & 74.43          & 86.35          & 55.24          & 68.38          & 79.7  & 55.65          & 64.68          \\
-            Supervised                  &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
+            Scratch                     & 90.34           & \textbf{96.84} & 80.63          & 0              & 25.5           & 56.95          & 62.93          & 75.14          & 86.21          & 65.77          & 72.12          & 76.91          & 56.4           & 65.06                   \\
+            Supervised                  & \textbf{91.81}  & 96.74          & \textbf{84.11} & \textbf{0.21}  & \textbf{34.68} & \textbf{57.64} & \textbf{81.53} & \textbf{77.63} & \textbf{89.82} & \textbf{85.76} & \textbf{76.03} & \textbf{79.27} & \textbf{57.67} & \textbf{70.22} (+5.16)  \\
             \hline
-            PointContrast               &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
-            PointContrast - No Overhang &                &                &                &      &        &        &                &                &                &                &                &       &                &                \\
-            CME                         & \textbf{91.07} & \textbf{96.74} & \textbf{82.84} & 0    & 28.38  & 55.98  & \textbf{70.34} & \textbf{75.01} & \textbf{87.34} & \textbf{63.24} & \textbf{71.25} & 75.01 & \textbf{56.42} & \textbf{65.67} \\
+            PointContrast               & \textbf{92.118} & \textbf{96.93} & 82.2           & 0              & 24.94          & 53.78          & \textbf{76.91} & \textbf{75.94} & 87.93          & 70.27          & 71.2           & \textbf{77.76} & 56.36          & \textbf{66.641} (+1.58) \\
+            PointContrast - No Overhang & 91.624          & 96.86          & 82.02          & 0              & 29.37          & \textbf{57.5}  & 64.93          & 75.49          & 87.36          & 59.32          & 71.42          & 76.61          & 56.11          & 65.28 (+0.22)           \\
+            CME                         & 90.8            & 96.57          & \textbf{82.52} & \textbf{0.110} & \textbf{30.94} & 56.94          & 68.71          & 75.12          & \textbf{88.48} & \textbf{71.7}  & \textbf{72.17} & 73.783         & \textbf{56.35} & 66.48 (+1.42)           \\
         \end{tabular}
@@
 
+Things to include in results
+\begin{itemize}
+    \item S3DIS semantic segmentation
+    \item S3DIS instance segmentation
+    \item Scannet semantic segmentation
+    \item Visualization of features in 2d
+    \item Visualization of features in 3D
+    \item Pretraining and training without colour information
+    \item Using projection of 3D points instead of direct scan using \cite{katz2007Direct}
+    \item Training and validation curves
+    \item Ablation studies
+\end{itemize}
+
 \subsection{Contrastive Loss in 2D}
@@
 
+\newpage
 %%%%%%%%% REFERENCES

### 2022-04-04 doc/paper.tex
Completes the first draft of the methods section

@@
 
+\AJ{Need to review how I do this in the code, I think I use a projector left over from BYOL}
+
 \subsection{Feature projection}
@@
 
-With the mapping, we can now
+Our last step, after generating 2D features and mapping them to 3D points is to train our 3D model to emulate our pre-trained features.
+We accomplish this using the same contrastive loss from \ref{eq:contrastive_loss} but with input embeddings $\mathbf{z}_i$ coming from the features of the 3D points and its positive and negative embeddings $\mathbf{z}_i^{+},\mathbf{z}_i^{-}$ coming from the corresponding 2D feature.
 
-Transformations applied to 3D points.
+Only the required transformations are applied to the 2D image.
+These being: resizing, square cropping and colour normalization.
+Point cloud coordinates are transformed with a random horizontal flip and random rotation around the $z$ axis while their colors are transformed using autocontrast, translation and jitter.
 
-How the loss was formulated in the 2D to 3D transfer.
+We use the same 3D model as in \cite{xie2020pointcontrast}. For downstream tasks, the final $1\times1$ convolution is thrown away as the decoder.
 
-What does the loss function look like?
+\AJ{Need to look at not using a projection head in 3D model.}
 
-\section{Downstream Tasks}
-\label{sec:downstream}
-
-\newpage
 \section{Experiment Results}

### 2022-04-05 doc/paper.tex
adds first set of results

@@
 
-\section{Experiment Results}
-\label{sec:results}
-
 \begin{table*}[t!]
@@
             PointContrast - No Overhang & 91.624          & 96.86          & 82.02          & 0              & 29.37          & \textbf{57.5}  & 64.93          & 75.49          & 87.36          & 59.32          & 71.42          & 76.61          & 56.11          & 65.28 (+0.22)           \\
-            CME                         & 90.8            & 96.57          & \textbf{82.52} & \textbf{0.110} & \textbf{30.94} & 56.94          & 68.71          & 75.12          & \textbf{88.48} & \textbf{71.7}  & \textbf{72.17} & 73.783         & \textbf{56.35} & 66.48 (+1.42)           \\
+            CME \textbf{(Ours)}         & 90.8            & 96.57          & \textbf{82.52} & \textbf{0.110} & \textbf{30.94} & 56.94          & 68.71          & 75.12          & \textbf{88.48} & \textbf{71.7}  & \textbf{72.17} & 73.783         & \textbf{56.35} & 66.48 (+1.42)           \\
         \end{tabular}
@@
 
+\section{Experiment Results}
+\label{sec:results}
+
+This section will show (1) that the learned 2D features improve 2D segmentation performance as well as show a clear mapping to their input image when visualized.
+Therefore these features serve as good candidates for initializing our 3D network.
+(2) That the transfer of features from 2D to 3D improve performance on various 3D tasks against training from scratch.
+(3) That the performance of PointContrast is wholly dependant on the use of overlapping scans that allow it to extrapolate data.
+
 \subsection{Datasets}
 \label{sec:results:datasets}
+
+Following the common practice of 3D contrastive learning \cite{xie2020pointcontrast,hou2021Exploring, zhang2021Self, jiang2021Guided}, we use the following the datasets.
+
+\textbf{Scannet \cite{Dai2017ScanNet}:} Comprises of roughly 1600 reconstructed scans of indoor scenes captured using a stereo camera, reconstructed using a pose graph and then formed into a mesh.
+The scenes are mostly of individual rooms and range in the square floor size of 2 meters to 10 meters with a standard ceiling height of about 3 meters.
+Included in the dataset is the raw data of individual scans with their corresponding images along with the semantic and instance labels for the reconstructed scenes.
+It contains 20 different classes.
+We use the predefined train and validation split, using the validation as our test set as well as is done in previous works.
+
+\textbf{S3DIS \cite{armeni20163D}}: Much smaller than Scannet, it comprises of only 300 scenes however their size varies quite drastically as some are of rooms and others are of entire auditoriums.
+The scenes were also captured using a stereo camera and are mainly of indoor office scenes.
+There are 13 different semantic classes and all scenes contain point level semantic and instance labels.
+As is commonly done \cite{xie2020pointcontrast, hou2021Exploring} we use the Area 5 split as our validation and test set.
+
+\textbf{semanticKITTI}: Time allowing.
+
+We use Scannet to pretrain out network as it is the largest.
+It shares many of the same classes and types of indoor scenes as S3DIS and so is a perfect candidate to show the use case of collecting large amounts of unlabeled data for a task and then labeling only a partial amount.
 
@@
 
-\subsection{Contrastive Loss in 3D}
+\textbf{Model:} The backbone network used is the same as that used in \cite{xie2020pointcontrast, hou2021Exploring} using the sparse convolutional library developed by \cite{choy20194d}.
+The backbone uses a Res16UNet34 structure with non-bottleneck blocks and a maximum feature embedding size of 256.
+The outputs of the model are directly used for semantic segmentation.
+
+Instance segmentation is done using the same backbone with the head from \cite{jiang2020pointgroup}.
+
+\textbf{Pretraining on 2D:} Batch size of 64 image pairs where for each image pair we select 4092 pixels to contrast in our loss function.
+The loss function uses a tau of 0.4, a learning rate of 0.01, SGD optimizer with a momentum of 0.9, dampning of 0.1 and weight decay of 0.004.
+We decay the learning rate according to an exponential scheduler with an exponential rate of 0.99.
+Trained for ??
+iterations.
+
+\textbf{Transfer:} Transfer learning is done using the scannet dataset.
+We use a batch size of 8 and select 2000 point-pixel combinations per datapoint.
+The images are only center cropped to fit into the 2D model.
+The 3D model uses a voxel size of 5cm.
+The hyperparameters are the same al those in 2D except the learning rate is set to 0.1.
+It is trained for ??
+time (I forget).
+
+\textbf{Downstream Tasks:} All downstream tasks are run using the pipeline and parameters from \cite{hou2021Exploring}. The only difference is that we train instead on a single GPU for 20k iterations
+
+\subsection{Contrastive Loss in 2D}
+\label{sec:results:2d}
+
+Before transferring features into 3D, we first want to verify that the features we are transferring are useful.
+To do this we first visualized the features by bringing them into a 1D color space using the t-SNE \cite{maaten2008Visualizing} algorithm.
+the colors are put together to make an image.
+The visualization along with the original image can be found in figure \ref{fig:features2dvis}.
+
+\AJ{Need to train a 2D segmentation with and without pre-training.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{images/experiments/25.02.2022-image-pretrain-vis2.png}
+    \includegraphics[width=8cm]{images/experiments/25.02.2022-image-pretrain-vis1.png}
+    \caption{Visualization of features from the pre-trained 2D backbone. Images on the left are the input and right are the output.}
+    \label{fig:features2dvis}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{images/experiments/image-to-point-vis1.png}
+    \includegraphics[width=8cm]{images/experiments/image-to-point-vis2.png}
+    \caption{\textbf{left} shows features of the 2D images and \textbf{right} shows the features that were transfered to the 3D point cloud.}
+    \label{fig:features2d-3dvis}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4cm]{images/experiments/scratch-3d.png}
+    \includegraphics[width=4cm]{images/experiments/scratch-3d-2.png}
+    \caption{Visualization of point features without pre-training.}
+    \label{fig:features2dScratchvis}
+\end{figure}
+
+\subsection{Transfering 2D features to 3D}
+\label{sec:results:2d3d}
+
+To verify that the features were trasnfered correctly, we used the same visualization approach taken in section \ref{sec:results:2d} but instead applied it to points.
+Figure \ref{fig:features2d-3dvis} shows the relation between the 2D and 3D features.
+Although they didn't transfer perfectly, they did on certain key areas for example on the bookshelves in the top image and the wall and table in the lower image.
+
+\subsection{Semantic Segmentation}
 \label{sec:results:3d}
+
+We first test the case of pretraining our model on ScannetV2 and finetuning on S3DIS.
+Our results are shown in table \ref{table:s3disResults}.
+Fully supervised pretraining has a drastic impact on the final downstream performance and is used as an upperbound on expected performance.
+Our algorithm is comparable to the performance of \cite{xie2020pointcontrast} however, that is without using overlapping 3D scans.
+
+When using point correspondences without overlapping scans, the performance gain disappears almost entirely.
+This was noted in their original work and suggests to us that it is learning better features only due to its ability to better extrapolate surfaces.
 
@@
     \item Scannet semantic segmentation
-    \item Visualization of features in 2d
+    \item Scannet semantic segmentation with different amounts of data
+    \item Visualization of features in 2D
     \item Visualization of features in 3D
@@
 \end{itemize}
-
-\subsection{Contrastive Loss in 2D}
-\label{sec:results:2d}
-
-\subsection{Transfering 2D features to 3D}
-\label{sec:results:2d3d}
 

### 2022-04-07 scripts/chains/old-loss.sh
cleanup

@@ -1,4 +1,0 @@
-#!/bin/bash
-
-bash scripts/pretrain.sh scripts/experiments/pretrain/old_loss.sh
-bash scripts/run.sh scripts/experiments/s3dis/minkowski-pretrained-old-loss.sh

### 2022-04-07 scripts/experiments/s3dis/minkowski-pretrained-byol.sh

@@ -1,24 +1,0 @@
-python src/main.py \
-    dataset_dir=$dataset_dir \
-    dataset=s3dis \
-    model=minkowski \
-    model.name=minkowski_byol \
-    tasks=["train","eval"] \
-    hydra.run.dir=$output_dir/pretrain/byol \
-    gpus=[0] \
-    dataset.batch_size=6 \
-    dataset.accumulate_grad_batches=8 \
-    dataset.voxel_size=0.05 \
-    model.optimizer.type=SGD \
-    model.optimizer.lr=0.2 \
-    model.scheduler.type=PolyLR \
-    model.scheduler.poly_power=0.9 \
-    model.scheduler.max_iter=10000 \
-    model.scheduler.interval=step \
-    model.scheduler.frequency=10 \
-    max_time="01:00:00:00" \
-    max_epochs=2000 \
-    max_steps=4000 \
-    check_val_every_n_epoch=20 \
-    model.train.train_workers=6 \
-    pretrain_checkpoint=\"last.ckpt\"

### 2022-04-07 scripts/experiments/s3dis/minkowski-pretrained-moco.sh

@@ -1,24 +1,0 @@
-python src/main.py \
-    dataset_dir=$dataset_dir \
-    dataset=s3dis \
-    model=minkowski \
-    model.name=minkowski_moco \
-    tasks=["train","eval"] \
-    hydra.run.dir=$output_dir/pretrain/moco \
-    gpus=[0] \
-    dataset.batch_size=6 \
-    dataset.accumulate_grad_batches=8 \
-    dataset.voxel_size=0.05 \
-    model.optimizer.type=SGD \
-    model.optimizer.lr=0.6 \
-    model.scheduler.type=PolyLR \
-    model.scheduler.poly_power=0.9 \
-    model.scheduler.max_iter=10000 \
-    model.scheduler.interval=step \
-    model.scheduler.frequency=10 \
-    max_time="01:00:00:00" \
-    max_epochs=2000 \
-    max_steps=4000 \
-    check_val_every_n_epoch=20 \
-    model.train.train_workers=6 \
-    pretrain_checkpoint=\"last.ckpt\"

### 2022-04-07 scripts/slurm/chains/new-loss.sh

@@ -1,8 +1,0 @@
-#!/bin/bash
-#SBATCH --gres=gpu:1       # Request GPU "generic resources"
-#SBATCH --cpus-per-task=10  # Cores proportional to GPUs: 6 on Cedar, 10 on Béluga, 16 on Graham.
-#SBATCH --mem=47000M       # Memory proportional to GPUs: 32000 Cedar, 47000 Béluga, 64000 Graham.
-#SBATCH --time=1-00:03:00     # DD-HH:MM:SS
-
-bash scripts/slurm/pretrain.sh scripts/experiments/pretrain/new_loss.sh
-bash scripts/slurm/run.sh scripts/experiments/s3dis/minkowski-pretrained-new-loss.sh

### 2022-04-07 scripts/slurm/run_scannet_narval.sh

@@ -1,15 +1,16 @@
 #!/bin/bash
+#SBATCH --account=rrg-jskelly
 #SBATCH --gres=gpu:1       # Request GPU "generic resources"
 #SBATCH --cpus-per-task=12  # Cores proportional to GPUs: 6 on Cedar, 10 on Béluga, 16 on Graham.
 #SBATCH --mem=40000M       # Memory proportional to GPUs: 32000 Cedar, 47000 Béluga, 64000 Graham.
 #SBATCH --time=1-00:00:00     # DD-HH:MM:SS
 
-# Setup default environment
-source scripts/slurm/base_narval.sh
+# # Setup default environment
+# source scripts/slurm/base_narval.sh
 
-# Untar data
-dataset=~/scratch/datasets/scannet_preprocessed_2cm_clean.tar
-tar -xf $dataset -C $SLURM_TMPDIR
+# # Untar data
+# dataset=~/scratch/datasets/scannet_preprocessed_2cm_clean.tar
+# tar -xf $dataset -C $SLURM_TMPDIR
 
 base_dir=$PWD
 source $SLURM_TMPDIR/env/bin/activate

### 2022-04-07 src/models/minkowski/trainer.py

@@ -98,366 +98,6 @@ class MinkowskiMLP(nn.Module):
 
         return output
 
-
-class MinkowskiBYOL(nn.Module):
-    def __init__(self, cfg: DictConfig, use_predictor=False):
-        nn.Module.__init__(self)
-
-        self.dataset_cfg = cfg.dataset
-        self.use_predictor = use_predictor
-        self.feature_dim = cfg.model.net.model_n_out
-        self.bn_momentum = cfg.model.net.bn_momentum
-        self.norm_type = NormType.BATCH_NORM
-
-        # Backbone
-        self.backbone = Res16UNet34C(3, self.feature_dim, cfg.model, D=3)
-
-        # Projection head
-        self.projector = MinkowskiMLP(
-            cfg,
-            self.feature_dim,
-            cfg.model.pretrain.loss.projector_hidden_size,
-            cfg.model.pretrain.loss.projector_output_size,
-        )
-
-        if use_predictor:
-            self.predictor = MinkowskiMLP(
-                cfg,
-                cfg.model.pretrain.loss.projector_output_size,
-                cfg.model.pretrain.loss.predictor_hidden_size,
-                cfg.model.pretrain.loss.projector_output_size,
-            )
-
-    def forward(self, input):
-        """Extract features and predict semantic class."""
-
-        # Get embedding
-        output = self.backbone(input)
-
-        # Get projected features
-        output = self.projector(output)
-
-        # Get predicted features if requested
-        if self.use_predictor:
-            output = self.predictor(output)
-
-        return output
-
-
-class MinkowskiBOYLBackboneTrainer(BackboneTrainer):
-    def __init__(
-        self,
-        cfg: DictConfig,
-    ):
-        super(MinkowskiBOYLBackboneTrainer, self).__init__(cfg)
-
-        # config
-        self.feature_dim = cfg.model.net.model_n_out
-        self.tau = cfg.model.pretrain.loss.byol_tau
-
-        # Loss type
-        self.criterion = nn.MSELoss()
-
-        # Encoders
-        self.model_online = MinkowskiBYOL(cfg, use_predictor=True)
-        self.model_target = MinkowskiBYOL(cfg, use_predictor=False)
-
-    @property
-    def model(self):
-        return self.model_online.backbone
-
-    def training_step(self, batch: MinkowskiPretrainInput, batch_idx: int):
-
-        # Get outputs
-        model_input = ME.SparseTensor(batch.features.float(), batch.points)
[…]

### 2022-04-08 src/models/minkowski/dataset.py
removes projection head from feature transfer

@@ -380,11 +380,7 @@ class MinkowskiPretrainDataset(PretrainDataset):
             for datapoint in batch
             if datapoint
         ]
-        # point_to_pixel_maps2 = [
-        #     datapoint["quantized_frames"][1]["point_to_pixel_map"]
-        #     for datapoint in batch
-        #     if datapoint
-        # ]
+
         images = [datapoint["image"] for datapoint in batch if datapoint]
         image_coordinates = [
             datapoint["image_coords"] for datapoint in batch if datapoint
@@ -402,25 +398,8 @@ class MinkowskiPretrainDataset(PretrainDataset):
             for datapoint in batch
             if datapoint
         ]
-        # coords2_list = [
-        #     datapoint["quantized_frames"][1]["discrete_coords"]
-        #     for datapoint in batch
-        #     if datapoint
-        # ]
-        # features2_list = [
-        #     datapoint["quantized_frames"][1]["unique_feats"]
-        #     for datapoint in batch
-        #     if datapoint
-        # ]
 
         points1, features1 = ME.utils.sparse_collate(coords1_list, features1_list)
-
-        # points2, features2 = ME.utils.sparse_collate(coords2_list, features2_list)
-
-        # # Point to Point map
-        # point_to_point_map = [
-        #     datapoint["point_to_point_map"] for datapoint in batch if datapoint
-        # ]
 
         pretrain_input = MinkowskiPretrainInputNew(
             points1=points1,
@@ -537,18 +516,7 @@ class MinkowskiPretrainDataset(PretrainDataset):
             frame.scan_points.shape[0], max_pos, replace=False
         )
 
-        # point_to_point_map = np.array(
-        #     [
-        #         [
-        #             quantized_frames[0]["inverse_mapping"][point_ind],
-        #             quantized_frames[1]["inverse_mapping"][point_ind],
-        #         ]
-        #         for point_ind in point_indices
-        #     ]
-        # )
-
         return {
-            # "point_to_point_map": point_to_point_map,
             "quantized_frames": quantized_frames,
             "image": image,
             "image_coords": coords.detach().cpu().numpy(),

### 2022-04-08 src/models/minkowski/trainer.py

@@ -98,6 +98,7 @@ class MinkowskiMLP(nn.Module):
 
         return output
 
+
 def embed_tsne(data):
     """
     N x D np.array data
@@ -120,10 +121,6 @@ class CMEBackboneTrainerFull(BackboneTrainer):
 
         # 3D feature extractor
         self.model = Res16UNet34C(3, self.feature_dim, cfg.model, D=3)
-
-        self.head = MinkowskiMLP(
-            cfg, self.feature_dim, self.feature_dim, self.dataset_cfg.classes
-        )
 
         self.criterion = nn.CrossEntropyLoss()
 

### 2022-04-11 scripts/experiments/pretrain/images_transfer.sh
puts back projection head

@@ -10,6 +10,6 @@ python src/main.py \
     model=minkowski \
     model.name=minkowski_cme \
     model.train.train_workers=12 \
-    model.net.model_n_out=16 \
+    dataset.classes=16 \
     check_val_every_n_epoch=1 \
     pretrain_checkpoint_2d=\"epoch=31-step=18879-val_loss=6.38.ckpt\"

### 2022-04-11 src/models/minkowski/trainer.py

@@ -32,44 +32,16 @@ from sklearn.manifold import TSNE
 log = logging.getLogger(__name__)
 
 
-class MinkovskiSemantic(nn.Module):
-    def __init__(
-        self, cfg: DictConfig, encoding_only=False, backbone=None, freeze_backbone=False
-    ):
-        nn.Module.__init__(self)
-
-        self.encoding_only = encoding_only
-        self.dataset_cfg = cfg.dataset
-        self.feature_dim = cfg.model.net.model_n_out
-        self.bn_momentum = cfg.model.net.bn_momentum
-        self.norm_type = NormType.BATCH_NORM
-
-        # Backbone
-        if backbone:
-            self.backbone = backbone
-        else:
-            self.backbone = Res16UNet34C(3, self.feature_dim, cfg.model, D=3)
-
-        # Freeze backbone if required
-        if cfg.model.net.freeze_backbone:
-            for param in self.backbone.parameters():
-                param.requires_grad = False
-
-        self.head = MinkowskiMLP(
-            cfg, self.feature_dim, self.feature_dim, self.dataset_cfg.classes
-        )
-
-    def forward(self, input):
-        """Extract features and predict semantic class."""
-
-        if self.encoding_only:
-            return self.backbone.encoder(input)
-        else:
-            # Get backbone features
-            output = self.backbone(input)
-            output = self.head(output)
-
-            return MinkowskiOutput(output=output, semantic_scores=output.F)
+def embed_tsne(data):
+    """
+    N x D np.array data
+    """
+    tsne = TSNE(n_components=1, verbose=1, perplexity=40, n_iter=300, random_state=0)
+    tsne_results = tsne.fit_transform(data)
+    tsne_results = np.squeeze(tsne_results)
+    tsne_min = np.min(tsne_results)
+    tsne_max = np.max(tsne_results)
+    return (tsne_results - tsne_min) / (tsne_max - tsne_min)
 
 
 class MinkowskiMLP(nn.Module):
@@ -99,18 +71,6 @@ class MinkowskiMLP(nn.Module):
         return output
 
 
-def embed_tsne(data):
-    """
-    N x D np.array data
-    """
-    tsne = TSNE(n_components=1, verbose=1, perplexity=40, n_iter=300, random_state=0)
-    tsne_results = tsne.fit_transform(data)
-    tsne_results = np.squeeze(tsne_results)
[…]

### 2022-04-11 src/main.py
adds image segmentation training

@@ -94,7 +94,9 @@ class Trainer:
 
         log.info("Loading data module")
         self.data_interface = DataInterfaceFactory(cfg).get_interface()
-        self.model_factory = ModelFactory(cfg, self.data_interface)
+        self.model_factory = ModelFactory(
+            cfg, self.data_interface, checkpoint_2d_path=self.checkpoint_2d_path
+        )
 
         if self.pretrain_checkpoint:
             log.info("Continuing pretraining from checkpoint.")

### 2022-04-11 src/models/factory.py

@@ -6,12 +6,13 @@ from models.minkowski.trainer import (
     MinkowskiBackboneTrainer,
     CMEBackboneTrainerFull,
     ImageTrainer,
+    ImageSegmentationTrainer,
 )
 from models.minkowski.dataset import (
     ImagePretrainDataset,
     MinkowskiDataset,
     MinkowskiPretrainDataset,
-    MinkowskiFrameDataset,
+    ImageSegmentationDataset,
     MinkowskiS3DISDataset,
     PointContrastPretrainDataset,
 )
@@ -21,21 +22,30 @@ minkowski_name = "minkowski"
 pointcontrast_name = "pointcontrast"
 cme_name = "minkowski_cme"
 image_name = "image_pretrain"
+image_seg_name = "image_segmentation"
 supported_models = [
     minkowski_name,
     cme_name,
     image_name,
     pointcontrast_name,
+    image_seg_name,
 ]
 
 
 class ModelFactory:
-    def __init__(self, cfg: DictConfig, data_interface: DataInterface, backbone=None):
+    def __init__(
+        self,
+        cfg: DictConfig,
+        data_interface: DataInterface,
+        backbone=None,
+        checkpoint_2d_path=None,
+    ):
         self.model_name = cfg.model.name
         self.dataset_name = cfg.dataset.name
         self.cfg = cfg
         self.data_interface = data_interface
         self.backbone = backbone
+        self.checkpoint_2d_path = checkpoint_2d_path
 
         # Ensure that the model is supported
         self.error_msg = f"model {self.model_name} is not supported"
@@ -43,7 +53,14 @@ class ModelFactory:
             raise RuntimeError(self.error_msg)
 
     def get_model(self):
-        return MinkowskiTrainer(self.cfg, self.data_interface, backbone=self.backbone)
+        if self.model_name == image_seg_name:
+            return ImageSegmentationTrainer(
+                self.cfg, checkpoint_2d_path=self.checkpoint_2d_path
+            )
+        else:
+            return MinkowskiTrainer(
+                self.cfg, self.data_interface, backbone=self.backbone
+            )
         # if minkowski_name in self.model_name:
         #     return MinkowskiTrainer(
         #         self.cfg, self.data_interface, backbone=self.backbone
@@ -75,6 +92,8 @@ class ModelFactory:
             raise RuntimeError(self.error_msg)
 
     def get_dataset_type(self):
+        if self.model_name == image_seg_name:
+            return ImageSegmentationDataset
[…]

### 2022-04-11 src/models/minkowski/dataset.py

@@ -15,14 +15,143 @@ from dataloaders import transform_coord
 from dataloaders import transforms
 
 from scipy.spatial import KDTree
+from util.utils import get_random_colour
 
 from models.minkowski.types import (
     MinkowskiInput,
-    MinkowskiPretrainInput,
     MinkowskiPretrainInputNew,
     ImagePretrainInput,
 )
 from dataloaders.datasets import PretrainDataset, SegmentationDataset
+
+
+class ImageSegmentationDataset(PretrainDataset):
+    def __init__(self, scenes, cfg, is_test=False):
+        super(ImageSegmentationDataset, self).__init__(scenes, cfg)
+
+        self.scenes = scenes
+        self.is_test = is_test
+        self.ignore_label = cfg.dataset.ignore_label
+
+        image_size = 224
+        crop = 0.4
+        self.image_augmentations = transform_coord.Compose(
+            [
+                transform_coord.RandomResizedCropCoord(image_size, scale=(crop, 1.0)),
+                transform_coord.RandomHorizontalFlipCoord(),
+                T.RandomApply([T.ColorJitter(0.4, 0.4, 0.4, 0.1)], p=0.8),
+                T.RandomGrayscale(p=0.2),
+                T.RandomApply([transforms.GaussianBlur([0.1, 2.0])], p=0.5),
+                T.ToTensor(),
+                T.Normalize(mean=[0.485, 0.456, 0.406], std=[0.229, 0.224, 0.225]),
+            ]
+        )
+
+        self.coord_augmentations = T.Compose(
+            [
+                T.ToTensor(),
+                T.Resize(256),
+                T.CenterCrop(image_size),
+            ]
+        )
+        self.test_augmentations = T.Compose(
+            [
+                T.Normalize(mean=[0.485, 0.456, 0.406], std=[0.229, 0.224, 0.225]),
+            ]
+        )
+
+    def get_coords(self, image):
+        height = image.shape[0]
+        width = image.shape[1]
+
+        # Create coordinate image
+        # dims are (2, W, H)
+        # where the first axis holds the (x,y) coordinates
+        coord_x = np.linspace(0, width - 1, width).reshape(1, 1, -1)
+        coord_x = np.repeat(coord_x, height, axis=1)
+        coord_y = np.linspace(0, height - 1, height).reshape(1, -1, 1)
+        coord_y = np.repeat(coord_y, width, axis=2)
+
[…]

### 2022-04-11 src/models/minkowski/trainer.py

@@ -272,6 +272,96 @@ class CMEBackboneTrainerFull(BackboneTrainer):
         return self._loss_fn(online_features, target_features)
 
 
+class ImageSegmentationTrainer(BackboneTrainer):
+    def __init__(self, cfg: DictConfig, checkpoint_2d_path=None):
+        super(ImageSegmentationTrainer, self).__init__(cfg)
+
+        # 2D feature extraction
+        if checkpoint_2d_path:
+            self.model = ImageTrainer.load_from_checkpoint(
+                cfg=cfg, checkpoint_path=checkpoint_2d_path
+            ).model
+        else:
+            self.model = FeatureDecoder()
+
+        self.head = MLP2d(16, inner_dim=16, out_dim=cfg.dataset.classes + 1)
+
+        self.semantic_criterion = nn.CrossEntropyLoss(
+            ignore_index=cfg.dataset.ignore_label
+        )
+
+    def training_step(self, batch: ImagePretrainInput, batch_idx: int):
+
+        # Get Encoder values
+        z1 = self.model(batch.images1)
+        output = self.head(z1)
+
+        semantic_predictions = output.reshape((-1, output.shape[1]))
+        semantic_labels = batch.labels.reshape((-1,)).long()
+        loss = self.semantic_criterion(semantic_predictions, semantic_labels)
+
+        # Log losses
+        log = functools.partial(self.log, on_step=True, on_epoch=True)
+        log("train_loss", loss)
+
+        return loss
+
+    def validation_step(self, batch: ImagePretrainInput, batch_idx: int):
+
+        # Get Encoder values
+        z1 = self.model(batch.images1)
+        output = self.head(z1)
+
+        semantic_predictions = output.reshape((-1, output.shape[1]))
+        semantic_labels = batch.labels.reshape((-1,)).long()
+        loss = self.semantic_criterion(semantic_predictions, semantic_labels)
+
+        self.log("val_loss", loss, sync_dist=True)
+
+    def _loss_fn(self, output_online, output_target):
+        q = nn.functional.normalize(output_online, dim=-1, p=2)
+        k = nn.functional.normalize(output_target, dim=-1, p=2)
+
+        tau = 0.4
+
+        # Labels
+        npos = q.shape[0]
+        labels = torch.arange(npos).to(q.device).long()
+
+        logits = torch.mm(q, k.transpose(1, 0))  # npos by npos
+        out = torch.div(logits, tau)
+        out = out.squeeze().contiguous()
+
[…]

### 2022-04-11 src/models/minkowski/types.py

@@ -38,6 +38,7 @@ class ImagePretrainInput:
     images2: torch.Tensor
     coords1: torch.Tensor
     coords2: torch.Tensor
+    labels: torch.Tensor
     correspondences: list
 
     batch_size: int

### 2022-04-11 src/util/utils.py

@@ -213,7 +213,7 @@ class Visualizer:
 
 
 def get_random_colour():
-    return np.random.choice(range(256), size=3).astype(np.float) / 255.0
+    return np.random.choice(range(256), size=3).astype(np.float) / 256.0
 
 
 def visualize_pointcloud(points, colours):

### 2022-04-12 src/datasets/measurements_new.py
temp changes

@@ -133,6 +133,11 @@ class SceneMeasurement:
         #     interpolation=cv2.INTER_NEAREST,
         # )
 
+        # Show label image
+        cv2.imwrite("instance.png", self.instance_image)
+        cv2.imwrite("semantic.png", self.label_image)
+        cv2.imwrite("color.png", self.color_image)
+
         # Convert depth map into point cloud
         self.scan_points, self.point_to_pixel_map = get_point_cloud(
             depth_image, self.depth_intrinsics
@@ -149,6 +154,13 @@ class SceneMeasurement:
 
         # Get points visible in camera frame
         frame, pixels = self.get_projected_scene(info, scene)
+
+        pcd = o3d.geometry.PointCloud()
+        pcd.points = o3d.utility.Vector3dVector(frame.points)
+        labels = np.zeros(frame.features.shape)
+        labels[:, 0] = frame.semantic_labels
+        pcd.colors = o3d.utility.Vector3dVector(labels)
+        points_map = self.remove_hidden_points(pcd)
 
         # # Get point cloud for camera view
         # pcd = o3d.geometry.PointCloud()
@@ -169,9 +181,13 @@ class SceneMeasurement:
         self.instance_labels = frame.instance_labels
 
         # Compute image labels
-        self.label_image[:, :, :] = frame.semantic_labels.max() + 1
-        for point_ind, (px, py) in enumerate(pixels.T):
-            self.label_image[int(py), int(px), :] = frame.semantic_labels[point_ind]
+        self.label_image[:, :, :] = frame.semantic_labels.max() + 20
+        tmp = frame.semantic_labels
+        tmp[frame.semantic_labels == -100] = 256
+        for point_ind in points_map:
+            # for point_ind, (px, py) in enumerate(pixels.T):
+            px, py = pixels.T[point_ind]
+            self.label_image[int(py), int(px), :] = tmp[point_ind]
 
         self.label_image = cv2.resize(
             self.label_image,
@@ -179,35 +195,44 @@ class SceneMeasurement:
             interpolation=cv2.INTER_NEAREST,
         )
 
-        # labels_to_colors = {}
-        # for i in range(100):
-        #     labels_to_colors[i] = (get_random_colour() * 255).astype(np.int)
-        # labels_to_colors[-100] = (get_random_colour() * 255).astype(np.int)
+        labels_to_colors = {}
+        for i in range(100):
[…]

### 2022-04-13 doc/paper.tex
adds initial results for scannet with varying data amounts
updates instance supervised results
adds partial data results
Adds results from their implementation of point contrast

@@
             \hline
-            PointContrast               & \textbf{92.118} & \textbf{96.93} & 82.2           & 0              & 24.94          & 53.78          & \textbf{76.91} & \textbf{75.94} & 87.93          & 70.27          & 71.2           & \textbf{77.76} & 56.36          & \textbf{66.641} (+1.58) \\
+            PointContrast               & \textbf{92.118} & \textbf{96.93} & 82.2           & 0              & 24.94          & 53.78          & \textbf{76.91} & 75.94          & 87.93          & 70.27          & 71.2           & \textbf{77.76} & 56.36          & \textbf{66.641} (+1.58) \\
+            PointContrast - theirs      & 91.97           & 96.67          & 82.4           & 0              & 17.84          & 56.49          & 75.39          & \textbf{77.63} & 87.35          & 63.3           & 73.53          & 79.55          & 58.76          & 66.22 (+1.16)           \\
             PointContrast - No Overhang & 91.624          & 96.86          & 82.02          & 0              & 29.37          & \textbf{57.5}  & 64.93          & 75.49          & 87.36          & 59.32          & 71.42          & 76.61          & 56.11          & 65.28 (+0.22)           \\
@@
     }
-    \caption{S3DIS results using ContrastiveSceneContexts Code}
+    \caption{S3DIS Semantic Segmentation results (mIOU)}
     \label{table:s3disResults}
 \end{table*}
+
+\begin{table*}[t!]
+    \centering
+    \resizebox{0.8\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c c c c c | c }
+                                   & beam & column        & window        & door          & table         & chair         & sofa          & bookcase      & board         & clutter       & mAP@0.5              \\
+            \hline
+            Scratch                & 0    & 12.8          & 61.7          & 92.0          & 51.4          & 87.3          & 62.4          & 38.5          & 87.0          & 37.1          & 53.0                 \\
+            Supervised             & 0    & 26.0          & 74.3          & 84.5          & 43.3          & 87.3          & \textbf{81.8} & 37.4          & \textbf{85.4} & \textbf{41.2} & \textbf{56.2} (+3.2) \\
+            \hline
+            PointContrast - theirs &      & 24.7          & 67.9          & 90.7          & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
+            PointContrast          & 0    & 17.4          & 68.2          & \textbf{93.0} & 45.7          & 88.5          & 63.6          & 34.1          & 82.1          & 38.2          & 53.1 (+0.1)          \\
+            CME \textbf{(Ours)}    & 0    & \textbf{35.0} & \textbf{70.0} & 83.4          & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 (+2.8)          \\
+        \end{tabular}
+    }
+    \caption{S3DIS Instance Segmentation Results (mAP@0.5)}
+    \label{table:s3disInstanceSegmention}
+\end{table*}
+
+\begin{table}[t!]
+    \centering
+
+    \begin{tabular}{ c | c c c c c c}
+                            & 5\%  & 10\% & 20\% & 30\% & 40\% & 100\% \\
+        \hline
+        Scratch             & 56.9 &      & 64.1 &      & 65.6 &       \\
+        \hline
+        PointContrast       &      &      &      &      &      &       \\
+        CME \textbf{(Ours)} & 60.8 &      & 64.8 &      & 64.7 &       \\
+    \end{tabular}
+
+    \caption{Scannet Semantic Segmentation with Partial Data  (mIOU)}
+    \label{table:scannetVaryingDataAmount}
+\end{table}
+
+% \begin{table}[t!]
+%     \centering
+
+%     \begin{tabular}{ c | c c c c c c }
+%                             & 5\% & 10\% & 20\% & 30\% & 40\% & 100\% \\
+%         \hline
+%         Scratch             &     &      &      &      &      &       \\
+%         \hline
+%         PointContrast       &     &      &      &      &      &       \\
+%         CME \textbf{(Ours)} &     &      &      &      &      &       \\
+%     \end{tabular}
+
+%     \caption{S3DIS Semantic Segmentation with Partial Data  (mIOU)}
+%     \label{table:s3disVaryingDataAmount}
+% \end{table}
 
@@
 We decay the learning rate according to an exponential scheduler with an exponential rate of 0.99.
-Trained for ??
-iterations.
+Trained for 20k iterations.
 
@@
 The hyperparameters are the same al those in 2D except the learning rate is set to 0.1.
-It is trained for ??
-time (I forget).
+It is trained for 20k iterations same as pre-training on 2D.
 

### 2022-04-13 doc/paper.tex
removes results of my implementation of pointcontrast

@@
             \hline
-            PointContrast               & \textbf{92.118} & \textbf{96.93} & 82.2           & 0              & 24.94          & 53.78          & \textbf{76.91} & 75.94          & 87.93          & 70.27          & 71.2           & \textbf{77.76} & 56.36          & \textbf{66.641} (+1.58) \\
-            PointContrast - theirs      & 91.97           & 96.67          & 82.4           & 0              & 17.84          & 56.49          & 75.39          & \textbf{77.63} & 87.35          & 63.3           & 73.53          & 79.55          & 58.76          & 66.22 (+1.16)           \\
+            PointContrast               & 91.97          & 96.67          & 82.4           & 0              & 17.84          & 56.49          & 75.39          & \textbf{77.63} & 87.35          & 63.3           & 73.53          & 79.55          & 58.76          & 66.22 (+1.16)          \\
             PointContrast - No Overhang & 91.624         & 96.86          & 82.02          & 0              & 29.37          & \textbf{57.5}  & 64.93          & 75.49          & 87.36          & 59.32          & 71.42          & 76.61          & 56.11          & 65.28 (+0.22)          \\
@@
             \hline
-            PointContrast - theirs &      & 24.7          & 67.9          & 90.7          & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
-            PointContrast          & 0    & 17.4          & 68.2          & \textbf{93.0} & 45.7          & 88.5          & 63.6          & 34.1          & 82.1          & 38.2          & 53.1 (+0.1)          \\
+            PointContrast       &      & 24.7          & 67.9          & 90.7 & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
             CME \textbf{(Ours)} & 0    & \textbf{35.0} & \textbf{70.0} & 83.4 & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 (+2.8)          \\

### 2022-04-18 doc/paper.tex
adds CSC semantic seg results

@@
             \hline
-            PointContrast               & 91.97          & 96.67          & 82.4           & 0              & 17.84          & 56.49          & 75.39          & \textbf{77.63} & 87.35          & 63.3           & 73.53          & 79.55          & 58.76          & 66.22 (+1.16)          \\
-            PointContrast - No Overhang & 91.624         & 96.86          & 82.02          & 0              & 29.37          & \textbf{57.5}  & 64.93          & 75.49          & 87.36          & 59.32          & 71.42          & 76.61          & 56.11          & 65.28 (+0.22)          \\
-            CME \textbf{(Ours)}         & 90.8           & 96.57          & \textbf{82.52} & \textbf{0.110} & \textbf{30.94} & 56.94          & 68.71          & 75.12          & \textbf{88.48} & \textbf{71.7}  & \textbf{72.17} & 73.783         & \textbf{56.35} & 66.48 (+1.42)          \\
+            PointContrast              & 91.97          & \textbf{96.67} & 82.4           & 0              & 17.84          & 56.49          & \textbf{75.39} & \textbf{77.63} & 87.35          & 63.3           & 73.53          & \textbf{79.55} & 58.76          & 66.22 (+1.16)          \\
+            PointContrast - Same Scan  & 91.03          & 96.27          & \textbf{82.8}  & 0              & 31.53          & 57.27          & 69.33          & 76.97          & \textbf{88.83} & 45.49          & 71.02          & 77.09          & 57.06          & 64.97 (-0.09)          \\
+            Contrastive Scene Contexts & \textbf{92.34} & 96.02          & 83.53          & 0              & \textbf{40.88} & \textbf{57.73} & 72.11          & 77.5           & 88.15          & \textbf{74.61} & \textbf{74.52} & 76.45          & \textbf{62.64} & \textbf{68.96} (+3.9)  \\
+            CME \textbf{(Ours)}        & 90.8           & 96.57          & 82.52          & \textbf{0.110} & 30.94          & 56.94          & 68.71          & 75.12          & 88.48          & 71.7           & 72.17          & 73.783         & 56.35          & 66.48 (+1.42)          \\
         \end{tabular}
@@
             \hline
-            PointContrast       &      & 24.7          & 67.9          & 90.7 & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
+            PointContrast       & 0    & 24.7          & 67.9          & 90.7 & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
             CME \textbf{(Ours)} & 0    & \textbf{35.0} & \textbf{70.0} & 83.4 & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 (+2.8)          \\
@@
 
+% \begin{table}[t!]
+%     \centering
+
+%     \begin{tabular}{ c | c c c c c c}
+%                             & 5\%  & 10\% & 20\% & 30\% & 40\% & 100\% \\
+%         \hline
+%         Scratch             & 56.9 &      & 64.1 &      & 65.6 &       \\
+%         \hline
+%         PointContrast       &      &      &      &      &      &       \\
+%         CME \textbf{(Ours)} & 60.8 &      & 64.8 &      & 64.7 &       \\
+%     \end{tabular}
+
+%     \caption{Scannet Semantic Segmentation with Partial Data  (mIOU)}
+%     \label{table:scannetVaryingDataAmount}
+% \end{table}
+
 \begin{table}[t!]
     \centering
-
+    \resizebox{0.4\textwidth}{!}{
         \begin{tabular}{ c | c c c c c c}
@@
             \hline
-        Scratch             & 56.9 &      & 64.1 &      & 65.6 &       \\
+            Scratch             & 42.2  & 51.09 & 58.23 & 59.23 & 62.17 & 67.43 \\
             \hline
-        PointContrast       &      &      &      &      &      &       \\
-        CME \textbf{(Ours)} & 60.8 &      & 64.8 &      & 64.7 &       \\
+            PointContrast       & 44.79 & 55.28 & 57.98 & 58.97 & 63.17 &       \\
+            CME \textbf{(Ours)} & 44.04 & 51.19 & 57.3  & 60.73 & 63.31 & 67.71 \\
         \end{tabular}
-
+    }
     \caption{Scannet Semantic Segmentation with Partial Data  (mIOU)}
@@
 \end{table}
-
-% \begin{table}[t!]
-%     \centering
-
-%     \begin{tabular}{ c | c c c c c c }
-%                             & 5\% & 10\% & 20\% & 30\% & 40\% & 100\% \\
-%         \hline
-%         Scratch             &     &      &      &      &      &       \\
-%         \hline
-%         PointContrast       &     &      &      &      &      &       \\
-%         CME \textbf{(Ours)} &     &      &      &      &      &       \\
-%     \end{tabular}
-
-%     \caption{S3DIS Semantic Segmentation with Partial Data  (mIOU)}
-%     \label{table:s3disVaryingDataAmount}
-% \end{table}
 
@@
 
+\textbf{IDEA:} To test the overlap hypothesis, try upping the min overlap ratio from 0.3 to 0.9.
+
 Things to include in results
@@
     \item Visualization of features in 3D
-    \item Pretraining and training without colour information
+          % \item Pretraining and training without colour information
     \item Using projection of 3D points instead of direct scan using \cite{katz2007Direct}

### 2022-04-19 doc/paper.tex
updates results

@@
             PointContrast - Same Scan    & 91.03          & 96.27          & \textbf{82.8}  & 0              & 31.53          & 57.27          & 69.33          & 76.97          & \textbf{88.83} & 45.49          & 71.02          & 77.09          & 57.06          & 64.97 (-0.09)          \\
+            PointContrast - 70\% overlap & 91.85          & 96.56          & 83.55          & 0              & 28.25          & 59.36          & 73.53          & 78.51          & 88.28          & 79.95          & 74.61          & 76             & 61.08          & 68.58 (+3.52)          \\
             Contrastive Scene Contexts   & \textbf{92.34} & 96.02          & 83.53          & 0              & \textbf{40.88} & \textbf{57.73} & 72.11          & 77.5           & 88.15          & \textbf{74.61} & \textbf{74.52} & 76.45          & \textbf{62.64} & \textbf{68.96} (+3.9)  \\
@@
             PointContrast              & 0    & 24.7          & 67.9          & 90.7 & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
+            Contrastive Scene Contexts & 0    & 32.9          & 69.2          & 93.0 & 47.6          & 87.5          & 80.9          & 39.6          & 87.5          & 40.4          & 57.8 (+4.8)          \\
             CME \textbf{(Ours)}        & 0    & \textbf{35.0} & \textbf{70.0} & 83.4 & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 (+2.8)          \\
@@
 
-% \begin{table}[t!]
-%     \centering
-
-%     \begin{tabular}{ c | c c c c c c}
-%                             & 5\%  & 10\% & 20\% & 30\% & 40\% & 100\% \\
-%         \hline
-%         Scratch             & 56.9 &      & 64.1 &      & 65.6 &       \\
-%         \hline
-%         PointContrast       &      &      &      &      &      &       \\
-%         CME \textbf{(Ours)} & 60.8 &      & 64.8 &      & 64.7 &       \\
-%     \end{tabular}
-
-%     \caption{Scannet Semantic Segmentation with Partial Data  (mIOU)}
-%     \label{table:scannetVaryingDataAmount}
-% \end{table}
-
-\begin{table}[t!]
+\begin{table*}[t!]
     \centering
-    \resizebox{0.4\textwidth}{!}{
+    \resizebox{0.8\textwidth}{!}{
         \begin{tabular}{ c | c c c c c c}
@@
             \hline
-            PointContrast       & 44.79 & 55.28 & 57.98 & 58.97 & 63.17 &       \\
-            CME \textbf{(Ours)} & 44.04 & 51.19 & 57.3  & 60.73 & 63.31 & 67.71 \\
+            PointContrast              & 44.79 (+2.59) & 55.28 (+4.19) & 57.98 (-0.25) & 58.97 (-0.26) & 63.17 (+1.00) & 66.88 (-0.55) \\
+            Contrastive Scene Contexts &               &               &               & 62.15 (+2.92) & 65.15 (+2.98) & 67.62 (+0.19) \\
+            CME \textbf{(Ours)}        & 44.04 (+1.84) & 52.97 (+1.88) & 58.6 (+0.37)  & 60.73 (+1.5)  & 63.31 (+1.14) & 67.71 (+0.28) \\
         \end{tabular}
@@
     \label{table:scannetVaryingDataAmount}
-\end{table}
+\end{table*}
+
+\begin{table*}[t!]
+    \centering
+    \resizebox{0.25\textwidth}{!}{
+        \begin{tabular}{c | c}
+                                       & mAP  \\
+            \hline
+            Scratch                    & 35.2 \\
+            \hline
+            PointContrast              & 36.7 \\
+            Contrastive Scene Contexts & 36.1 \\
+            CME \textbf{(Ours)}        & 37.7 \\
+        \end{tabular}
+    }
+    \caption{Scannet Object Detection using VoteNet  (mAP)}
+    \label{table:scannetObjectDetection}
+\end{table*}
 

### 2022-04-25 doc/paper.tex
updates results

--- 2022-04-18 doc/paper.tex
@@
 \usepackage{booktabs}
+\usepackage{rotating}
 
@@
             PointContrast - Same Scan              & 91.03          & 96.27          & \textbf{82.8}  & 0              & 31.53          & 57.27          & 69.33          & 76.97          & \textbf{88.83} & 45.49          & 71.02          & 77.09          & 57.06          & 64.97 (-0.09)          \\
+            PointContrast - 70\% overlap           & 91.85          & 96.56          & 83.55          & 0              & 28.25          & 59.36          & 73.53          & 78.51          & 88.28          & 79.95          & 74.61          & 76             & 61.08          & 68.58 (+3.52)          \\
             Contrastive Scene Contexts             & \textbf{92.34} & 96.02          & 83.53          & 0              & \textbf{40.88} & \textbf{57.73} & 72.11          & 77.5           & 88.15          & \textbf{74.61} & \textbf{74.52} & 76.45          & \textbf{62.64} & \textbf{68.96} (+3.9)  \\
+            Contrastive Scene Contexts - Same Scan & 91.68          & 96.59          & 83.21          & 0              & 40.38          & 58.12          & 70.42          & 76.77          & 87.13          & 69.33          & 68.77          & 82.34          & 58.45          & 67.94 (+2.88)          \\
             CME \textbf{(Ours)}                    & 90.8           & 96.57          & 82.52          & \textbf{0.110} & 30.94          & 56.94          & 68.71          & 75.12          & 88.48          & 71.7           & 72.17          & 73.783         & 56.35          & 66.48 (+1.42)          \\
@@
             PointContrast                          & 0    & 24.7          & 67.9          & 90.7 & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
+            PointContrast - Same Scan              & 0    & 19.8          & 70.7          & 92.1 & 48.4          & 90.5          & 81.8          & 38.4          & 90.4          & 34.1          & 56.6 (+3.6)          \\
+            PointContrast - 70\% overlap           & 0    & 22.6          & 71.6          & 94.6 & 45.8          & 87.9          & 71.7          & 37.0          & 81.0          & 40.9          & 55.3 (+2.3)          \\
+            Contrastive Scene Contexts             & 0    & 32.9          & 69.2          & 93.0 & 47.6          & 87.5          & 80.9          & 39.6          & 87.5          & 40.4          & 57.8 (+4.8)          \\
+            Contrastive Scene Contexts - Same Scan & 0    & 32.0          & 69.3          & 96.1 & 38.7          & 89.4          & 87.6          & 38.8          & 81.3          & 38.8          & 57.2 (+4.2)          \\
             CME \textbf{(Ours)}                    & 0    & \textbf{35.0} & \textbf{70.0} & 83.4 & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 (+2.8)          \\
@@
             \hline
-            PointContrast       & 44.79 & 55.28 & 57.98 & 58.97 & 63.17 &       \\
-            CME \textbf{(Ours)} & 44.04 & 51.19 & 57.3  & 60.73 & 63.31 & 67.71 \\
+            PointContrast              & 44.79 \textbf{(+2.59)} & 55.28 \textbf{(+4.19)} & 57.98 (-0.25)         & 58.97 (-0.26)          & 63.17 \textbf{(+1.00)} & 66.88 (-0.55)          \\
+            Contrastive Scene Contexts & 42.6 \textbf{(+0.4)}   & 53.03 \textbf{(+1.94)} & 57.28 (-0.95)         & 62.15 \textbf{(+2.92)} & 65.15 \textbf{(+2.98)} & 67.62 \textbf{(+0.19)} \\
+            CME \textbf{(Ours)}        & 44.04 \textbf{(+1.84)} & 52.97 \textbf{(+1.88)} & 58.6 \textbf{(+0.37)} & 60.73 \textbf{(+1.5)}  & 63.31 \textbf{(+1.14)} & 67.71 \textbf{(+0.28)} \\
         \end{tabular}
@@
     \label{table:scannetVaryingDataAmount}
-\end{table}
+\end{table*}
+
+\begin{table*}[t!]
+    \centering
+    \resizebox{\textwidth}{!}{
+        \begin{tabular}[b]{ c | l c c c c c c c c c c c c c c c c c | c }
+                                       & \begin{turn}{-90}cabinet \end{turn} & \begin{turn}{-90} bed  \end{turn} & \begin{turn}{-90} chair \end{turn} & \begin{turn}{-90} sofa \end{turn} & \begin{turn}{-90} table \end{turn} & \begin{turn}{-90} door \end{turn} & \begin{turn}{-90}  window \end{turn} & \begin{turn}{-90} bookshelf \end{turn} & \begin{turn}{-90} picture \end{turn} & \begin{turn}{-90} counter \end{turn} & \begin{turn}{-90} desk \end{turn} & \begin{turn}{-90} curtain \end{turn} & \begin{turn}{-90} refrigerator \end{turn} & \begin{turn}{-90}  Shower curtain \end{turn} & \begin{turn}{-90}  toilet \end{turn} & \begin{turn}{-90} sink  \end{turn} & \begin{turn}{-90} bathtub  \end{turn} & \begin{turn}{-90} otherfurniture \end{turn} & mAP@0.5     \\
+
+            scratch                    & 45.8                       & 72.7                       & 79.8                       & 60.6                       & 69.5                       & 37.8                       & 28.9                       & 44.2                       & 17.8                       & 16.6                       & 24.8                       & 24.0                       & 45.3                       & 63.4                       & 93.1                       & 46.6                       & 78.5                       & 33.4                       & 49.0        \\
+            \hline
+            PointContrast              & 45.6                       & 68.9                       & 76.0                       & 62.7                       & 65.2                       & 38.5                       & 33.2                       & 53.7                       & 15.9                       & 20.9                       & 31.6                       & 23.2                       & 36.6                       & 55.5                       & 1.0                        & 48.0                       & 76.9                       & 31.6                       & 49.1 (+0.1) \\
+            Contrastive Scene Contexts & 48.6                       & 63.5                       & 79.0                       & 64.3                       & 65.0                       & 35.9                       & 32.8                       & 48.8                       & 15.0                       & 27.1                       & 38.3                       & 19.1                       & 44.9                       & 50.4                       & 94.6                       & 49.1                       & 76.3                       & 34.3                       & 49.3 (+0.3) \\
+            CME \textbf{(Ours)}        & 45.3                       & 72.8                       & 76.2                       & 66.2                       & 63.2                       & 38.2                       & 25.4                       & 49.7                       & 18.6                       & 21.0                       & 34.3                       & 20.2                       & 39.2                       & 58.9                       & 94.8                       & 45.0                       & 73.9                       & 30.0                       & 48.5 (-0.5) \\
+            \hline
+        \end{tabular}
+    }
+    \caption{Scannet Instance Segmentation Results (mAP@0.5)}
+    \label{table:scannetInstanceSegmention}
+\end{table*}
+
+\begin{table*}[t!]
+    \centering
+    \resizebox{0.6\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c}
+                                       & 5\% & 10\% & 20\% & 30\% & 40\% & 100\%       \\
+            \hline
+            Scratch                    &     &      &      &      &      & 49.0        \\
+            \hline
+            PointContrast              &     &      & 40.5 & 43.0 & 43.9 & 49.1 (+0.1) \\
+            Contrastive Scene Contexts &     &      & 41.6 & 42.9 & 44.6 & 49.3 (+0.3) \\
+            CME \textbf{(Ours)}        &     &      & 39.1 & 42.4 & 43.6 & 48.5 (-0.5) \\
+        \end{tabular}
+    }
+    \caption{Scannet Instance Segmentation with Partial Data  (mAP@0.5)}
+    \label{table:scannetInstanceVaryingDataAmount}
+\end{table*}
+
+\begin{table*}[t!]
+    \centering
+    \resizebox{\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c c c c c c c c | c }
+                                       & ceiling & floor & wall & beam & column & window & door & table & chair & sofa & bookcase & board & clutter & mIOU \\
+            \hline
+            Scratch                    &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
+            \hline
+            PointContrast              &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
+            Contrastive Scene Contexts &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
+            CME \textbf{(Ours)}        &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
+        \end{tabular}
+    }
+    \caption{KITTI Semantic Segmentation results (mIOU)}
+    \label{table:kittiSemanticResults}
+\end{table*}
+
+\begin{table*}[t!]
+    \centering
+    \resizebox{0.5\textwidth}{!}{
+        \begin{tabular}{c | c c c c}
+                                       & Scannet & S3DIS & SunRGBD & semanticKITTI \\
+            \hline
+            Scratch                    & 35.2    &       & 32.0    &               \\
+            \hline
+            PointContrast              & 36.7    &       & 34.2    &               \\
+            Contrastive Scene Contexts & 36.1    &       & 33.2    &               \\
+            CME \textbf{(Ours)}        & 37.7    &       & 33.1    &               \\
+        \end{tabular}
+    }
+    \caption{Object Detection using VoteNet  (mAP)}
+    \label{table:scannetObjectDetection}
+\end{table*}
 

### 2022-04-25 doc/paper.tex
updates results

@@
             PointContrast - 70\% overlap              & 91.85          & 96.56          & 83.55          & 0              & 28.25          & 59.36          & 73.53          & 78.51          & 88.28          & 79.95          & 74.61          & 76             & 61.08          & 68.58 (+3.52)          \\
-            Contrastive Scene Contexts             & \textbf{92.34} & 96.02          & 83.53          & 0              & \textbf{40.88} & \textbf{57.73} & 72.11          & 77.5           & 88.15          & \textbf{74.61} & \textbf{74.52} & 76.45          & \textbf{62.64} & \textbf{68.96} (+3.9)  \\
+            Contrastive Scene Contexts                & \textbf{92.34} & 96.02          & 83.53          & 0              & \textbf{40.88} & \textbf{57.73} & 72.11          & 77.5           & 88.15          & \textbf{74.61} & \textbf{74.52} & 76.45          & \textbf{62.64} & 68.96 (+3.9)           \\
             Contrastive Scene Contexts - Same Scan    & 91.68          & 96.59          & 83.21          & 0              & 40.38          & 58.12          & 70.42          & 76.77          & 87.13          & 69.33          & 68.77          & 82.34          & 58.45          & 67.94 (+2.88)          \\
+            Contrastive Scene Contexts - 70\% overlap & 92.68          & 96.63          & 83.99          & 0              & 42.37          & 55.95          & 77.17          & 75.74          & 89.57          & 82.74          & 73.45          & 80.67          & 60.55          & \textbf{70.12} (+5.06) \\
             CME \textbf{(Ours)}                       & 90.8           & 96.57          & 82.52          & \textbf{0.110} & 30.94          & 56.94          & 68.71          & 75.12          & 88.48          & 71.7           & 72.17          & 73.783         & 56.35          & 66.48 (+1.42)          \\
@@
             Contrastive Scene Contexts - Same Scan    & 0    & 32.0          & 69.3          & 96.1 & 38.7          & 89.4          & 87.6          & 38.8          & 81.3          & 38.8          & 57.2 (+4.2)          \\
+            Contrastive Scene Contexts - 70\% overlap & 0    & 27.2          & 76.9          & 95.4 & 42.2          & 90.3          & 90.9          & 42.6          & 84.7          & 38.7          & \textbf{58.9} (+5.9) \\
             CME \textbf{(Ours)}                       & 0    & \textbf{35.0} & \textbf{70.0} & 83.4 & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 (+2.8)          \\
@@
     \centering
+    \resizebox{0.6\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c}
+                                       & 5\% & 10\% & 20\% & 30\% & 40\% & 100\%       \\
+            \hline
+            Scratch                    &     &      &      &      &      & 49.0        \\
+            \hline
+            PointContrast              &     &      & 40.5 & 43.0 & 43.9 & 49.1 (+0.1) \\
+            Contrastive Scene Contexts &     &      & 41.6 & 42.9 & 44.6 & 49.3 (+0.3) \\
+            CME \textbf{(Ours)}        &     &      & 39.1 & 42.4 & 43.6 & 48.5 (-0.5) \\
+        \end{tabular}
+    }
+    \caption{Scannet Instance Segmentation with Partial Data  (mAP@0.5)}
+    \label{table:scannetInstanceVaryingDataAmount}
+\end{table*}
+
+\begin{table*}[t!]
+    \centering
     \resizebox{\textwidth}{!}{
@@
     \centering
-    \resizebox{0.6\textwidth}{!}{
-        \begin{tabular}{ c | c c c c c c}
-                                       & 5\% & 10\% & 20\% & 30\% & 40\% & 100\%       \\
+    \resizebox{0.3\textwidth}{!}{
+        \begin{tabular}{ c | c }
+                                       & mIOU  \\
             \hline
-            Scratch                    &     &      &      &      &      & 49.0        \\
+            Scratch                    & 41.82 \\
             \hline
-            PointContrast              &     &      & 40.5 & 43.0 & 43.9 & 49.1 (+0.1) \\
-            Contrastive Scene Contexts &     &      & 41.6 & 42.9 & 44.6 & 49.3 (+0.3) \\
-            CME \textbf{(Ours)}        &     &      & 39.1 & 42.4 & 43.6 & 48.5 (-0.5) \\
-        \end{tabular}
-    }
-    \caption{Scannet Instance Segmentation with Partial Data  (mAP@0.5)}
-    \label{table:scannetInstanceVaryingDataAmount}
-\end{table*}
-
-\begin{table*}[t!]
-    \centering
-    \resizebox{\textwidth}{!}{
-        \begin{tabular}{ c | c c c c c c c c c c c c c | c }
-                                       & ceiling & floor & wall & beam & column & window & door & table & chair & sofa & bookcase & board & clutter & mIOU \\
-            \hline
-            Scratch                    &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
-            \hline
-            PointContrast              &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
-            Contrastive Scene Contexts &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
-            CME \textbf{(Ours)}        &         &       &      &      &        &        &      &       &       &      &          &       &         &      \\
+            PointContrast              & 38.67 \\
+            Contrastive Scene Contexts & 41.12 \\
+            CME \textbf{(Ours)}        & 40.37 \\
         \end{tabular}

### 2022-04-25 doc/paper.tex
updates results

@@
     \centering
-    \resizebox{0.6\textwidth}{!}{
+    \resizebox{0.5\textwidth}{!}{
         \begin{tabular}{ c | c c c c c c}
@@
             \hline
-            Scratch                    &     &      &      &      &      & 49.0        \\
+            Scratch                    & 26.9 & 31.1 & 40.7        & 42.2        & 45.7        & 49.0        \\
             \hline
-            PointContrast              &     &      & 40.5 & 43.0 & 43.9 & 49.1 (+0.1) \\
-            Contrastive Scene Contexts &     &      & 41.6 & 42.9 & 44.6 & 49.3 (+0.3) \\
-            CME \textbf{(Ours)}        &     &      & 39.1 & 42.4 & 43.6 & 48.5 (-0.5) \\
+            PointContrast              &      & 35.3 & 40.5 (-0.2) & 43.0 (+0.8) & 43.9 (-1.8) & 49.1 (+0.1) \\
+            Contrastive Scene Contexts &      & 35.5 & 41.6 (+0.9) & 42.9 (+0.7) & 44.6 (-1.1) & 49.3 (+0.3) \\
+            CME \textbf{(Ours)}        &      & 35.5 & 39.1 (-1.6) & 42.4 (+0.2) & 43.6 (-2.1) & 48.5 (-0.5) \\
         \end{tabular}
@@
     \resizebox{0.5\textwidth}{!}{
-        \begin{tabular}{c | c c c c}
-                                       & Scannet & S3DIS & SunRGBD & semanticKITTI \\
+        \begin{tabular}{c | c c c}
+                                       & Scannet & S3DIS & SunRGBD \\
             \hline
-            Scratch                    & 35.2    &       & 32.0    &               \\
+            Scratch                    & 35.2    &       & 32.0    \\
             \hline
-            PointContrast              & 36.7    &       & 34.2    &               \\
-            Contrastive Scene Contexts & 36.1    &       & 33.2    &               \\
-            CME \textbf{(Ours)}        & 37.7    &       & 33.1    &               \\
+            PointContrast              & 36.7    &       & 34.2    \\
+            Contrastive Scene Contexts & 36.1    &       & 33.2    \\
+            CME \textbf{(Ours)}        & 37.7    &       & 33.1    \\
         \end{tabular}

### 2022-04-29 doc/paper.tex
updates results

@@
             Contrastive Scene Contexts                & 42.6 \textbf{(+0.4)}   & 53.03 \textbf{(+1.94)} & 57.28 (-0.95)          & 62.15 \textbf{(+2.92)} & 65.15 \textbf{(+2.98)} & 67.62 \textbf{(+0.19)} \\
+            Contrastive Scene Contexts - 70\% overlap & 45.44 \textbf{(+3.24)} & 50.84 (-0.25)          & 59.26 \textbf{(+1.03)} & 59.72 \textbf{(+0.49)} & 62.72 \textbf{(+0.55)} & 68.01 \textbf{(+0.58)} \\
+            Contrastive Scene Contexts - same scan    & 44.09 \textbf{(+1.89)} & 50.88 (-0.21)          & 56.64 (-1.59)          & 58.9 (-0.33)           & 61.50 (-0.67)          & 67.37 (-0.06)          \\
             CME \textbf{(Ours)}                       & 44.04 \textbf{(+1.84)} & 52.97 \textbf{(+1.88)} & 58.6 \textbf{(+0.37)}  & 60.73 \textbf{(+1.5)}  & 63.31 \textbf{(+1.14)} & 67.71 \textbf{(+0.28)} \\
@@
     \centering
-    \resizebox{0.5\textwidth}{!}{
+    \resizebox{0.7\textwidth}{!}{
         \begin{tabular}{ c | c c c c c c}
@@
             \hline
-            PointContrast              &      & 35.3 & 40.5 (-0.2) & 43.0 (+0.8) & 43.9 (-1.8) & 49.1 (+0.1) \\
-            Contrastive Scene Contexts &      & 35.5 & 41.6 (+0.9) & 42.9 (+0.7) & 44.6 (-1.1) & 49.3 (+0.3) \\
-            CME \textbf{(Ours)}        &      & 35.5 & 39.1 (-1.6) & 42.4 (+0.2) & 43.6 (-2.1) & 48.5 (-0.5) \\
+            PointContrast              & 27.0 (+0.1) & 35.3 & 40.5 (-0.2) & 43.0 (+0.8) & 43.9 (-1.8) & 49.1 (+0.1) \\
+            Contrastive Scene Contexts & 29.8 (+2.9) & 35.5 & 41.6 (+0.9) & 42.9 (+0.7) & 44.6 (-1.1) & 49.3 (+0.3) \\
+            CME \textbf{(Ours)}        & 25.5 (-1.4) & 35.5 & 39.1 (-1.6) & 42.4 (+0.2) & 43.6 (-2.1) & 48.5 (-0.5) \\
         \end{tabular}
@@
     \centering
-    \resizebox{0.3\textwidth}{!}{
-        \begin{tabular}{ c | c }
-                                       & mIOU  \\
+    \resizebox{0.7\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c }
+                                       & 5\%                    & 10\%                   & 20\%                   & 30\%                   & 40\%                   & 100\%                 \\
             \hline
-            Scratch                    & 41.82 \\
+            Scratch                    & 39.2                   & 40.43                  & 41.06                  & 41.67                  & 40.79                  & 41.01                 \\
             \hline
-            PointContrast              & 38.67 \\
-            Contrastive Scene Contexts & 41.12 \\
-            CME \textbf{(Ours)}        & 40.37 \\
+            PointContrast              & 40.29 \textbf{(1.09)}  & 42.59 \textbf{(+2.16)} & 42.2  \textbf{(+1.14)} & 41.65 (-0.02)          & 42.99 \textbf{(+2.2)}  &                       \\
+            Contrastive Scene Contexts & 40.6 \textbf{(+1.4)}   & 42.21 \textbf{(+1.78)} & 42.26 \textbf{(+1.2)}  & 42.24 \textbf{(+0.57)} & 44.29 \textbf{(+3.5)}  & 43.01 \textbf{(+2.0)} \\
+            CME \textbf{(Ours)}        & 39.49 \textbf{(+0.29)} & 40.90 \textbf{(+0.47)} & 40.91 (-0.15)          & 41.82 \textbf{(+0.15)} & 41.77 \textbf{(+0.98)} &                       \\
         \end{tabular}
@@
         \begin{tabular}{c | c c c}
-                                       & Scannet & S3DIS & SunRGBD \\
+                                       & Scannet              & SunRGBD              \\
             \hline
-            Scratch                    & 35.2    &       & 32.0    \\
+            Scratch                    & 35.2                 & 32.0                 \\
             \hline
-            PointContrast              & 36.7    &       & 34.2    \\
-            Contrastive Scene Contexts & 36.1    &       & 33.2    \\
-            CME \textbf{(Ours)}        & 37.7    &       & 33.1    \\
+            PointContrast              & 36.7 \textbf{(+1.5)} & 34.2 \textbf{(+2.2)} \\
+            Contrastive Scene Contexts & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
+            CME \textbf{(Ours)}        & 37.7 \textbf{(+2.5)} & 33.1 \textbf{(+1.1)} \\
         \end{tabular}

### 2022-05-02 doc/paper.tex
updates results

@@
             \hline
-            Scratch                    & 26.9        & 31.1 & 40.7        & 42.2        & 45.7        & 49.0        \\
+            Scratch                    & 26.9                 & \textbf{31.1} & 40.7                 & 42.2                 & 45.7        & 49.0                 \\
             \hline
-            PointContrast              & 27.0 (+0.1) & 35.3 & 40.5 (-0.2) & 43.0 (+0.8) & 43.9 (-1.8) & 49.1 (+0.1) \\
-            Contrastive Scene Contexts & 29.8 (+2.9) & 35.5 & 41.6 (+0.9) & 42.9 (+0.7) & 44.6 (-1.1) & 49.3 (+0.3) \\
-            CME \textbf{(Ours)}        & 25.5 (-1.4) & 35.5 & 39.1 (-1.6) & 42.4 (+0.2) & 43.6 (-2.1) & 48.5 (-0.5) \\
+            PointContrast              & 27.0 \textbf{(+0.1)} & 35.3          & 40.5 (-0.2)          & 43.0 \textbf{(+0.8)} & 43.9 (-1.8) & 49.1 \textbf{(+0.1)} \\
+            Contrastive Scene Contexts & 29.8 \textbf{(+2.9)} & 35.5          & 41.6 \textbf{(+0.9)} & 42.9 \textbf{(+0.7)} & 44.6 (-1.1) & 49.3 \textbf{(+0.3)} \\
+            CME \textbf{(Ours)}        & 25.5 (-1.4)          & 35.5          & 39.1 (-1.6)          & 42.4 \textbf{(+0.2)} & 43.6 (-2.1) & 48.5 (-0.5)          \\
         \end{tabular}
@@
 
-\begin{table*}[t!]
-    \centering
-    \resizebox{\textwidth}{!}{
-        \begin{tabular}[b]{ c | l c c c c c c c c c c c c c c c c c | c }
-                                       & \begin{turn}{-90}cabinet \end{turn} & \begin{turn}{-90} bed  \end{turn} & \begin{turn}{-90} chair \end{turn} & \begin{turn}{-90} sofa \end{turn} & \begin{turn}{-90} table \end{turn} & \begin{turn}{-90} door \end{turn} & \begin{turn}{-90}  window \end{turn} & \begin{turn}{-90} bookshelf \end{turn} & \begin{turn}{-90} picture \end{turn} & \begin{turn}{-90} counter \end{turn} & \begin{turn}{-90} desk \end{turn} & \begin{turn}{-90} curtain \end{turn} & \begin{turn}{-90} refrigerator \end{turn} & \begin{turn}{-90}  Shower curtain \end{turn} & \begin{turn}{-90}  toilet \end{turn} & \begin{turn}{-90} sink  \end{turn} & \begin{turn}{-90} bathtub  \end{turn} & \begin{turn}{-90} otherfurniture \end{turn} & mAP@0.5     \\
+% \begin{table*}[t!]
+%     \centering
+%     \resizebox{\textwidth}{!}{
+%         \begin{tabular}[b]{ c | l c c c c c c c c c c c c c c c c c | c }
+%                                        & \begin{turn}{-90}cabinet \end{turn} & \begin{turn}{-90} bed  \end{turn} & \begin{turn}{-90} chair \end{turn} & \begin{turn}{-90} sofa \end{turn} & \begin{turn}{-90} table \end{turn} & \begin{turn}{-90} door \end{turn} & \begin{turn}{-90}  window \end{turn} & \begin{turn}{-90} bookshelf \end{turn} & \begin{turn}{-90} picture \end{turn} & \begin{turn}{-90} counter \end{turn} & \begin{turn}{-90} desk \end{turn} & \begin{turn}{-90} curtain \end{turn} & \begin{turn}{-90} refrigerator \end{turn} & \begin{turn}{-90}  Shower curtain \end{turn} & \begin{turn}{-90}  toilet \end{turn} & \begin{turn}{-90} sink  \end{turn} & \begin{turn}{-90} bathtub  \end{turn} & \begin{turn}{-90} otherfurniture \end{turn} & mAP@0.5     \\
 
-            scratch                    & 45.8                       & 72.7                       & 79.8                       & 60.6                       & 69.5                       & 37.8                       & 28.9                       & 44.2                       & 17.8                       & 16.6                       & 24.8                       & 24.0                       & 45.3                       & 63.4                       & 93.1                       & 46.6                       & 78.5                       & 33.4                       & 49.0        \\
-            \hline
-            PointContrast              & 45.6                       & 68.9                       & 76.0                       & 62.7                       & 65.2                       & 38.5                       & 33.2                       & 53.7                       & 15.9                       & 20.9                       & 31.6                       & 23.2                       & 36.6                       & 55.5                       & 1.0                        & 48.0                       & 76.9                       & 31.6                       & 49.1 (+0.1) \\
-            Contrastive Scene Contexts & 48.6                       & 63.5                       & 79.0                       & 64.3                       & 65.0                       & 35.9                       & 32.8                       & 48.8                       & 15.0                       & 27.1                       & 38.3                       & 19.1                       & 44.9                       & 50.4                       & 94.6                       & 49.1                       & 76.3                       & 34.3                       & 49.3 (+0.3) \\
-            CME \textbf{(Ours)}        & 45.3                       & 72.8                       & 76.2                       & 66.2                       & 63.2                       & 38.2                       & 25.4                       & 49.7                       & 18.6                       & 21.0                       & 34.3                       & 20.2                       & 39.2                       & 58.9                       & 94.8                       & 45.0                       & 73.9                       & 30.0                       & 48.5 (-0.5) \\
-            \hline
-        \end{tabular}
-    }
-    \caption{Scannet Instance Segmentation Results (mAP@0.5)}
-    \label{table:scannetInstanceSegmention}
-\end{table*}
+%             scratch                    & 45.8                       & 72.7                       & 79.8                       & 60.6                       & 69.5                       & 37.8                       & 28.9                       & 44.2                       & 17.8                       & 16.6                       & 24.8                       & 24.0                       & 45.3                       & 63.4                       & 93.1                       & 46.6                       & 78.5                       & 33.4                       & 49.0        \\
+%             \hline
+%             PointContrast              & 45.6                       & 68.9                       & 76.0                       & 62.7                       & 65.2                       & 38.5                       & 33.2                       & 53.7                       & 15.9                       & 20.9                       & 31.6                       & 23.2                       & 36.6                       & 55.5                       & 1.0                        & 48.0                       & 76.9                       & 31.6                       & 49.1 (+0.1) \\
+%             Contrastive Scene Contexts & 48.6                       & 63.5                       & 79.0                       & 64.3                       & 65.0                       & 35.9                       & 32.8                       & 48.8                       & 15.0                       & 27.1                       & 38.3                       & 19.1                       & 44.9                       & 50.4                       & 94.6                       & 49.1                       & 76.3                       & 34.3                       & 49.3 (+0.3) \\
+%             CME \textbf{(Ours)}        & 45.3                       & 72.8                       & 76.2                       & 66.2                       & 63.2                       & 38.2                       & 25.4                       & 49.7                       & 18.6                       & 21.0                       & 34.3                       & 20.2                       & 39.2                       & 58.9                       & 94.8                       & 45.0                       & 73.9                       & 30.0                       & 48.5 (-0.5) \\
+%             \hline
+%         \end{tabular}
+%     }
+%     \caption{Scannet Instance Segmentation Results (mAP@0.5)}
+%     \label{table:scannetInstanceSegmention}
+% \end{table*}
 
@@
             \hline
-            PointContrast              & 40.29 \textbf{(1.09)}  & 42.59 \textbf{(+2.16)} & 42.2  \textbf{(+1.14)} & 41.65 (-0.02)          & 42.99 \textbf{(+2.2)}  &                       \\
+            PointContrast              & 40.29 \textbf{(1.09)}  & 42.59 \textbf{(+2.16)} & 42.2  \textbf{(+1.14)} & 41.65 (-0.02)          & 42.99 \textbf{(+2.2)}  & 42.09 \textbf{(+1.08)} \\
             Contrastive Scene Contexts & 40.6 \textbf{(+1.4)}   & 42.21 \textbf{(+1.78)} & 42.26 \textbf{(+1.2)}  & 42.24 \textbf{(+0.57)} & 44.29 \textbf{(+3.5)}  & 43.01 \textbf{(+2.0)}  \\
-            CME \textbf{(Ours)}        & 39.49 \textbf{(+0.29)} & 40.90 \textbf{(+0.47)} & 40.91 (-0.15)          & 41.82 \textbf{(+0.15)} & 41.77 \textbf{(+0.98)} &                       \\
+            CME \textbf{(Ours)}        & 39.49 \textbf{(+0.29)} & 40.90 \textbf{(+0.47)} & 40.91 (-0.15)          & 41.82 \textbf{(+0.15)} & 41.77 \textbf{(+0.98)} & 42.04 \textbf{(+1.03)} \\
         \end{tabular}
@@
     \centering
-    \resizebox{0.5\textwidth}{!}{
+    \resizebox{0.4\textwidth}{!}{
         \begin{tabular}{c | c c c}

### 2022-05-04 doc/paper.tex
updates paper results section

@@
 \usepackage{rotating}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 
+\clearpage
+\section{Experiment Results}
+\label{sec:results}
+
+This section will analyze the performance of our pre-trained backbone on a variety of datasets from indoor stereo to outdoor lidar scenes.
+It will also compare our method to state of the art baselines on three different downstream tasks; these being semantic segmentation, instance segmentation and object detection.
+
+\subsection{Datasets}
+\label{sec:results:datasets}
+
+Following the common practice of 3D contrastive learning \cite{xie2020pointcontrast, hou2021Exploring, zhang2021Self, jiang2021Guided}, we use the following indoor stereo datasets and add an extra outdoor lidar dataset.
+
+\textbf{Scannet \cite{Dai2017ScanNet}:} Comprises of roughly 1600 reconstructed scenes of indoor environments captured using a stereo camera.
+The scenes are mostly of individual rooms and range in the square floor size of 2 meters to 10 meters with a standard ceiling height of about 3 meters.
+It contains 20 different classes.
+We use the dataset-defined train and validation split, using the validation as our test set.
+
+\textbf{S3DIS \cite{armeni20163D}:} Much smaller than Scannet, it comprises of only 300 reconstructed scenes however their size varies quite drastically as some are of rooms and others are of entire auditoriums.
+The scenes were also captured using a stereo camera and are mainly of indoor office scenes.
+There are 13 different semantic classes and all scenes contain point level semantic and instance labels.
+As is commonly done \cite{xie2020pointcontrast, hou2021Exploring} we use the Area 5 split as our validation and test set.
+
+\textbf{SunRGBD \cite{song2015sunrgbd, janoch2011category, xiao2013database, silberman2012indoor}:} Similar to Scannet and S3DIS, this dataset contains roughly 10k RGBD scans of indoor office environments.
+It has 3D bounding box annotations for 10 different classes, with large class overlap with Scannet and S3DIS.
+There is no scene reconstruction available.
+We use the dataset-defined train and validation split.
+
+\textbf{semanticKITTI \cite{behley2019semantic, geiger2012are}: } Contains roughly 25k laser scans of outdoor driving environments.
+It contains 20 different classes with 3D semantic and instance labels.
+The scans are of a full rotation around the vehicle, going out to a range of about 20 meters with the spatial resolution decreasing with distance.
+We use the sequences 1-7 and 9-10 as our training set and sequence 8 as our validation.
+
+We use Scannet to pretrain our backbone network as it is the largest indoor dataset.
+It shares many of the same classes and types of indoor scenes as S3DIS and SUNRBGD and so is a perfect candidate to show the use case of collecting large amounts of unlabeled data for a task and then labeling only a partial amount.
+We also use it to pre-train for downstream tasks with semanticKITTI.
+This is meant to explore how versatile the learned features are, since semanticKITTI uses a different sensor type in a different environment and with different object types than Scannet.
+
+In some experiments, we vary the amount of training labeled data on downstream tasks to see the effectiveness of dataset size on the self-supervised pre-training.
+
+\subsection{Baselines}
+\label{sec:results:baselines}
+
+To verify the effectiveness of our method, we compare against three state of the art baselines.
+These being PointContrast \cite{xie2020pointcontrast}, Contrastive Scene Contexts (CSC) \cite{hou2021Exploring} and DepthContrast \cite{zhang2021Self}.
+
+Both PointContrast and Contrastive Scene Contexts require pairs of scans with known poses and at least 30\% overlap while DepthContrast and our method operate on single scans.
+To test more fairly, we tried variations of PointContrast and CSC using the same scan for both positive pairs.
+We also tried another variation with a more strict overlap threshold of 70\% to see whether they were learning to extrapolate global scene contexts.
+Note also that due to resource limitations because DepthContrast originally requires 32 GPUs to train, we
+
+In cases where it's applicable, we also compare against a fully-supervised backbone, to give us an idea of a reasonable upper-bound.
+
+\subsection{Implementation Details}
+\label{sec:results:implementation}
+
+\textbf{Model:} The backbone network used is the same as that used in \cite{xie2020pointcontrast, hou2021Exploring} using the sparse convolutional library developed by \cite{choy20194d}.
+The backbone uses a Res16UNet34 structure with non-bottleneck blocks and a maximum feature embedding size of 256.
+The outputs of the model are directly used for semantic segmentation.
+
+Instance segmentation is done using the same backbone to extract features which are subsequently fed into the algorithm from \cite{jiang2020pointgroup}.
+
+Object detection is performed with VoteNet \cite{qi2019deep}, using the sparse voxel backbone instead of PointNet++.
+For more information regarding the network structure for these tasks, please refer to the supplementary material.
+
+\textbf{Pretraining on 2D:} We use a batch size of 64 image pairs.
+For each pair we sample 4092 pixels to contrast in our loss function from eqn.
+\ref{eq:contrastive_loss}.
+The loss function uses a $\tau$ of 0.4, a learning rate of 0.01, SGD optimizer with a momentum of 0.9, dampening of 0.1 and weight decay of 0.004.
+We decay the learning rate according to an exponential scheduler with an exponential rate of 0.99.
+Positive samples are generated using the following transformations: random resized crop, random horizontal flip, color jitter, random gray scale, gaussian blur and color normalization.
+We pre-train the image network for 20k iterations.
+
+\textbf{Transfer from 2D to 3D:} Transfer learning is done using the scannet dataset.
+We use a batch size of 8 scan-image pairs and select 2000 point-pixel correspondences per pair.
+The images are only center cropped to fit into the 2D model.
+We voxelize the points using a size of 5cm and transform them using a Random horizontal flip and random rotation around the z-axis.
+The colors are also transformed randomly with chromatic autocontrast, translation and jitter.
+The hyperparameters are the same as those in 2D except the learning rate is set to 0.1.
+It is trained for 20k iterations same as pre-training on 2D.
+
+\textbf{Downstream Tasks:} All downstream tasks are run using the pipeline and parameters from \cite{hou2021Exploring}.
+It may be noted that we get slightly different numbers from those oringinaly reported.
+This comes down to a couple of factors, mainly an update to the core sparse convolution library that is required to run on modern GPUs, as well as using a single GPU instead of 8.
+Because of the library update, all pre-training was run from scratch.
+
+\subsection{Contrastive Loss in 2D}
+\label{sec:results:2d}
+
+Before transferring features into 3D, we first want to verify that the features we are transferring have some connection to the original image.
+To do this we visualized the features by bringing them into a 1D color space using the t-SNE \cite{maaten2008Visualizing} algorithm.
+the colors are put together to make an image.
+The visualization along with the original image can be found in figure \ref{fig:features2dvis}.
+The visualizations show a clear mapping between input image and output feature.
+This suggests the features should act as a good baseline for initializing network backbones.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=7cm]{images/experiments/25.02.2022-image-pretrain-vis2.png}
+    \includegraphics[width=7cm]{images/experiments/25.02.2022-image-pretrain-vis1.png}
+    \caption{\textbf{Left} Input image. \textbf{Right} Visualization of features from the pre-trained 2D backbone}
+    \label{fig:features2dvis}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=7cm]{images/experiments/image-to-point-vis1.png}
+    \includegraphics[width=7cm]{images/experiments/image-to-point-vis2.png}
+    \caption{\textbf{Left} Visualization of transferred features on the 3D point cloud. \textbf{Right}  Visualization of features from the pre-trained 2D backbone.}
+    \label{fig:features2d-3dvis}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4cm]{images/experiments/scratch-3d.png}
+    \includegraphics[width=4cm]{images/experiments/scratch-3d-2.png}
+    \caption{Visualization of point features without pre-training.}
+    \label{fig:features2dScratchvis}
+\end{figure}
+
+\subsection{Transfering 2D features to 3D}
+\label{sec:results:2d3d}
+
+To verify that the features were transferred correctly, we used the same visualization approach taken in section \ref{sec:results:2d} but instead applied it to points.
+Figure \ref{fig:features2d-3dvis} shows the relation between the 2D and 3D features.
+Although the features did not transfer perfectly, they did on certain key areas.
+For example, on the wall and table in the top image and on the bookshelves in the bottom image.
+
+\subsection{Semantic Segmentation}
+\label{sec:results:semantic}
+
+We first test the case of pretraining our model on ScannetV2 and finetuning on S3DIS.
+Our results are shown in table \ref{table:s3disResults}.
+Fully supervised pretraining has a drastic impact on the final downstream performance and is used as an upperbound on expected performance.
+Our algorithm is comparable to the performance of \cite{xie2020pointcontrast} however, that is without using overlapping 3D scans.
+
+When using point correspondences without overlapping scans, the performance gain disappears almost entirely.
+This was noted in their original work and suggests to us that it is learning better features only due to its ability to better extrapolate surfaces.
+
+\subsection{Instance Segmentation}
+\label{sec:results:instance}
+
+\subsection{Object Detection}
+\label{sec:results:object}
+
+\subsection{Object Detection}
+\label{sec:results:object}
+
+Things to include in results
+\begin{itemize}
+    \item Training and validation curves
+    \item Ablation studies
+\end{itemize}
+
+\section{Discussion}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+\begin{itemize}
+    \item that transferring 2D features to 3D improves performance across different domains.
+    \item that although some algorithms perform extremely well in some domains, they fail in others. Or they might work well on one task but not another.
+    \item PointContrast \& CSC both ideally require overlapping scenes to perform properly whereas our algorithm does not.
+    \item Pre-training does little when large labeled datasets are used but are more effective when datasets are very small
+    \item Pre-training can help speed up training time on smaller datasets but has no impact on larger datasets.
+\end{itemize}
+
+\newpage
+%%%%%%%%% REFERENCES
+{\small
+    \bibliographystyle{ieee_fullname}
+    \bibliography{contrastive-learning}
+}
+
+\clearpage
+\section{Supplement}
+
 \begin{table*}[t!]
@@
             Scratch                                   & 90.34          & \textbf{96.84} & 80.63          & 0              & 25.5           & 56.95          & 62.93          & 75.14          & 86.21          & 65.77          & 72.12          & 76.91          & 56.4           & 65.06                           \\
-            Supervised                                & \textbf{91.81} & 96.74          & \textbf{84.11} & \textbf{0.21}  & \textbf{34.68} & \textbf{57.64} & \textbf{81.53} & \textbf{77.63} & \textbf{89.82} & \textbf{85.76} & \textbf{76.03} & \textbf{79.27} & \textbf{57.67} & \textbf{70.22} (+5.16) \\
+            Supervised                                & \textbf{91.81} & 96.74          & \textbf{84.11} & \textbf{0.21}  & \textbf{34.68} & \textbf{57.64} & \textbf{81.53} & \textbf{77.63} & \textbf{89.82} & \textbf{85.76} & \textbf{76.03} & \textbf{79.27} & \textbf{57.67} & \textbf{70.22} \textbf{(+5.16)} \\
             \hline
-            PointContrast                             & 91.97          & \textbf{96.67} & 82.4           & 0              & 17.84          & 56.49          & \textbf{75.39} & \textbf{77.63} & 87.35          & 63.3           & 73.53          & \textbf{79.55} & 58.76          & 66.22 (+1.16)          \\
+            PointContrast                             & 91.97          & \textbf{96.67} & 82.4           & 0              & 17.84          & 56.49          & \textbf{75.39} & \textbf{77.63} & 87.35          & 63.3           & 73.53          & \textbf{79.55} & 58.76          & 66.22 \textbf{(+1.16)}          \\
             PointContrast - Same Scan                 & 91.03          & 96.27          & \textbf{82.8}  & 0              & 31.53          & 57.27          & 69.33          & 76.97          & \textbf{88.83} & 45.49          & 71.02          & 77.09          & 57.06          & 64.97 (-0.09)                   \\
-            PointContrast - 70\% overlap              & 91.85          & 96.56          & 83.55          & 0              & 28.25          & 59.36          & 73.53          & 78.51          & 88.28          & 79.95          & 74.61          & 76             & 61.08          & 68.58 (+3.52)          \\
-            Contrastive Scene Contexts                & \textbf{92.34} & 96.02          & 83.53          & 0              & \textbf{40.88} & \textbf{57.73} & 72.11          & 77.5           & 88.15          & \textbf{74.61} & \textbf{74.52} & 76.45          & \textbf{62.64} & 68.96 (+3.9)           \\
-            Contrastive Scene Contexts - Same Scan    & 91.68          & 96.59          & 83.21          & 0              & 40.38          & 58.12          & 70.42          & 76.77          & 87.13          & 69.33          & 68.77          & 82.34          & 58.45          & 67.94 (+2.88)          \\
-            Contrastive Scene Contexts - 70\% overlap & 92.68          & 96.63          & 83.99          & 0              & 42.37          & 55.95          & 77.17          & 75.74          & 89.57          & 82.74          & 73.45          & 80.67          & 60.55          & \textbf{70.12} (+5.06) \\
-            CME \textbf{(Ours)}                       & 90.8           & 96.57          & 82.52          & \textbf{0.110} & 30.94          & 56.94          & 68.71          & 75.12          & 88.48          & 71.7           & 72.17          & 73.783         & 56.35          & 66.48 (+1.42)          \\
+            PointContrast - 70\% overlap              & 91.85          & 96.56          & 83.55          & 0              & 28.25          & 59.36          & 73.53          & 78.51          & 88.28          & 79.95          & 74.61          & 76             & 61.08          & 68.58 \textbf{(+3.52)}          \\
+            Contrastive Scene Contexts                & \textbf{92.34} & 96.02          & 83.53          & 0              & \textbf{40.88} & \textbf{57.73} & 72.11          & 77.5           & 88.15          & \textbf{74.61} & \textbf{74.52} & 76.45          & \textbf{62.64} & 68.96 \textbf{(+3.9)}           \\
+            Contrastive Scene Contexts - Same Scan    & 91.68          & 96.59          & 83.21          & 0              & 40.38          & 58.12          & 70.42          & 76.77          & 87.13          & 69.33          & 68.77          & 82.34          & 58.45          & 67.94 \textbf{(+2.88)}          \\
+            Contrastive Scene Contexts - 70\% overlap & 92.68          & 96.63          & 83.99          & 0              & 42.37          & 55.95          & 77.17          & 75.74          & 89.57          & 82.74          & 73.45          & 80.67          & 60.55          & \textbf{70.12} \textbf{(+5.06)} \\
+            DepthContrast                             & 90.99          & 95.45          & 80.99          & 0              & 32.52          & 51.94          & 61.57          & 74.55          & 87.28          & 71.62          & 71.34          & 67.71          & 57.85          & 64.91 (-0.15)                   \\
+            CME \textbf{(Ours)}                       & 90.8           & 96.57          & 82.52          & \textbf{0.110} & 30.94          & 56.94          & 68.71          & 75.12          & 88.48          & 71.7           & 72.17          & 73.783         & 56.35          & 66.48 \textbf{(+1.42)}          \\
         \end{tabular}
@@
             \hline
-            PointContrast                             & 0    & 24.7          & 67.9          & 90.7 & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 (+1.8)          \\
-            PointContrast - Same Scan                 & 0    & 19.8          & 70.7          & 92.1 & 48.4          & 90.5          & 81.8          & 38.4          & 90.4          & 34.1          & 56.6 (+3.6)          \\
-            PointContrast - 70\% overlap              & 0    & 22.6          & 71.6          & 94.6 & 45.8          & 87.9          & 71.7          & 37.0          & 81.0          & 40.9          & 55.3 (+2.3)          \\
-            Contrastive Scene Contexts                & 0    & 32.9          & 69.2          & 93.0 & 47.6          & 87.5          & 80.9          & 39.6          & 87.5          & 40.4          & 57.8 (+4.8)          \\
-            Contrastive Scene Contexts - Same Scan    & 0    & 32.0          & 69.3          & 96.1 & 38.7          & 89.4          & 87.6          & 38.8          & 81.3          & 38.8          & 57.2 (+4.2)          \\
-            Contrastive Scene Contexts - 70\% overlap & 0    & 27.2          & 76.9          & 95.4 & 42.2          & 90.3          & 90.9          & 42.6          & 84.7          & 38.7          & \textbf{58.9} (+5.9) \\
-            CME \textbf{(Ours)}                       & 0    & \textbf{35.0} & \textbf{70.0} & 83.4 & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 (+2.8)          \\
+            PointContrast                             & 0    & 24.7          & 67.9          & 90.7 & 43.8          & 92.0          & 72.7          & 32.3          & 85.6          & 38.0          & 54.8 \textbf{(+1.8)}          \\
+            PointContrast - Same Scan                 & 0    & 19.8          & 70.7          & 92.1 & 48.4          & 90.5          & 81.8          & 38.4          & 90.4          & 34.1          & 56.6 \textbf{(+3.6)}          \\
+            PointContrast - 70\% overlap              & 0    & 22.6          & 71.6          & 94.6 & 45.8          & 87.9          & 71.7          & 37.0          & 81.0          & 40.9          & 55.3 \textbf{(+2.3)}          \\
+            Contrastive Scene Contexts                & 0    & 32.9          & 69.2          & 93.0 & 47.6          & 87.5          & 80.9          & 39.6          & 87.5          & 40.4          & 57.8 \textbf{(+4.8)}          \\
+            Contrastive Scene Contexts - Same Scan    & 0    & 32.0          & 69.3          & 96.1 & 38.7          & 89.4          & 87.6          & 38.8          & 81.3          & 38.8          & 57.2 \textbf{(+4.2)}          \\
+            Contrastive Scene Contexts - 70\% overlap & 0    & 27.2          & 76.9          & 95.4 & 42.2          & 90.3          & 90.9          & 42.6          & 84.7          & 38.7          & \textbf{58.9} \textbf{(+5.9)} \\
+            DepthContrast                             & 0    & 40.3          & 68.4          & 74.9 & 42.2          & 87.8          & 62.4          & 33.3          & 78.1          & 35.5          & 52.3 (-0.7)                   \\
+            CME \textbf{(Ours)}                       & 0    & \textbf{35.0} & \textbf{70.0} & 83.4 & \textbf{53.9} & \textbf{88.9} & 71.7          & \textbf{42.2} & 76.1          & 36.7          & 55.8 \textbf{(+2.8)}          \\
         \end{tabular}
@@
 
-% \begin{table*}[t!]
-%     \centering
-%     \resizebox{\textwidth}{!}{
-%         \begin{tabular}[b]{ c | l c c c c c c c c c c c c c c c c c | c }
-%                                        & \begin{turn}{-90}cabinet \end{turn} & \begin{turn}{-90} bed  \end{turn} & \begin{turn}{-90} chair \end{turn} & \begin{turn}{-90} sofa \end{turn} & \begin{turn}{-90} table \end{turn} & \begin{turn}{-90} door \end{turn} & \begin{turn}{-90}  window \end{turn} & \begin{turn}{-90} bookshelf \end{turn} & \begin{turn}{-90} picture \end{turn} & \begin{turn}{-90} counter \end{turn} & \begin{turn}{-90} desk \end{turn} & \begin{turn}{-90} curtain \end{turn} & \begin{turn}{-90} refrigerator \end{turn} & \begin{turn}{-90}  Shower curtain \end{turn} & \begin{turn}{-90}  toilet \end{turn} & \begin{turn}{-90} sink  \end{turn} & \begin{turn}{-90} bathtub  \end{turn} & \begin{turn}{-90} otherfurniture \end{turn} & mAP@0.5     \\
-
-%             scratch                    & 45.8                       & 72.7                       & 79.8                       & 60.6                       & 69.5                       & 37.8                       & 28.9                       & 44.2                       & 17.8                       & 16.6                       & 24.8                       & 24.0                       & 45.3                       & 63.4                       & 93.1                       & 46.6                       & 78.5                       & 33.4                       & 49.0        \\
-%             \hline
-%             PointContrast              & 45.6                       & 68.9                       & 76.0                       & 62.7                       & 65.2                       & 38.5                       & 33.2                       & 53.7                       & 15.9                       & 20.9                       & 31.6                       & 23.2                       & 36.6                       & 55.5                       & 1.0                        & 48.0                       & 76.9                       & 31.6                       & 49.1 (+0.1) \\
-%             Contrastive Scene Contexts & 48.6                       & 63.5                       & 79.0                       & 64.3                       & 65.0                       & 35.9                       & 32.8                       & 48.8                       & 15.0                       & 27.1                       & 38.3                       & 19.1                       & 44.9                       & 50.4                       & 94.6                       & 49.1                       & 76.3                       & 34.3                       & 49.3 (+0.3) \\
-%             CME \textbf{(Ours)}        & 45.3                       & 72.8                       & 76.2                       & 66.2                       & 63.2                       & 38.2                       & 25.4                       & 49.7                       & 18.6                       & 21.0                       & 34.3                       & 20.2                       & 39.2                       & 58.9                       & 94.8                       & 45.0                       & 73.9                       & 30.0                       & 48.5 (-0.5) \\
-%             \hline
-%         \end{tabular}
-%     }
-%     \caption{Scannet Instance Segmentation Results (mAP@0.5)}
-%     \label{table:scannetInstanceSegmention}
-% \end{table*}
-
 \begin{table*}[t!]
@@
             Contrastive Scene Contexts & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
+            DepthContrast              & 33.9 (-1.3)          &                      \\
             CME \textbf{(Ours)}        & 37.7 \textbf{(+2.5)} & 33.1 \textbf{(+1.1)} \\
@@
 
-\section{Experiment Results}
-\label{sec:results}
-
-This section will show (1) that the learned 2D features improve 2D segmentation performance as well as show a clear mapping to their input image when visualized.
-Therefore these features serve as good candidates for initializing our 3D network.
-(2) That the transfer of features from 2D to 3D improve performance on various 3D tasks against training from scratch.
-(3) That the performance of PointContrast is wholly dependant on the use of overlapping scans that allow it to extrapolate data.
-
-\subsection{Datasets}
-\label{sec:results:datasets}
-
-Following the common practice of 3D contrastive learning \cite{xie2020pointcontrast,hou2021Exploring, zhang2021Self, jiang2021Guided}, we use the following the datasets.
-
-\textbf{Scannet \cite{Dai2017ScanNet}:} Comprises of roughly 1600 reconstructed scans of indoor scenes captured using a stereo camera, reconstructed using a pose graph and then formed into a mesh.
-The scenes are mostly of individual rooms and range in the square floor size of 2 meters to 10 meters with a standard ceiling height of about 3 meters.
-Included in the dataset is the raw data of individual scans with their corresponding images along with the semantic and instance labels for the reconstructed scenes.
-It contains 20 different classes.
-We use the predefined train and validation split, using the validation as our test set as well as is done in previous works.
-
-\textbf{S3DIS \cite{armeni20163D}}: Much smaller than Scannet, it comprises of only 300 scenes however their size varies quite drastically as some are of rooms and others are of entire auditoriums.
-The scenes were also captured using a stereo camera and are mainly of indoor office scenes.
-There are 13 different semantic classes and all scenes contain point level semantic and instance labels.
-As is commonly done \cite{xie2020pointcontrast, hou2021Exploring} we use the Area 5 split as our validation and test set.
-
-\textbf{semanticKITTI}: Time allowing.
-
-We use Scannet to pretrain out network as it is the largest.
-It shares many of the same classes and types of indoor scenes as S3DIS and so is a perfect candidate to show the use case of collecting large amounts of unlabeled data for a task and then labeling only a partial amount.
-
-\subsection{Implementation}
-\label{sec:results:implementation}
-
-\textbf{Model:} The backbone network used is the same as that used in \cite{xie2020pointcontrast, hou2021Exploring} using the sparse convolutional library developed by \cite{choy20194d}.
-The backbone uses a Res16UNet34 structure with non-bottleneck blocks and a maximum feature embedding size of 256.
-The outputs of the model are directly used for semantic segmentation.
-
-Instance segmentation is done using the same backbone with the head from \cite{jiang2020pointgroup}.
-
-\textbf{Pretraining on 2D:} Batch size of 64 image pairs where for each image pair we select 4092 pixels to contrast in our loss function.
-The loss function uses a tau of 0.4, a learning rate of 0.01, SGD optimizer with a momentum of 0.9, dampning of 0.1 and weight decay of 0.004.
-We decay the learning rate according to an exponential scheduler with an exponential rate of 0.99.
-Trained for 20k iterations.
-
-\textbf{Transfer:} Transfer learning is done using the scannet dataset.
-We use a batch size of 8 and select 2000 point-pixel combinations per datapoint.
-The images are only center cropped to fit into the 2D model.
-The 3D model uses a voxel size of 5cm.
-The hyperparameters are the same al those in 2D except the learning rate is set to 0.1.
-It is trained for 20k iterations same as pre-training on 2D.
-
-\textbf{Downstream Tasks:} All downstream tasks are run using the pipeline and parameters from \cite{hou2021Exploring}. The only difference is that we train instead on a single GPU for 20k iterations
-
-\subsection{Contrastive Loss in 2D}
-\label{sec:results:2d}
-
-Before transferring features into 3D, we first want to verify that the features we are transferring are useful.
-To do this we first visualized the features by bringing them into a 1D color space using the t-SNE \cite{maaten2008Visualizing} algorithm.
-the colors are put together to make an image.
-The visualization along with the original image can be found in figure \ref{fig:features2dvis}.
-
-\AJ{Need to train a 2D segmentation with and without pre-training.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=8cm]{images/experiments/25.02.2022-image-pretrain-vis2.png}
-    \includegraphics[width=8cm]{images/experiments/25.02.2022-image-pretrain-vis1.png}
-    \caption{Visualization of features from the pre-trained 2D backbone. Images on the left are the input and right are the output.}
-    \label{fig:features2dvis}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=8cm]{images/experiments/image-to-point-vis1.png}
-    \includegraphics[width=8cm]{images/experiments/image-to-point-vis2.png}
-    \caption{\textbf{left} shows features of the 2D images and \textbf{right} shows the features that were transfered to the 3D point cloud.}
-    \label{fig:features2d-3dvis}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=4cm]{images/experiments/scratch-3d.png}
-    \includegraphics[width=4cm]{images/experiments/scratch-3d-2.png}
-    \caption{Visualization of point features without pre-training.}
-    \label{fig:features2dScratchvis}
-\end{figure}
-
-\subsection{Transfering 2D features to 3D}
-\label{sec:results:2d3d}
-
-To verify that the features were trasnfered correctly, we used the same visualization approach taken in section \ref{sec:results:2d} but instead applied it to points.
-Figure \ref{fig:features2d-3dvis} shows the relation between the 2D and 3D features.
-Although they didn't transfer perfectly, they did on certain key areas for example on the bookshelves in the top image and the wall and table in the lower image.
-
-\subsection{Semantic Segmentation}
-\label{sec:results:3d}
-
-We first test the case of pretraining our model on ScannetV2 and finetuning on S3DIS.
-Our results are shown in table \ref{table:s3disResults}.
-Fully supervised pretraining has a drastic impact on the final downstream performance and is used as an upperbound on expected performance.
-Our algorithm is comparable to the performance of \cite{xie2020pointcontrast} however, that is without using overlapping 3D scans.
-
-When using point correspondences without overlapping scans, the performance gain disappears almost entirely.
-This was noted in their original work and suggests to us that it is learning better features only due to its ability to better extrapolate surfaces.
-
-\textbf{IDEA:} To test the overlap hypothesis, try upping the min overlap ratio from 0.3 to 0.9.
-
-Things to include in results
-\begin{itemize}
-    \item S3DIS semantic segmentation
-    \item S3DIS instance segmentation
-    \item Scannet semantic segmentation
-    \item Scannet semantic segmentation with different amounts of data
-    \item Visualization of features in 2D
-    \item Visualization of features in 3D
-          % \item Pretraining and training without colour information
-    \item Using projection of 3D points instead of direct scan using \cite{katz2007Direct}
-    \item Training and validation curves
-    \item Ablation studies
-\end{itemize}
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-\newpage
-%%%%%%%%% REFERENCES
-{\small
-    \bibliographystyle{ieee_fullname}
-    \bibliography{contrastive-learning}
-}
-
 \end{document}

### 2022-05-04 doc/paper.tex
reformat tables

@@
 This section will analyze the performance of our pre-trained backbone on a variety of datasets from indoor stereo to outdoor lidar scenes.
-It will also compare our method to state of the art baselines on three different downstream tasks; these being semantic segmentation, instance segmentation and object detection.
+It will also compare our method to state of the art baselines on three different downstream tasks: semantic segmentation, instance segmentation and object detection.
 
@@
 Both PointContrast and Contrastive Scene Contexts require pairs of scans with known poses and at least 30\% overlap while DepthContrast and our method operate on single scans.
-To test more fairly, we tried variations of PointContrast and CSC using the same scan for both positive pairs.
+For ablations, we tried variations of PointContrast and CSC using the same scan for both positive pairs.
 We also tried another variation with a more strict overlap threshold of 70\% to see whether they were learning to extrapolate global scene contexts.
-Note also that due to resource limitations because DepthContrast originally requires 32 GPUs to train, we
 
-In cases where it's applicable, we also compare against a fully-supervised backbone, to give us an idea of a reasonable upper-bound.
+Due to our own resource limitations, we had to run DepthContrast with a batch size of 32 on a single GPU instead of a batch size of 1024 split across 32 GPUs.
+We also ran it for 40 epochs instead of 400, which still takes twice as long as any other method.
+Since they claim to match PointContrast on semantic segmentation with their original setup, we are already comparing against state of the art and so any performance degradation would suggest using the less computationally intensive method is better.
+
+In cases where it's applicable, we also compare against a fully-supervised backbone, to give us an idea of a reasonable upper-bound on performance improvement.
 
@@
 
+\begin{table*}[t]
+    \centering
+    \resizebox{\textwidth}{!}{
+        \begin{tabular}{ c | c c | c c c c c c | c c c c c c |}
+                                &          &      & \multicolumn{6}{|c|}{Semantic Segmentation (mIOU)} & \multicolumn{6}{|c|}{Instance Segmentation (mAP@0.5)}                                                                                                                                                                                                                               \\
+                                & overlap  & Scan & 5\%                                                & 10\%                                                  & 20\%                   & 30\%                   & 40\%                   & 100\%                  & 5\%                  & 10\%          & 20\%                 & 30\%                 & 40\%        & 100\%                \\
+            \hline
+            Scratch             & --       & --   & 42.2                                               & 51.09                                                 & 58.23                  & 59.23                  & 62.17                  & 67.43                  & 26.9                 & \textbf{31.1} & 40.7                 & 42.2                 & 45.7        & 49.0                 \\
+            \hline
+            PointContrast       & $>$ 30\% & Pair & 44.79 \textbf{(+2.59)}                             & 55.28 \textbf{(+4.19)}                                & 57.98 (-0.25)          & 58.97 (-0.26)          & 63.17 \textbf{(+1.00)} & 66.88 (-0.55)          & 27.0 \textbf{(+0.1)} & 35.3          & 40.5 (-0.2)          & 43.0 \textbf{(+0.8)} & 43.9 (-1.8) & 49.1 \textbf{(+0.1)} \\
+            CSC                 & $>$ 30\% & Pair & 42.6 \textbf{(+0.4)}                               & 53.03 \textbf{(+1.94)}                                & 57.28 (-0.95)          & 62.15 \textbf{(+2.92)} & 65.15 \textbf{(+2.98)} & 67.62 \textbf{(+0.19)} & 29.8 \textbf{(+2.9)} & 35.5          & 41.6 \textbf{(+0.9)} & 42.9 \textbf{(+0.7)} & 44.6 (-1.1) & 49.3 \textbf{(+0.3)} \\
+            CSC                 & $>$ 70\% & Pair & 45.44 \textbf{(+3.24)}                             & 50.84 (-0.25)                                         & 59.26 \textbf{(+1.03)} & 59.72 \textbf{(+0.49)} & 62.72 \textbf{(+0.55)} & 68.01 \textbf{(+0.58)} & --                   & --            & --                   & --                   & --          & --                   \\
+            CSC                 & --       & Same & 44.09 \textbf{(+1.89)}                             & 50.88 (-0.21)                                         & 56.64 (-1.59)          & 58.9 (-0.33)           & 61.50 (-0.67)          & 67.37 (-0.06)          & --                   & --            & --                   & --                   & --          & --                   \\
+            CME \textbf{(Ours)} & --       & Same & 44.04 \textbf{(+1.84)}                             & 52.97 \textbf{(+1.88)}                                & 58.6 \textbf{(+0.37)}  & 60.73 \textbf{(+1.5)}  & 63.31 \textbf{(+1.14)} & 67.71 \textbf{(+0.28)} & 25.5 (-1.4)          & 35.5          & 39.1 (-1.6)          & 42.4 \textbf{(+0.2)} & 43.6 (-2.1) & 48.5 (-0.5)          \\
+        \end{tabular}
+    }
+    \caption{Scannet downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet)}
+    \label{table:scannetVaryingDataAmount}
+\end{table*}
+
+\begin{table}
+    \centering
+    \resizebox{0.45\textwidth}{!}{
+        \begin{tabular}{ c | c c c c }
+                                & min overlap & Same Scan  & Semantic (mIOU)                 & Instance (mAP@0.5)            \\
+            \hline
+            Scratch             & --          & --         & 65.06                           & 53.0                          \\
+            Supervised          & --          & --         & \textbf{70.22} \textbf{(+5.16)} & \textbf{56.2} \textbf{(+3.2)} \\
+            \hline
+            PointContrast       & 30\%        &            & 66.22 \textbf{(+1.16)}          & 54.8 \textbf{(+1.8)}          \\
+            PointContrast       & 70\%        &            & 68.58 \textbf{(+3.52)}          & 55.3 \textbf{(+2.3)}          \\
+            PointContrast       &             & \checkmark & 64.97 (-0.09)                   & 56.6 \textbf{(+3.6)}          \\
+            CSC                 & 30\%        &            & 68.96 \textbf{(+3.9)}           & 57.8 \textbf{(+4.8)}          \\
+            CSC                 & 70\%        &            & \textbf{70.12} \textbf{(+5.06)} & \textbf{58.9} \textbf{(+5.9)} \\
+            CSC                 &             & \checkmark & 67.94 \textbf{(+2.88)}          & 57.2 \textbf{(+4.2)}          \\
+            DepthContrast       &             & \checkmark & 64.91 (-0.15)                   & 52.3 (-0.7)                   \\
+            CME \textbf{(Ours)} &             & \checkmark & 66.48 \textbf{(+1.42)}          & 55.8 \textbf{(+2.8)}          \\
+        \end{tabular}
+    }
+    \caption{S3DIS downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet}
+    \label{table:s3disResults}
+\end{table}
+
+\begin{table*}[t]
+    \centering
+    \resizebox{0.7\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c}
+                                       & 5\%                  & 10\%          & 20\%                 & 30\%                 & 40\%        & 100\%                \\
+            \hline
+            Scratch                    & 26.9                 & \textbf{31.1} & 40.7                 & 42.2                 & 45.7        & 49.0                 \\
+            \hline
+            PointContrast              & 27.0 \textbf{(+0.1)} & 35.3          & 40.5 (-0.2)          & 43.0 \textbf{(+0.8)} & 43.9 (-1.8) & 49.1 \textbf{(+0.1)} \\
+            Contrastive Scene Contexts & 29.8 \textbf{(+2.9)} & 35.5          & 41.6 \textbf{(+0.9)} & 42.9 \textbf{(+0.7)} & 44.6 (-1.1) & 49.3 \textbf{(+0.3)} \\
+            CME \textbf{(Ours)}        & 25.5 (-1.4)          & 35.5          & 39.1 (-1.6)          & 42.4 \textbf{(+0.2)} & 43.6 (-2.1) & 48.5 (-0.5)          \\
+        \end{tabular}
+    }
+    \caption{Scannet Instance Segmentation with Partial Data  (mAP@0.5)}
+    \label{table:scannetInstanceVaryingDataAmount}
+\end{table*}
+
 \subsection{Instance Segmentation}
@@
 
-\newpage
+\clearpage
 %%%%%%%%% REFERENCES

### 2022-05-05 doc/paper.tex
updates tables

@@
 
+% \subsection{Evaluation Metrics}
+% \label{sec:results:metrics}
+% \dots
+
 \subsection{Contrastive Loss in 2D}
@@
 
-\begin{table*}[t]
+\begin{table*}[t!]
     \centering
@@
             CSC                 & --       & Same & 44.09 \textbf{(+1.89)}                             & 50.88 (-0.21)                                         & 56.64 (-1.59)          & 58.9 (-0.33)           & 61.50 (-0.67)          & 67.37 (-0.06)          & --                   & --            & --                   & --                   & --          & --                   \\
+            DepthContrast       & --       & Same & 49.27 \textbf{(+7.07)}                             & 54.27 \textbf{(+3.18)}                                & 58.21 (-0.02)          &                        &                        &                        & 24.9 (-2.0)          & 33.8          & 38.2 (-2.5)          & --                   & --          & --                   \\
             CME \textbf{(Ours)} & --       & Same & 44.04 \textbf{(+1.84)}                             & 52.97 \textbf{(+1.88)}                                & 58.6 \textbf{(+0.37)}  & 60.73 \textbf{(+1.5)}  & 63.31 \textbf{(+1.14)} & 67.71 \textbf{(+0.28)} & 25.5 (-1.4)          & 35.5          & 39.1 (-1.6)          & 42.4 \textbf{(+0.2)} & 43.6 (-2.1) & 48.5 (-0.5)          \\
@@
         \begin{tabular}{ c | c c c c }
-                                & min overlap & Same Scan  & Semantic (mIOU)                 & Instance (mAP@0.5)            \\
+                                & overlap  & Scan & Semantic (mIOU)                 & Instance (mAP@0.5)            \\
             \hline
@@
             \hline
-            PointContrast       & 30\%        &            & 66.22 \textbf{(+1.16)}          & 54.8 \textbf{(+1.8)}          \\
-            PointContrast       & 70\%        &            & 68.58 \textbf{(+3.52)}          & 55.3 \textbf{(+2.3)}          \\
-            PointContrast       &             & \checkmark & 64.97 (-0.09)                   & 56.6 \textbf{(+3.6)}          \\
-            CSC                 & 30\%        &            & 68.96 \textbf{(+3.9)}           & 57.8 \textbf{(+4.8)}          \\
-            CSC                 & 70\%        &            & \textbf{70.12} \textbf{(+5.06)} & \textbf{58.9} \textbf{(+5.9)} \\
-            CSC                 &             & \checkmark & 67.94 \textbf{(+2.88)}          & 57.2 \textbf{(+4.2)}          \\
-            DepthContrast       &             & \checkmark & 64.91 (-0.15)                   & 52.3 (-0.7)                   \\
-            CME \textbf{(Ours)} &             & \checkmark & 66.48 \textbf{(+1.42)}          & 55.8 \textbf{(+2.8)}          \\
+            PointContrast       & $>$ 30\% & Pair & 66.22 \textbf{(+1.16)}          & 54.8 \textbf{(+1.8)}          \\
+            PointContrast       & $>$ 70\% & Pair & 68.58 \textbf{(+3.52)}          & 55.3 \textbf{(+2.3)}          \\
+            PointContrast       &          & Same & 64.97 (-0.09)                   & 56.6 \textbf{(+3.6)}          \\
+            CSC                 & $>$ 30\% & Pair & 68.96 \textbf{(+3.9)}           & 57.8 \textbf{(+4.8)}          \\
+            CSC                 & $>$ 70\% & Pair & \textbf{70.12} \textbf{(+5.06)} & \textbf{58.9} \textbf{(+5.9)} \\
+            CSC                 &          & Same & 67.94 \textbf{(+2.88)}          & 57.2 \textbf{(+4.2)}          \\
+            DepthContrast       &          & Same & 64.91 (-0.15)                   & 52.3 (-0.7)                   \\
+            CME \textbf{(Ours)} &          & Same & 66.48 \textbf{(+1.42)}          & 55.8 \textbf{(+2.8)}          \\
         \end{tabular}
@@
     \label{table:s3disResults}
+\end{table}
+
+\begin{table}
+    \centering
+    \resizebox{0.4\textwidth}{!}{
+        \begin{tabular}{c | c c c}
+                                       & Scannet              & SunRGBD              \\
+            \hline
+            Scratch                    & 35.2                 & 32.0                 \\
+            \hline
+            PointContrast              & 36.7 \textbf{(+1.5)} & 34.2 \textbf{(+2.2)} \\
+            Contrastive Scene Contexts & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
+            DepthContrast              & 33.9 (-1.3)          & 32.9 \textbf{(+0.9)} \\
+            CME \textbf{(Ours)}        & 37.7 \textbf{(+2.5)} & 33.1 \textbf{(+1.1)} \\
+        \end{tabular}
+    }
+    \caption{Object Detection using VoteNet  (mAP)}
+    \label{table:scannetObjectDetection}
 \end{table}

### 2022-05-05 doc/paper.tex
fix up tables

--- 2022-05-04 doc/paper.tex
@@
 Because of the library update, all pre-training was run from scratch.
+
+% \subsection{Evaluation Metrics}
+% \label{sec:results:metrics}
+% \dots
 
@@
 
+\begin{table*}[t!]
+    \centering
+    \resizebox{\textwidth}{!}{
+        \begin{tabular}{ c | c c | c c c c c c | c c c c c c |}
+                                &      &          & \multicolumn{6}{|c|}{Semantic Segmentation (mIOU)} & \multicolumn{6}{|c|}{Instance Segmentation (mAP@0.5)}                                                                                                                                                                                                                        \\
+                                & Scan & Overlap  & 5\%                                                & 10\%                                                  & 20\%                 & 30\%                 & 40\%                 & 100\%                 & 5\%                  & 10\%          & 20\%                 & 30\%                 & 40\%        & 100\%                \\
+            \hline
+            Scratch             & --   & --       & 42.2                                               & 51.1                                                  & 58.2                 & 59.2                 & 62.2                 & 67.4                  & 26.9                 & \textbf{31.1} & 40.7                 & 42.2                 & 45.7        & 49.0                 \\
+            \hline
+            PointContrast       & Pair & $>$ 30\% & 44.8 \textbf{(+2.6)}                               & 55.3 \textbf{(+4.2)}                                  & 58.0 (-0.2)          & 59.0 (-0.2)          & 63.2 \textbf{(+1.0)} & 66.9 (-0.5)           & 27.0 \textbf{(+0.1)} & 35.3          & 40.5 (-0.2)          & 43.0 \textbf{(+0.8)} & 43.9 (-1.8) & 49.1 \textbf{(+0.1)} \\
+            PointContrast       & Pair & $>$ 70\% & 46.1 \textbf{(+3.9)}                               & 50.7 (-0.4)                                           & 58.0 (-0.2)          & 61.9 \textbf{(+2.7)} & 62.6 \textbf{(+0.4)} & 67.3 (-0.1)           & --                   & --            & --                   & --                   & --          & --                   \\
+            PointContrast       & Same & --       & 45.5 \textbf{(+3.3)}                               & 48.8 (-2.3)                                           & 56.7 (-1.5)          & 59.3 \textbf{(+0.1)} & 61.7 (-0.5)          & 66.5 (-0.9)           & --                   & --            & --                   & --                   & --          & --                   \\
+            CSC                 & Pair & $>$ 30\% & 42.6 \textbf{(+0.4)}                               & 53.0 \textbf{(+1.9)}                                  & 57.3 (-0.9)          & 62.2 \textbf{(+3.0)} & 65.2 \textbf{(+3.0)} & 67.6 \textbf{(+0.2)}  & 29.8 \textbf{(+2.9)} & 35.5          & 41.6 \textbf{(+0.9)} & 42.9 \textbf{(+0.7)} & 44.6 (-1.1) & 49.3 \textbf{(+0.3)} \\
+            CSC                 & Pair & $>$ 70\% & 45.4 \textbf{(+3.2)}                               & 50.8 (-0.3)                                           & 59.3 \textbf{(+1.1)} & 59.7 \textbf{(+0.5)} & 62.7 \textbf{(+0.5)} & 68.0 \textbf{(+0.6)}  & --                   & --            & --                   & --                   & --          & --                   \\
+            CSC                 & Same & --       & 44.1 \textbf{(+1.9)}                               & 50.9 (-0.2)                                           & 56.6 (-1.6)          & 58.9 (-0.3)          & 61.5 (-0.7)          & 67.4 \textbf{(+0.0)}  & --                   & --            & --                   & --                   & --          & --                   \\
+            DepthContrast       & Same & --       & 49.3 \textbf{(+7.1)}                               & 54.3 \textbf{(+3.2)}                                  & 58.2 \textbf{(+0.0)} & --                   & --                   & --                    & 24.9 (-2.0)          & 33.8          & 38.2 (-2.5)          & --                   & --          & --                   \\
+            CME \textbf{(Ours)} & Same & --       & 44.0 \textbf{(+1.8)}                               & 53.0 \textbf{(+1.9)}                                  & 58.6 \textbf{(+0.4)} & 60.7 \textbf{(+1.5)} & 63.3 \textbf{(+1.1)} & 67.7 \textbf{(+0.28)} & 25.5 (-1.4)          & 35.5          & 39.1 (-1.6)          & 42.4 \textbf{(+0.2)} & 43.6 (-2.1) & 48.5 (-0.5)          \\
+        \end{tabular}
+    }
+    \caption{Scannet downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet}
+    \label{table:scannetVaryingDataAmount}
+\end{table*}
+
+\begin{table}
+    \centering
+    \resizebox{0.45\textwidth}{!}{
+        \begin{tabular}{ c | c c c c }
+                                & Scan & Overlap  & Semantic (mIOU)               & Instance (mAP@0.5)            \\
+            \hline
+            Scratch             & --   & --       & 65.1                          & 53.0                          \\
+            Supervised          & --   & --       & \textbf{70.2} \textbf{(+5.1)} & \textbf{56.2} \textbf{(+3.2)} \\
+            \hline
+            PointContrast       & Pair & $>$ 30\% & 66.2 \textbf{(+1.1)}          & 54.8 \textbf{(+1.8)}          \\
+            PointContrast       & Pair & $>$ 70\% & 68.6 \textbf{(+3.5)}          & 55.3 \textbf{(+2.3)}          \\
+            PointContrast       & Same & --       & 65.0 (-0.1)                   & 56.6 \textbf{(+3.6)}          \\
+            CSC                 & Pair & $>$ 30\% & 69.0 \textbf{(+3.9)}          & 57.8 \textbf{(+4.8)}          \\
+            CSC                 & Pair & $>$ 70\% & \textbf{70.1} \textbf{(+5.0)} & \textbf{58.9} \textbf{(+5.9)} \\
+            CSC                 & Same & --       & 67.9 \textbf{(+2.8)}          & 57.2 \textbf{(+4.2)}          \\
+            DepthContrast       & Same & --       & 64.9 (-0.2)                   & 52.3 (-0.7)                   \\
+            CME \textbf{(Ours)} & Same & --       & 66.5 \textbf{(+1.4)}          & 55.8 \textbf{(+2.8)}          \\
+        \end{tabular}
+    }
+    \caption{S3DIS downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet}
+    \label{table:s3disResults}
+\end{table}
+
+\begin{table}
+    \centering
+    \resizebox{0.5\textwidth}{!}{
+        \begin{tabular}{ c | c c c c c c }
+                                & 5\%                  & 10\%                 & 20\%                  & 30\%                 & 40\%                 & 100\%                \\
+            \hline
+            Scratch             & 39.2                 & 40.4                 & 41.1                  & 41.7                 & 40.8                 & 41.0                 \\
+            \hline
+            PointContrast       & 40.3 \textbf{(1.1)}  & 42.6 \textbf{(+2.2)} & 42.2  \textbf{(+1.1)} & 41.7 \textbf{(+0.0)} & 43.0 \textbf{(+2.2)} & 42.1 \textbf{(+1.1)} \\
+            CSC                 & 40.6 \textbf{(+1.4)} & 42.2 \textbf{(+1.8)} & 42.3 \textbf{(+1.2)}  & 42.2 \textbf{(+0.6)} & 44.3 \textbf{(+3.5)} & 43.0 \textbf{(+2.0)} \\
+            DepthContrast       & 39.7 \textbf{(+0.5)} & 40.9 \textbf{(+0.5)} & --                    & --                   & --                   & --                   \\
+            CME \textbf{(Ours)} & 39.5 \textbf{(+0.3)} & 40.9 \textbf{(+0.5)} & 40.9 (-0.2)           & 41.8 \textbf{(+0.1)} & 41.8 \textbf{(+1.0)} & 42.0 \textbf{(+1.0)} \\
+        \end{tabular}
+    }
+    \caption{KITTI Semantic Segmentation results (mIOU)}
+    \label{table:kittiSemanticResults}
+\end{table}
+
+\begin{table}
+    \centering
+    \resizebox{0.3\textwidth}{!}{
+        \begin{tabular}{c | c c c}
+                                       & Scannet              & SunRGBD              \\
+            \hline
+            Scratch                    & 35.2                 & 32.0                 \\
+            \hline
+            PointContrast              & 36.7 \textbf{(+1.5)} & 34.2 \textbf{(+2.2)} \\
+            Contrastive Scene Contexts & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
+            DepthContrast              & 33.9 (-1.3)          & 32.9 \textbf{(+0.9)} \\
+            CME \textbf{(Ours)}        & 37.7 \textbf{(+2.5)} & 33.1 \textbf{(+1.1)} \\
+        \end{tabular}
+    }
+    \caption{Object Detection using VoteNet  (mAP)}
+    \label{table:scannetObjectDetection}
+\end{table}
+
 \subsection{Instance Segmentation}
 \label{sec:results:instance}
-
-\subsection{Object Detection}
-\label{sec:results:object}
 
@@
     \caption{S3DIS Semantic Segmentation results (mIOU)}
-    \label{table:s3disResults}
 \end{table*}
@@
     \caption{S3DIS Instance Segmentation Results (mAP@0.5)}
-    \label{table:s3disInstanceSegmention}
 \end{table*}
@@
     \caption{Scannet Semantic Segmentation with Partial Data  (mIOU)}
-    \label{table:scannetVaryingDataAmount}
 \end{table*}
@@
     \caption{Scannet Instance Segmentation with Partial Data  (mAP@0.5)}
-    \label{table:scannetInstanceVaryingDataAmount}
 \end{table*}
@@
     \caption{KITTI Semantic Segmentation results (mIOU)}
-    \label{table:kittiSemanticResults}
 \end{table*}
@@
     \caption{Object Detection using VoteNet  (mAP)}
-    \label{table:scannetObjectDetection}
 \end{table*}

### 2022-05-06 doc/paper.tex
adds plots to paper
updates results + semantic seg discussion

@@
 Only the required transformations are applied to the 2D image.
-These being: resizing, square cropping and colour normalization.
+These being: resizing, square cropping and color normalization.
 Point cloud coordinates are transformed with a random horizontal flip and random rotation around the $z$ axis while their colors are transformed using autocontrast, translation and jitter.
@@
 \clearpage
-\section{Experiment Results}
+\section{Experiments}
 \label{sec:results}
@@
 \textbf{Pretraining on 2D:} We use a batch size of 64 image pairs.
-For each pair we sample 4092 pixels to contrast in our loss function from eqn.
-\ref{eq:contrastive_loss}.
+For each pair we sample 4092 pixels to contrast in our loss function from Eq \ref{eq:contrastive_loss}.
 The loss function uses a $\tau$ of 0.4, a learning rate of 0.01, SGD optimizer with a momentum of 0.9, dampening of 0.1 and weight decay of 0.004.
@@
 
-\textbf{Transfer from 2D to 3D:} Transfer learning is done using the scannet dataset.
+\textbf{Transfer from 2D to 3D:} Transfer learning is done using the Scannet dataset.
 We use a batch size of 8 scan-image pairs and select 2000 point-pixel correspondences per pair.
@@
 The colors are also transformed randomly with chromatic autocontrast, translation and jitter.
-The hyperparameters are the same as those in 2D except the learning rate is set to 0.1.
+The hyper-parameters are the same as those in 2D except the learning rate is set to 0.1.
 It is trained for 20k iterations same as pre-training on 2D.
@@
 \textbf{Downstream Tasks:} All downstream tasks are run using the pipeline and parameters from \cite{hou2021Exploring}.
-It may be noted that we get slightly different numbers from those oringinaly reported.
+It may be noted that we get slightly different numbers from those originally reported.
 This comes down to a couple of factors, mainly an update to the core sparse convolution library that is required to run on modern GPUs, as well as using a single GPU instead of 8.
@@
 
-We first test the case of pretraining our model on ScannetV2 and finetuning on S3DIS.
-Our results are shown in table \ref{table:s3disResults}.
-Fully supervised pretraining has a drastic impact on the final downstream performance and is used as an upperbound on expected performance.
-Our algorithm is comparable to the performance of \cite{xie2020pointcontrast} however, that is without using overlapping 3D scans.
+We first test the case of pre-training our model on ScannetV2 and fine tuning on S3DIS.
+Detailed results are shown in table \ref{table:s3disResults}.
+Fully supervised pre-training has a drastic impact on the final downstream performance ($+5.1$ mIOU) and is used as a pseudo upper bound on expected performance.
+
+Almost all base algorithms improve downstream performance, except DepthContrast. Our algorithm is comparable to PointContrast, both achieving a 1 mIOU increase  but perform worse than CSC.
+
+Table \ref{table:scannetVaryingDataAmount} looks at using weights pre-trained on the entire Scannet dataset but then fine-tuned using only a portion of the labels.
+This simulates the use-case of capturing a large amount of data and only labeling a subset.
+All methods show no improvement when given access to labels for the entire dataset.
+The performance gains are unpredictable, with
+
+Interestingly, tightening the minimum overlap ratio between scan pairs improved performance on S3DIS.
+
+Our method has the advantage of not requiring the transformation between adjacent scans, and so to test how PointContrast and CSC perform under the same restriction we used the same scan
 
@@
 
-\begin{table*}[t!]
-    \centering
-    \resizebox{\textwidth}{!}{
-        \begin{tabular}{ c | c c | c c c c c c | c c c c c c |}
-                                &      &          & \multicolumn{6}{|c|}{Semantic Segmentation (mIOU)} & \multicolumn{6}{|c|}{Instance Segmentation (mAP@0.5)}                                                                                                                                                                                                                        \\
-                                & Scan & Overlap  & 5\%                                                & 10\%                                                  & 20\%                 & 30\%                 & 40\%                 & 100\%                 & 5\%                  & 10\%          & 20\%                 & 30\%                 & 40\%        & 100\%                \\
-            \hline
-            Scratch             & --   & --       & 42.2                                               & 51.1                                                  & 58.2                 & 59.2                 & 62.2                 & 67.4                  & 26.9                 & \textbf{31.1} & 40.7                 & 42.2                 & 45.7        & 49.0                 \\
-            \hline
-            PointContrast       & Pair & $>$ 30\% & 44.8 \textbf{(+2.6)}                               & 55.3 \textbf{(+4.2)}                                  & 58.0 (-0.2)          & 59.0 (-0.2)          & 63.2 \textbf{(+1.0)} & 66.9 (-0.5)           & 27.0 \textbf{(+0.1)} & 35.3          & 40.5 (-0.2)          & 43.0 \textbf{(+0.8)} & 43.9 (-1.8) & 49.1 \textbf{(+0.1)} \\
-            PointContrast       & Pair & $>$ 70\% & 46.1 \textbf{(+3.9)}                               & 50.7 (-0.4)                                           & 58.0 (-0.2)          & 61.9 \textbf{(+2.7)} & 62.6 \textbf{(+0.4)} & 67.3 (-0.1)           & --                   & --            & --                   & --                   & --          & --                   \\
-            PointContrast       & Same & --       & 45.5 \textbf{(+3.3)}                               & 48.8 (-2.3)                                           & 56.7 (-1.5)          & 59.3 \textbf{(+0.1)} & 61.7 (-0.5)          & 66.5 (-0.9)           & --                   & --            & --                   & --                   & --          & --                   \\
-            CSC                 & Pair & $>$ 30\% & 42.6 \textbf{(+0.4)}                               & 53.0 \textbf{(+1.9)}                                  & 57.3 (-0.9)          & 62.2 \textbf{(+3.0)} & 65.2 \textbf{(+3.0)} & 67.6 \textbf{(+0.2)}  & 29.8 \textbf{(+2.9)} & 35.5          & 41.6 \textbf{(+0.9)} & 42.9 \textbf{(+0.7)} & 44.6 (-1.1) & 49.3 \textbf{(+0.3)} \\
-            CSC                 & Pair & $>$ 70\% & 45.4 \textbf{(+3.2)}                               & 50.8 (-0.3)                                           & 59.3 \textbf{(+1.1)} & 59.7 \textbf{(+0.5)} & 62.7 \textbf{(+0.5)} & 68.0 \textbf{(+0.6)}  & --                   & --            & --                   & --                   & --          & --                   \\
-            CSC                 & Same & --       & 44.1 \textbf{(+1.9)}                               & 50.9 (-0.2)                                           & 56.6 (-1.6)          & 58.9 (-0.3)          & 61.5 (-0.7)          & 67.4 \textbf{(+0.0)}  & --                   & --            & --                   & --                   & --          & --                   \\
-            DepthContrast       & Same & --       & 49.3 \textbf{(+7.1)}                               & 54.3 \textbf{(+3.2)}                                  & 58.2 \textbf{(+0.0)} & --                   & --                   & --                    & 24.9 (-2.0)          & 33.8          & 38.2 (-2.5)          & --                   & --          & --                   \\
-            CME \textbf{(Ours)} & Same & --       & 44.0 \textbf{(+1.8)}                               & 53.0 \textbf{(+1.9)}                                  & 58.6 \textbf{(+0.4)} & 60.7 \textbf{(+1.5)} & 63.3 \textbf{(+1.1)} & 67.7 \textbf{(+0.28)} & 25.5 (-1.4)          & 35.5          & 39.1 (-1.6)          & 42.4 \textbf{(+0.2)} & 43.6 (-2.1) & 48.5 (-0.5)          \\
-        \end{tabular}
-    }
-    \caption{Scannet downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet}
-    \label{table:scannetVaryingDataAmount}
-\end{table*}
+\begin{itemize}
+    \item Comparable to point contrast
+    \item CSC does best with tight overlap
+    \item DepthContrast has no improvement
+    \item Point Contrast has no improvement when using the same scan x
+\end{itemize}
 
@@
             \hline
-            PointContrast       & Pair & $>$ 30\% & 66.2 \textbf{(+1.1)}          & 54.8 \textbf{(+1.8)}          \\
-            PointContrast       & Pair & $>$ 70\% & 68.6 \textbf{(+3.5)}          & 55.3 \textbf{(+2.3)}          \\
-            PointContrast       & Same & --       & 65.0 (-0.1)                   & 56.6 \textbf{(+3.6)}          \\
-            CSC                 & Pair & $>$ 30\% & 69.0 \textbf{(+3.9)}          & 57.8 \textbf{(+4.8)}          \\
-            CSC                 & Pair & $>$ 70\% & \textbf{70.1} \textbf{(+5.0)} & \textbf{58.9} \textbf{(+5.9)} \\
-            CSC                 & Same & --       & 67.9 \textbf{(+2.8)}          & 57.2 \textbf{(+4.2)}          \\
-            DepthContrast       & Same & --       & 64.9 (-0.2)                   & 52.3 (-0.7)                   \\
-            CME \textbf{(Ours)} & Same & --       & 66.5 \textbf{(+1.4)}          & 55.8 \textbf{(+2.8)}          \\
+            $\dagger$ PointContrast \cite{xie2020pointcontrast} & Pair & $>$ 30\% & 66.2 \textbf{(+1.1)}          & 54.8 \textbf{(+1.8)}          \\
+            PointContrast \cite{xie2020pointcontrast}           & Pair & $>$ 70\% & 68.6 \textbf{(+3.5)}          & 55.3 \textbf{(+2.3)}          \\
+            PointContrast  \cite{xie2020pointcontrast}          & Same & --       & 65.0 (-0.1)                   & 56.6 \textbf{(+3.6)}          \\
+            $\dagger$ CSC \cite{hou2021Exploring}               & Pair & $>$ 30\% & 69.0 \textbf{(+3.9)}          & 57.8 \textbf{(+4.8)}          \\
+            CSC \cite{hou2021Exploring}                         & Pair & $>$ 70\% & \textbf{70.1} \textbf{(+5.0)} & \textbf{58.9} \textbf{(+5.9)} \\
+            CSC \cite{hou2021Exploring}                         & Same & --       & 67.9 \textbf{(+2.8)}          & 57.2 \textbf{(+4.2)}          \\
+            $\dagger$ DepthContrast \cite{zhang2021Self}        & Same & --       & 64.9 (-0.2)                   & 52.3 (-0.7)                   \\
+            \hline
+            $\dagger$ CME \textbf{(Ours)}                       & Same & --       & 66.5 \textbf{(+1.4)}          & 55.8 \textbf{(+2.8)}          \\
         \end{tabular}
     }
-    \caption{S3DIS downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet}
+    \caption{\textbf{S3DIS downstream performance.} All methods use weights pre-trained on Scannet. Those with a $\dagger$ indicate the original algorithm whereas all others are variants. }
     \label{table:s3disResults}
@@
     \centering
-    \resizebox{0.5\textwidth}{!}{
+    \resizebox{0.47\textwidth}{!}{
         \begin{tabular}{ c | c c c c c c }
@@
             \hline
-            PointContrast       & 40.3 \textbf{(1.1)}  & 42.6 \textbf{(+2.2)} & 42.2  \textbf{(+1.1)} & 41.7 \textbf{(+0.0)} & 43.0 \textbf{(+2.2)} & 42.1 \textbf{(+1.1)} \\
-            CSC                 & 40.6 \textbf{(+1.4)} & 42.2 \textbf{(+1.8)} & 42.3 \textbf{(+1.2)}  & 42.2 \textbf{(+0.6)} & 44.3 \textbf{(+3.5)} & 43.0 \textbf{(+2.0)} \\
-            DepthContrast       & 39.7 \textbf{(+0.5)} & 40.9 \textbf{(+0.5)} & --                    & --                   & --                   & --                   \\
+            PointContrast \cite{xie2020pointcontrast} & 40.3 \textbf{(1.1)}  & 42.6 \textbf{(+2.2)} & 42.2  \textbf{(+1.1)} & 41.7 \textbf{(+0.0)} & 43.0 \textbf{(+2.2)} & 42.1 \textbf{(+1.1)} \\
+            CSC \cite{hou2021Exploring}               & 40.6 \textbf{(+1.4)} & 42.2 \textbf{(+1.8)} & 42.3 \textbf{(+1.2)}  & 42.2 \textbf{(+0.6)} & 44.3 \textbf{(+3.5)} & 43.0 \textbf{(+2.0)} \\
+            DepthContrast \cite{zhang2021Self}        & 39.7 \textbf{(+0.5)} & 40.9 \textbf{(+0.5)} & 41.2 \textbf{(+0.1)}  & 41.9 \textbf{(+0.2)} & --                   & --                   \\
+            \hline
             CME \textbf{(Ours)}                       & 39.5 \textbf{(+0.3)} & 40.9 \textbf{(+0.5)} & 40.9 (-0.2)           & 41.8 \textbf{(+0.1)} & 41.8 \textbf{(+1.0)} & 42.0 \textbf{(+1.0)} \\
@@
     }
-    \caption{KITTI Semantic Segmentation results (mIOU)}
+    \caption{\textbf{KITTI Semantic Segmentation results (mIOU)}. All methods use weights pre-trained on Scannet.  }
     \label{table:kittiSemanticResults}
@@
             PointContrast       & 36.7 \textbf{(+1.5)} & 34.2 \textbf{(+2.2)} \\
-            Contrastive Scene Contexts & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
+            CSC                 & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
             DepthContrast       & 33.9 (-1.3)          & 32.9 \textbf{(+0.9)} \\
+            \hline
             CME \textbf{(Ours)} & 37.7 \textbf{(+2.5)} & 33.1 \textbf{(+1.1)} \\
@@
     }
-    \caption{Object Detection using VoteNet  (mAP)}
+    \caption{\textbf{Object Detection using VoteNet (mAP)}}
     \label{table:scannetObjectDetection}
 \end{table}
+
+\begin{table*}[ht!]
+    \centering
+    \resizebox{\textwidth}{!}{
+        \begin{tabular}{ c | c c | c c c c c c | c c c c c c |}
+                                                                &      &          & \multicolumn{6}{|c|}{Semantic Segmentation (mIOU)} & \multicolumn{6}{|c|}{Instance Segmentation (mAP@0.5)}                                                                                                                                                                                                                               \\
+                                                                & Scan & Overlap  & 5\%                                                & 10\%                                                  & 20\%                 & 30\%                 & 40\%                 & 100\%                 & 5\%                  & 10\%                 & 20\%                 & 30\%                 & 40\%        & 100\%                \\
+            \hline
+            Scratch                                             & --   & --       & 42.2                                               & 51.1                                                  & 58.2                 & 59.2                 & 62.2                 & 67.4                  & 26.9                 & 36.3                 & 40.7                 & 42.2                 & 45.7        & 49.0                 \\
+            \hline
+            $\dagger$ PointContrast \cite{xie2020pointcontrast} & Pair & $>$ 30\% & 44.8 \textbf{(+2.6)}                               & 55.3 \textbf{(+4.2)}                                  & 58.0 (-0.2)          & 59.0 (-0.2)          & 63.2 \textbf{(+1.0)} & 66.9 (-0.5)           & 27.0 \textbf{(+0.1)} & 35.3 (-1.0)          & 40.5 (-0.2)          & 43.0 \textbf{(+0.8)} & 43.9 (-1.8) & 49.1 \textbf{(+0.1)} \\
+            PointContrast \cite{xie2020pointcontrast}           & Pair & $>$ 70\% & 46.1 \textbf{(+3.9)}                               & 50.7 (-0.4)                                           & 58.0 (-0.2)          & 61.9 \textbf{(+2.7)} & 62.6 \textbf{(+0.4)} & 67.3 (-0.1)           & 28.0 \textbf{(+1.1)} & 35.7 (-0.6)          & 40.5 (-0.2)          & 42.8 \textbf{(+0.6)} & 44.4 (-1.3) & --                   \\
+            PointContrast \cite{xie2020pointcontrast}           & Same & --       & 45.5 \textbf{(+3.3)}                               & 48.8 (-2.3)                                           & 56.7 (-1.5)          & 59.3 \textbf{(+0.1)} & 61.7 (-0.5)          & 66.5 (-0.9)           & --                   & --                   & --                   & --                   & --          & --                   \\
+            $\dagger$ CSC \cite{hou2021Exploring}               & Pair & $>$ 30\% & 42.6 \textbf{(+0.4)}                               & 53.0 \textbf{(+1.9)}                                  & 57.3 (-0.9)          & 62.2 \textbf{(+3.0)} & 65.2 \textbf{(+3.0)} & 67.6 \textbf{(+0.2)}  & 29.8 \textbf{(+2.9)} & 35.5 (-0.8)          & 41.6 \textbf{(+0.9)} & 42.9 \textbf{(+0.7)} & 44.6 (-1.1) & 49.3 \textbf{(+0.3)} \\
+            CSC \cite{hou2021Exploring}                         & Pair & $>$ 70\% & 45.4 \textbf{(+3.2)}                               & 50.8 (-0.3)                                           & 59.3 \textbf{(+1.1)} & 59.7 \textbf{(+0.5)} & 62.7 \textbf{(+0.5)} & 68.0 \textbf{(+0.6)}  & 29.8 \textbf{(+2.9)} & 37.6 \textbf{(+1.3)} & 41.4 \textbf{(+0.7)} & --                   & --          & --                   \\
+            CSC \cite{hou2021Exploring}                         & Same & --       & 44.1 \textbf{(+1.9)}                               & 50.9 (-0.2)                                           & 56.6 (-1.6)          & 58.9 (-0.3)          & 61.5 (-0.7)          & 67.4 \textbf{(+0.0)}  & 28.8 \textbf{(+1.9)} & 37.2 \textbf{(+0.9)} & 42.0 \textbf{(+1.3)} & --                   & --          & --                   \\
+            $\dagger$ DepthContrast \cite{zhang2021Self}        & Same & --       & 49.3 \textbf{(+7.1)}                               & 54.3 \textbf{(+3.2)}                                  & 58.2 \textbf{(+0.0)} & 59.6 \textbf{(+0.4)} & 63.2 \textbf{(+1.0)} & 67.4 \textbf{(+0.0)}  & 24.9 (-2.0)          & 33.8 (-2.5)          & 38.2 (-2.5)          & 42.0 (-0.2)          & 42.0 (-3.7) & 48.7 (-0.3)          \\
+            \hline
+            $\dagger$ CME \textbf{(Ours)}                       & Same & --       & 44.0 \textbf{(+1.8)}                               & 53.0 \textbf{(+1.9)}                                  & 58.6 \textbf{(+0.4)} & 60.7 \textbf{(+1.5)} & 63.3 \textbf{(+1.1)} & 67.7 \textbf{(+0.28)} & 25.5 (-1.4)          & 35.5 (-0.8)          & 39.1 (-1.6)          & 42.4 \textbf{(+0.2)} & 43.6 (-2.1) & 48.5 (-0.5)          \\
+        \end{tabular}
+    }
+    \caption{Scannet downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet}
+    \label{table:scannetVaryingDataAmount}
+\end{table*}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{images/plots/s3dis_semantic.pdf}
+    \includegraphics[width=8cm]{images/plots/scannet_0.05_semantic.pdf}
+    \includegraphics[width=8cm]{images/plots/scannet_semantic.pdf}
+    \caption{Comparison of validation performance over training steps for models pre-trained on Scannet and fine-tuned for semantic segmentation on, \textbf{Top:} S3DIS, \textbf{Middle:} Scannet with only 5\% labeled data, \textbf{Bottom:} Scannet}
+    \label{fig:validationComparison}
+\end{figure}
 
@@
 
-Things to include in results
-\begin{itemize}
-    \item Training and validation curves
-    \item Ablation studies
-\end{itemize}
+\section{Analysis}
+\label{sec:analysis}
 
-\section{Discussion}
+\subsection{Varying Overlap}
+\label{sec:results:varying}
+
+\subsection{Training Speedup}
+\label{sec:results:speedup}
+
+Talk about graphs showing that you can get close to final performance within a fraction of the time.
+
+\subsection{Inconsistencies in Performance Boost}
+\label{sec:results:inconsistencies}
 
@@
     \item Pre-training can help speed up training time on smaller datasets but has no impact on larger datasets.
+    \item Overal can expect about a 1\% increase in performance
+    \item Our method, which is easy to implement and does not require a pose transformation between frames, is a good start in trying to find best performance.
+    \item Also when using PointContrast \& CSC, you can experiment with overlap ratio and using tha same scan to get better results
 \end{itemize}
@@
     \bibliographystyle{ieee_fullname}
-    \bibliography{contrastive-learning}
+    \bibliography{paper}
 }

### 2022-05-06 doc/paper.tex
adds semantic segmentation results section

@@
 
-In some experiments, we vary the amount of training labeled data on downstream tasks to see the effectiveness of dataset size on the self-supervised pre-training.
+As is common in self-supervised learning experiments, we investigate the effect of varying the amount of labeled data to see the impact pre-training on a large unlabeled dataset has on smaller labeled datasets.
 
@@
     \centering
-    \includegraphics[width=7cm]{images/experiments/25.02.2022-image-pretrain-vis2.png}
-    \includegraphics[width=7cm]{images/experiments/25.02.2022-image-pretrain-vis1.png}
+    \includegraphics[width=6.5cm]{images/experiments/25.02.2022-image-pretrain-vis2.png}
+    \includegraphics[width=6.5cm]{images/experiments/25.02.2022-image-pretrain-vis1.png}
     \caption{\textbf{Left} Input image. \textbf{Right} Visualization of features from the pre-trained 2D backbone}
@@
 
-We first test the case of pre-training our model on ScannetV2 and fine tuning on S3DIS.
+We first test the case of pre-training our model on ScannetV2 and fine-tuning on S3DIS.
 Detailed results are shown in table \ref{table:s3disResults}.
 Fully supervised pre-training has a drastic impact on the final downstream performance ($+5.1$ mIOU) and is used as a pseudo upper bound on expected performance.
+Almost all base algorithms improve downstream results, except DepthContrast.
+Our algorithm is comparable to PointContrast, both achieving a 1\% mIOU increase but perform worse than CSC.
 
-Almost all base algorithms improve downstream performance, except DepthContrast. Our algorithm is comparable to PointContrast, both achieving a 1 mIOU increase  but perform worse than CSC.
+To simulate the use-case of capturing a large amount of un-labeled data for pre-training and only labeling a subset of it, we took Scannet and partitioned it into different portions of labeled data.
+The results can be found in table \ref{table:scannetVaryingDataAmount}.
+All methods show no improvement when given access to labels for the entire dataset.
+However, as the labeled data is restricted, the pre-trained models start to have an impact, with all methods improving upon training from scratch when given access to only 5\% of the labels.
 
-Table \ref{table:scannetVaryingDataAmount} looks at using weights pre-trained on the entire Scannet dataset but then fine-tuned using only a portion of the labels.
-This simulates the use-case of capturing a large amount of data and only labeling a subset.
-All methods show no improvement when given access to labels for the entire dataset.
-The performance gains are unpredictable, with
-
-Interestingly, tightening the minimum overlap ratio between scan pairs improved performance on S3DIS.
-
-Our method has the advantage of not requiring the transformation between adjacent scans, and so to test how PointContrast and CSC perform under the same restriction we used the same scan
-
-When using point correspondences without overlapping scans, the performance gain disappears almost entirely.
-This was noted in their original work and suggests to us that it is learning better features only due to its ability to better extrapolate surfaces.
-
-\begin{itemize}
-    \item Comparable to point contrast
-    \item CSC does best with tight overlap
-    \item DepthContrast has no improvement
-    \item Point Contrast has no improvement when using the same scan x
-\end{itemize}
+Similarly, we also test varying data amounts on the semanticKITTI dataset which can be found in table \ref{table:kittiSemanticResults}.
+We pre-trained all methods on Scannet to see how well the learned representations transferred to different data domains.
+In almost all cases there are performance improvements with CSC achieving the biggest increases.
+Unlike in Scannet, all algorithms have some impact across all labeled data ratios.
+Also having more labeled data increased the performance boost of the pre-trained model, which is the opposite of what we saw in Scannet and S3DIS.
+This suggests that pre-training on a dataset which is out-of-distribution from the labeled dataset is more informative than pre-training on in-distribution.
 
@@
     }
-    \caption{\textbf{S3DIS downstream performance.} All methods use weights pre-trained on Scannet. Those with a $\dagger$ indicate the original algorithm whereas all others are variants. }
+    \caption{\textbf{S3DIS downstream performance.} All methods use weights pre-trained on Scannet.
+Those with a $\dagger$ indicate the original algorithm whereas all others are variants.
+The difference in brackets is with respect to training from scratch.}
     \label{table:s3disResults}
@@
     }
-    \caption{\textbf{KITTI Semantic Segmentation results (mIOU)}. All methods use weights pre-trained on Scannet.  }
+    \caption{\textbf{KITTI Semantic Segmentation results (mIOU)}. All methods use weights pre-trained on Scannet. The difference in brackets is with respect to training from scratch.}
     \label{table:kittiSemanticResults}
-\end{table}
-
-\begin{table}
-    \centering
-    \resizebox{0.3\textwidth}{!}{
-        \begin{tabular}{c | c c c}
-                                & Scannet              & SunRGBD              \\
-            \hline
-            Scratch             & 35.2                 & 32.0                 \\
-            \hline
-            PointContrast       & 36.7 \textbf{(+1.5)} & 34.2 \textbf{(+2.2)} \\
-            CSC                 & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
-            DepthContrast       & 33.9 (-1.3)          & 32.9 \textbf{(+0.9)} \\
-            \hline
-            CME \textbf{(Ours)} & 37.7 \textbf{(+2.5)} & 33.1 \textbf{(+1.1)} \\
-        \end{tabular}
-    }
-    \caption{\textbf{Object Detection using VoteNet (mAP)}}
-    \label{table:scannetObjectDetection}
 \end{table}
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=8cm]{images/plots/s3dis_semantic.pdf}
-    \includegraphics[width=8cm]{images/plots/scannet_0.05_semantic.pdf}
-    \includegraphics[width=8cm]{images/plots/scannet_semantic.pdf}
-    \caption{Comparison of validation performance over training steps for models pre-trained on Scannet and fine-tuned for semantic segmentation on, \textbf{Top:} S3DIS, \textbf{Middle:} Scannet with only 5\% labeled data, \textbf{Bottom:} Scannet}
-    \label{fig:validationComparison}
-\end{figure}
-
 \subsection{Instance Segmentation}
@@
 \label{sec:results:object}
+
+\begin{table}
+    \centering
+    \resizebox{0.3\textwidth}{!}{
+        \begin{tabular}{c | c c c}
+                                & Scannet              & SunRGBD              \\
+            \hline
+            Scratch             & 35.2                 & 32.0                 \\
+            \hline
+            PointContrast       & 36.7 \textbf{(+1.5)} & 34.2 \textbf{(+2.2)} \\
+            CSC                 & 36.1 \textbf{(+0.9)} & 35.1 \textbf{(+3.1)} \\
+            DepthContrast       & 33.9 (-1.3)          & 32.9 \textbf{(+0.9)} \\
+            \hline
+            CME \textbf{(Ours)} & 37.7 \textbf{(+2.5)} & 33.1 \textbf{(+1.1)} \\
+        \end{tabular}
+    }
+    \caption{\textbf{Object Detection using VoteNet (mAP)}}
+    \label{table:scannetObjectDetection}
+\end{table}
 
@@
 
+When using point correspondences without overlapping scans, the performance gain disappears almost entirely.
+This was noted in their original work and suggests to us that it is learning better features only due to its ability to better extrapolate surfaces.
+
+Our algorithm is the only one that shows a consistent increase across varying Scannet data portions as well as S3DIS.
+Unlike PointContrast and CSC, our algorithm does not require pose transformations between adjacent scans making it easier to implement.
+
+The performance gains are unpredictable as the fraction of labeled data is varied.
+A performance boost of 3\% at 30\% labeled data can turn into a performance drop at 20\% labeled data.
+At the extreme of using 5\% labeled data, all algorithms have a performance boost.
+This further reinforces the idea that pre-training is most useful when the amount of labeled data is extremely limited.
+Interestingly, DepthContrast which was the worst performing on S3DIS is the best performing by a large margin on Scannet at 5\% labeled data, again showing the unpredictability of the results.
+
 \section{Conclusion}
@@
 \end{itemize}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=8cm]{images/plots/s3dis_semantic.pdf}
+    \includegraphics[width=8cm]{images/plots/scannet_0.05_semantic.pdf}
+    \includegraphics[width=8cm]{images/plots/scannet_semantic.pdf}
+    \caption{Comparison of validation performance over training steps for models pre-trained on Scannet and fine-tuned for semantic segmentation on: \textbf{Top:} S3DIS, \textbf{Middle:} Scannet with only 5\% labeled data, \textbf{Bottom:} Scannet}
+    \label{fig:validationComparison}
+\end{figure}
 

### 2022-05-06 doc/paper.tex
adds discussion on instance segmentation and object detection

@@
     }
-    \caption{Scannet downstream semantic (mIOU) and instance (mAP@0.5) segmentation performance using weights pre-trained in Scannet}
+    \caption{\textbf{Scannet downstream performance.} All methods use weights pre-trained on Scannet.
+Those with a † indicate the original algorithm whereas all others are variants.
+The difference in brackets is with respect to training from scratch.
+}
     \label{table:scannetVaryingDataAmount}
@@
 
+As in the previous section, we evaluate the performance of pre-training on Scannet and training on S3DIS for instance segmentation which can be found in table \ref{table:s3disResults}.
+Almost all pre-training methods get close or even surpass supervised results with the exception again being DepthContrast.
+The best performing method, as before, is Contrastive Scene Contexts which gets a performance boost of 4.8 mAP@0.5.
+However, our own method gets very close to supervised level performance with a boost of 2.8.
+
+However, looking at the effects on Scannet in table \ref{table:scannetVaryingDataAmount}, all methods struggle to improve upon training from scratch.
+In many cases they even perform substantially worse, with out method not being able to improve in almost all scenarios.
+Contrastive Scene Contexts is the only method that shows some consistent improvement.
+It is interesting to note that even though semantic segmentation receives a boost, this does not necessarily translate into instance segmentation.
+
 \subsection{Object Detection}
 \label{sec:results:object}
+
+Object detection is an interesting downstream task because it operates at an object level instead of a point level like the pre-training objective.
+Here we investigated the performance on Scannet and SunRGBD, which can be found in table \ref{table:scannetObjectDetection}.
+
+All methods besides DepthContrast show a strong improvement in both datasets.
+Contrastive Scene Contexts perform get the largest boost on SunRGBD of 3.1 mAP while our method gets the largest boost on Scannet with an increase of 2.5 mAP.
+All methods are still pre-trained on Scannet only and so to see performance improvements when using the entire Scannet dataset indicated that in the case of object detection, pre-training can help even on larger datasets.
 
@@
     \item Also when using PointContrast \& CSC, you can experiment with overlap ratio and using tha same scan to get better results
+    \item It may be worth trying a self-supervised method to see performance boost before spending extra time labeling the entire dataset
+    \item Using out-of-distribution data for pre-training can have a bigger impact that pre-training on in-distribution.
 \end{itemize}

### 2022-05-09 doc/paper.tex
adds some analysis sections

@@
 Detailed results are shown in table \ref{table:s3disResults}.
-Fully supervised pre-training has a drastic impact on the final downstream performance ($+5.1$ mIOU) and is used as a pseudo upper bound on expected performance.
+Fully supervised pre-training on Scannet has a drastic impact on the final downstream performance ($+5.1$ mIOU) and is used as a pseudo upper bound on expected performance.
 Almost all base algorithms improve downstream results, except DepthContrast.
@@
 However, looking at the effects on Scannet in table \ref{table:scannetVaryingDataAmount}, all methods struggle to improve upon training from scratch.
-In many cases they even perform substantially worse, with out method not being able to improve in almost all scenarios.
+In many cases they even perform substantially worse, with our method not being able to improve in almost all scenarios.
 Contrastive Scene Contexts is the only method that shows some consistent improvement.
@@
 
+Since our method does not rely on registered point clouds, we wanted to see how sensitive PointContrast and CSC are to using the same scan for both embeddings.
+We present out findings in tables \ref{table:s3disResults} and \ref{table:scannetVaryingDataAmount}.
+We found that for S3DIS using the same scan only performed worse overall however was able to do better in a few areas.
+Contrastive Scene Contexts also saw a small drop in performance.
+This was also noted in \cite{xie2020pointcontrast}.
+However, using a smaller minimum overlap ratio in most cases improved performance.
+Our method is easier to implement as it does not rely on knowing the transformation between point-clouds and generally gets more consistent increases across the range on available data labels than using the same scan for PointContrast and Contrastive Scene contexts.
+
+\subsection{Pre-training on Out-of-Distribution data}
+\label{sec:results:varying}
+
+Although the performance of using Scannet as the pre-training and downstream dataset has relatively low gains, those same weights have a much larger impact on S3DIS and consistently improve on KITTI across all data ratios.
+This discrepancy suggests that using a pre-training dataset which is out-of-distribution of the downstream dataset has a more consistent and larger performance gain than using the same dataset.
+This is potentially because using the same dataset already exposes the network to the same information, thus it is not learning as much new information as it could from a different dataset.
+
+Also, on KITTI, when using more of the data we see a larger performance improvement.
+Since the pre-training is done on different data, the algorithms benefit more from a larger labeled dataset.
+Unlike when using the same dataset for both pre-training and training where the two overlap more.
+
 \subsection{Training Speedup}
@@
 
-Talk about graphs showing that you can get close to final performance within a fraction of the time.
+Table \ref{fig:validationComparison} shows the validation curves over the training steps of training from scratch vs other methods.
+All methods show an immediate improvement early on in the training cycle.
+This is potentially useful during development on downstream tasks as the final performance can be known within the first 2000 cycles instead of 20000.
+In S3DIS for example, algorithms like point contrast can get to within 1 mIOU of the final scratch performance in just 1000 cycles.
+This can drastically reduce training time and development time.
 
@@
 \label{sec:results:inconsistencies}
+
+Across all of our experiments, its been impossible to see a consistent improvement in performance of each method across different datasets and labeled data ratios.
+For example, DepthContrast might get the largest increase by far when using 5\% of semantic Scannet labels but fails when applied to Scannet.
 
@@
     \item Using out-of-distribution data for pre-training can have a bigger impact that pre-training on in-distribution.
+    \item Weights that work on one task could be detrimental to performance on another task.
 \end{itemize}

### 2022-05-09 doc/paper.tex
finishes first draft of results

@@
 
-% \subsection{Evaluation Metrics}
-% \label{sec:results:metrics}
-% \dots
-
 \subsection{Contrastive Loss in 2D}
@@
 \subsection{Pre-training on Out-of-Distribution data}
-\label{sec:results:varying}
+\label{sec:results:out-of-distribution}
 
@@
 
-When using point correspondences without overlapping scans, the performance gain disappears almost entirely.
-This was noted in their original work and suggests to us that it is learning better features only due to its ability to better extrapolate surfaces.
-
-Our algorithm is the only one that shows a consistent increase across varying Scannet data portions as well as S3DIS.
-Unlike PointContrast and CSC, our algorithm does not require pose transformations between adjacent scans making it easier to implement.
-
-The performance gains are unpredictable as the fraction of labeled data is varied.
+In table \ref{table:scannetVaryingDataAmount}, the performance gains are unpredictable as the fraction of labeled data is varied.
 A performance boost of 3\% at 30\% labeled data can turn into a performance drop at 20\% labeled data.
@@
 
+Our algorithm is the only one that shows a consistent increase across varying Scannet data portions as well as S3DIS.
+Unlike PointContrast and CSC, our algorithm does not require pose transformations between adjacent scans making it easier to implement.
+
 \section{Conclusion}
@@
 
-\begin{itemize}
-    \item that transferring 2D features to 3D improves performance across different domains.
-    \item that although some algorithms perform extremely well in some domains, they fail in others. Or they might work well on one task but not another.
-    \item PointContrast \& CSC both ideally require overlapping scenes to perform properly whereas our algorithm does not.
-    \item Pre-training does little when large labeled datasets are used but are more effective when datasets are very small
-    \item Pre-training can help speed up training time on smaller datasets but has no impact on larger datasets.
-    \item Overal can expect about a 1\% increase in performance
-    \item Our method, which is easy to implement and does not require a pose transformation between frames, is a good start in trying to find best performance.
-    \item Also when using PointContrast \& CSC, you can experiment with overlap ratio and using tha same scan to get better results
-    \item It may be worth trying a self-supervised method to see performance boost before spending extra time labeling the entire dataset
-    \item Using out-of-distribution data for pre-training can have a bigger impact that pre-training on in-distribution.
-    \item Weights that work on one task could be detrimental to performance on another task.
-\end{itemize}
+In this work, we presented a method of transferring self-supervised features derived from dense images to models working on sparse point clouds.
+We showed how these features performed on three different downstream tasks and three different datasets.
+We also compared our approach to other methods.
+
+We found that our method was relatively consistent at improving downstream performance when compared to state of the art, with around a 1\% increase in performance on average.
+On larger datasets, using all labeled data for pre-training and training showed no performance increase, however when pre-trained on another dataset, performance did improve.
+
+It is very hard to tell which weights will perform best on which downstream task.
+From our experiments we found that using images had a benefit much more often than methods that used only point clouds.
+Since it works on raw captured 3D data, it is also easier to implement than those on registered point clouds.
+When looking for pre-training weights, our method is the best starting point to see if it can improve on a specific downstream task.
+
+Across all methods, using different datasets for pre-training than for training offered a bigger performance boost than using the same dataset.
+However, it the effects are still very inconsistent across domains.
+
+When starting to work on a problem, it may be worth while to try a self-supervised technique on a smaller labeled dataset before labeling it entirely.
+It can also be used to improve development iterations as it achieves close to final performance in a fraction of the training steps.
 
@@
 
+\section{Conclusion}
+
+\begin{itemize}
+    \item that transferring 2D features to 3D improves performance across different domains.
+    \item that although some algorithms perform extremely well in some domains, they fail in others. Or they might work well on one task but not another.
+    \item PointContrast \& CSC both ideally require overlapping scenes to perform properly whereas our algorithm does not.
+    \item Pre-training does little when large labeled datasets are used but are more effective when datasets are very small
+    \item Pre-training can help speed up training time on smaller datasets but has no impact on larger datasets.
+    \item Overal can expect about a 1\% increase in performance
+    \item Our method, which is easy to implement and does not require a pose transformation between frames, is a good start in trying to find best performance.
+    \item Also when using PointContrast \& CSC, you can experiment with overlap ratio and using tha same scan to get better results
+    \item It may be worth trying a self-supervised method to see performance boost before spending extra time labeling the entire dataset
+    \item Using out-of-distribution data for pre-training can have a bigger impact that pre-training on in-distribution.
+    \item Weights that work on one task could be detrimental to performance on another task.
+\end{itemize}
+
 \end{document}

### 2022-05-09 doc/paper.tex
adds some relevant works

@@
     \item Demonstrates that previous work in 3D \cite{xie2020pointcontrast} works by learning to extrapolate 3D scenes.
-When purely learning improved features on fully-overlapping regions the performance gain disapears.
+When purely learning improved features on fully-overlapping regions the performance gain disappears.
     \item Provide a 2D feature extractor and pre-training algorithm. Demonstrate that these features can improve 2D segmentation.
@@
 
-\textbf{Sparse 3D convolutions.} Talk about submanifold networks, Minkowski Engine, etc
+\textbf{Sparse 3D convolutions.} Talk about submanifold networks, Minkowski Engine \cite{choy20194d}, pointnet \cite{charles2017PointNet} PointNet++ \cite{qi2017pointnet++}.
 
-\textbf{Contrastive Learning.} Talk about SimCLR, BOYL, etc. Then talk about PointContrast.
+\textbf{Downstream tasks.} PointGroup \cite{}
 
-\textbf{Knowledge Distilation} Need to read up on this
+\textbf{Contrastive Learning.} Talk about SimCLR \cite{chen_simple_2020}, BOYL \cite{grill_bootstrap_nodate}, Moco \cite{he_momentum_2020}. Then talk about PointContrast. \cite{choy2019Fully}
+
+\textbf{Contrastive learning variations}
+Supervised contrastive losses \cite{khosla_supervised_2021} Guided contrastive learning in 3D \cite{jiang2021Guided}. Prototypical learning \cite{li_prototypical_2021}
+
+\textbf{Contrastive Learning in 3D}
+DepthContrast \cite{zhang2021Self}
+
+\textbf{Using 2D images.} 3D-sis \cite{hou20193d}, Using 2D images to learn 3D segmentation \cite{genova2021learning} \cite{frey_continual_2021}. 3dmv \cite{angela2018joint}
+
+Closest to my work: \cite{yunze2020p4contrast}.
+Their approach is to augment 3D points with features from 2D images and use their combined multi-modal features as the output of their model.
+They require however that their model always have access to 2D images and cannot operate on solely colorized point clouds at runtime.
 

### 2022-05-10 doc/paper.tex
adds more results

@@
 
+DenseCL \cite{wang_dense_2021} Does contrastive learning at the pixel level of images giving a good background into extracting image level features.
+
 \textbf{Contrastive learning variations}
@@
 \textbf{Contrastive Learning in 3D}
-DepthContrast \cite{zhang2021Self}
+DepthContrast \cite{zhang2021Self}, PointContrast \cite{xie2020pointcontrast}, Contrastive scene contexts \cite{hou2021Exploring}
+
+Occlusion completion \cite{wang_unsupervised_nodate}
 
@@
             $\dagger$ PointContrast \cite{xie2020pointcontrast} & Pair & $>$ 30\% & 44.8 \textbf{(+2.6)}                               & 55.3 \textbf{(+4.2)}                                  & 58.0 (-0.2)          & 59.0 (-0.2)          & 63.2 \textbf{(+1.0)} & 66.9 (-0.5)           & 27.0 \textbf{(+0.1)} & 35.3 (-1.0)          & 40.5 (-0.2)          & 43.0 \textbf{(+0.8)} & 43.9 (-1.8)          & 49.1 \textbf{(+0.1)} \\
-            PointContrast \cite{xie2020pointcontrast}           & Pair & $>$ 70\% & 46.1 \textbf{(+3.9)}                               & 50.7 (-0.4)                                           & 58.0 (-0.2)          & 61.9 \textbf{(+2.7)} & 62.6 \textbf{(+0.4)} & 67.3 (-0.1)           & 28.0 \textbf{(+1.1)} & 35.7 (-0.6)          & 40.5 (-0.2)          & 42.8 \textbf{(+0.6)} & 44.4 (-1.3) & --                   \\
-            PointContrast \cite{xie2020pointcontrast}           & Same & --       & 45.5 \textbf{(+3.3)}                               & 48.8 (-2.3)                                           & 56.7 (-1.5)          & 59.3 \textbf{(+0.1)} & 61.7 (-0.5)          & 66.5 (-0.9)           & --                   & --                   & --                   & --                   & --          & --                   \\
+            PointContrast \cite{xie2020pointcontrast}           & Pair & $>$ 70\% & 46.1 \textbf{(+3.9)}                               & 50.7 (-0.4)                                           & 58.0 (-0.2)          & 61.9 \textbf{(+2.7)} & 62.6 \textbf{(+0.4)} & 67.3 (-0.1)           & 28.0 \textbf{(+1.1)} & 35.7 (-0.6)          & 40.5 (-0.2)          & 42.8 \textbf{(+0.6)} & 44.4 (-1.3)          & 48.4 (-0.6)          \\
+            PointContrast \cite{xie2020pointcontrast}           & Same & --       & 45.5 \textbf{(+3.3)}                               & 48.8 (-2.3)                                           & 56.7 (-1.5)          & 59.3 \textbf{(+0.1)} & 61.7 (-0.5)          & 66.5 (-0.9)           & 27.9 \textbf{(+1.0)} & 36.0 (-0.3)          & 36.6 (-3.4)          & 43.5 \textbf{(+1.3)} & 44.5 (-1.2)          & 48.5 (-0.5)          \\
             $\dagger$ CSC \cite{hou2021Exploring}               & Pair & $>$ 30\% & 42.6 \textbf{(+0.4)}                               & 53.0 \textbf{(+1.9)}                                  & 57.3 (-0.9)          & 62.2 \textbf{(+3.0)} & 65.2 \textbf{(+3.0)} & 67.6 \textbf{(+0.2)}  & 29.8 \textbf{(+2.9)} & 35.5 (-0.8)          & 41.6 \textbf{(+0.9)} & 42.9 \textbf{(+0.7)} & 44.6 (-1.1)          & 49.3 \textbf{(+0.3)} \\
-            CSC \cite{hou2021Exploring}                         & Pair & $>$ 70\% & 45.4 \textbf{(+3.2)}                               & 50.8 (-0.3)                                           & 59.3 \textbf{(+1.1)} & 59.7 \textbf{(+0.5)} & 62.7 \textbf{(+0.5)} & 68.0 \textbf{(+0.6)}  & 29.8 \textbf{(+2.9)} & 37.6 \textbf{(+1.3)} & 41.4 \textbf{(+0.7)} & --                   & --          & --                   \\
-            CSC \cite{hou2021Exploring}                         & Same & --       & 44.1 \textbf{(+1.9)}                               & 50.9 (-0.2)                                           & 56.6 (-1.6)          & 58.9 (-0.3)          & 61.5 (-0.7)          & 67.4 \textbf{(+0.0)}  & 28.8 \textbf{(+1.9)} & 37.2 \textbf{(+0.9)} & 42.0 \textbf{(+1.3)} & --                   & --          & --                   \\
+            CSC \cite{hou2021Exploring}                         & Pair & $>$ 70\% & 45.4 \textbf{(+3.2)}                               & 50.8 (-0.3)                                           & 59.3 \textbf{(+1.1)} & 59.7 \textbf{(+0.5)} & 62.7 \textbf{(+0.5)} & 68.0 \textbf{(+0.6)}  & 29.8 \textbf{(+2.9)} & 37.6 \textbf{(+1.3)} & 41.4 \textbf{(+0.7)} & 43.9 \textbf{(+1.7)} & 47.3 \textbf{(+1.6)} & 48.7 (-0.3)          \\
+            CSC \cite{hou2021Exploring}                         & Same & --       & 44.1 \textbf{(+1.9)}                               & 50.9 (-0.2)                                           & 56.6 (-1.6)          & 58.9 (-0.3)          & 61.5 (-0.7)          & 67.4 \textbf{(+0.0)}  & 28.8 \textbf{(+1.9)} & 37.2 \textbf{(+0.9)} & 42.0 \textbf{(+1.3)} & 43.8 \textbf{(+1.6)} & 44.5 (-0.2)          & 48.5 (-0.5)          \\
             $\dagger$ DepthContrast \cite{zhang2021Self}        & Same & --       & 49.3 \textbf{(+7.1)}                               & 54.3 \textbf{(+3.2)}                                  & 58.2 \textbf{(+0.0)} & 59.6 \textbf{(+0.4)} & 63.2 \textbf{(+1.0)} & 67.4 \textbf{(+0.0)}  & 24.9 (-2.0)          & 33.8 (-2.5)          & 38.2 (-2.5)          & 42.0 (-0.2)          & 42.0 (-3.7)          & 48.7 (-0.3)          \\

### 2022-05-10 doc/paper.tex
adds related works

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=7cm]{images/experiments/image-to-point-vis1.png}
+    \includegraphics[width=7cm]{images/experiments/image-to-point-vis2.png}
+    \caption{\textbf{Left} Visualization of transferred features on the 3D point cloud. \textbf{Right}  Visualization of features from the pre-trained 2D backbone.}
+    \label{fig:features2d-3dvis}
+\end{figure}
+
 It is common practice in image tasks to use a model that was pre-trained in a supervised way on a larger but different dataset.
@@
 
-\textbf{Sparse 3D convolutions.} Talk about submanifold networks, Minkowski Engine \cite{choy20194d}, pointnet \cite{charles2017PointNet} PointNet++ \cite{qi2017pointnet++}.
+Our approach builds upon work on feature extraction and self-supervised learning in images and point clouds. Here we provide a brief summary of work in these fields.
 
-\textbf{Downstream tasks.} PointGroup \cite{}
+\textbf{Feature Extraction in 3D.} Raw sensor measurements capture 3D scenes as either depth or points.
+These points require processing for downstream tasks such as semantic segmentation or object detection.
+This can be accomplished in two ways.
+The first is to use the raw points themselves \cite{charles2017PointNet, qi2017pointnet++, Tatarchenko2018TangentConvolutions, thomas2019KPConv, wu2019PointConv, engelmann2020Dilated, zhao2020point} which does not require any other expensive transformations.
+The other is to transform these points into an intermediate representation, such as discrete voxels \cite{cicek20163d, riegler2017octnet, graham20183D, choy20194d}.
+Any of these approaches can equally be used for extracting meaningful features for downstream tasks.
+However, in this work we chose to use the 3D U-Net voxel-based convolution presented in Chow \etal \cite{choy20194d}.
+Not only does it offer close to state of the art performance but many recent state of the art methods \cite{jiang2020pointgroup, chen2021hierarchical} have adopted this voxel-based feature extractor.
 
-\textbf{Contrastive Learning.} Talk about SimCLR \cite{chen_simple_2020}, BOYL \cite{grill_bootstrap_nodate}, Moco \cite{he_momentum_2020}. Then talk about PointContrast. \cite{choy2019Fully}
+\textbf{Self-Supervised Transfer learning.} Self-supervised learning uses a pretext task that operates on raw data directly to form a loss function that learns meaningful representations.
+It can take the form of generative models such as auto-encoders \cite{goodfellow2014generative, Kingma2014}, which learn how to model the input and make examples that are close to the real distribution.
+Or discriminative models, that aim to discriminative between inputs in high dimensional space.
 
-DenseCL \cite{wang_dense_2021} Does contrastive learning at the pixel level of images giving a good background into extracting image level features.
+Contrastive learning is a discriminative learning approach which contrasts different inputs to learn how to discriminate between positive and negative samples.
+It has been shown to be incredibly powerful in image classification.
+SimCLR \cite{chen_simple_2020}  uses a very simple framework which achieves close to supervised level performance.
+Some methods focus on increasing the batch size through a memory bank \cite{wu2018unsupervied} or momentum encoder \cite{he_momentum_2020}.
 
-\textbf{Contrastive learning variations}
-Supervised contrastive losses \cite{khosla_supervised_2021} Guided contrastive learning in 3D \cite{jiang2021Guided}. Prototypical learning \cite{li_prototypical_2021}
+Choosing which points to contrast is a huge challenge as some points are uninformative and others may belong to the same class.
+To address this issue, some methods select samples in a more informative way using negative mining techniques \cite{robinson_contrastive_2021, kalantidis_hard_nodate, chuang_debiased_2020, ho_contrastive_nodate} to improve the strength of the self-supervised signal.
+Other methods try to bootstrap which points should be contrasted.
+Li \etal \cite{li2021prototypical} iteratively groups representations into prototypes to avoid contrasting points from the same cluster.
+Khosla \etal \cite{khosla_supervised_2021} use a mix of supervised learning to generate pseudo labels that inform the selection of points during self-supervised learning.
+
+The bulk of this work has been on high-level image classification, but it has proved useful in pixel level image segmentation as well \cite{wang_dense_2021, xie2021propagate, wang2021exploring}.
+Since these features were useful in 2D segmentation, it suggests that they can be equally as useful in 3D.
 
 \textbf{Contrastive Learning in 3D}
-DepthContrast \cite{zhang2021Self}, PointContrast \cite{xie2020pointcontrast}, Contrastive scene contexts \cite{hou2021Exploring}
 
-Occlusion completion \cite{wang_unsupervised_nodate}
+The same techniques that were used in images can translate directly into 3D point clouds.
+This was first used for generating unique features for point clouds registration \cite{choy2019Fully}.
+This was subsequently adapted by PointContrast \cite{xie2020pointcontrast} to work for semantic segmentation.
+Point pairs are generated using two scans of registered overlapping point clouds and finding points which are spatially close in both.
+The scans are randomly transformed to learn transformation invariant features.
+Contrastive scene contexts \cite{hou2021Exploring} expands on this idea by sampling points from partitioned spatial contexts.
+This work extends point contrast to use instead of a second scan, to use features from an image of the same scan.
 
-\textbf{Using 2D images.} 3D-sis \cite{hou20193d}, Using 2D images to learn 3D segmentation \cite{genova2021learning} \cite{frey_continual_2021}. 3dmv \cite{angela2018joint}
+Alternatively, DepthContrast \cite{zhang2021Self} seeks to learn point-level representations using scan level features.
+They implement a momentum encoder \cite{he_momentum_2020} to operate on scene level descriptors generated from individual scans which are then contrasted with each other.
+Using other high level tasks like occlusion completion \cite{wang_unsupervised_nodate} has also proven useful.
 
-Closest to my work: \cite{yunze2020p4contrast}.
+The same sampling challenges exist in point clouds as they do for images. To handle this \cite{jiang2021Guided} use pseudo labels to pick points to avoid contrasting from the same class.
+
+\textbf{Sensor Fusion.} Our work seeks to leverage features learned exclusively from images and transfer them to a 3D model.
+Combining the two modalities has proven successful in learning 3D segmentation \cite{hou20193d, genova2021learning, frey_continual_2021, angela2018joint} as well as using multiple images of a single scene in multi-view fusion \cite{kundu2020virtual}.
+
+P4Contrast \cite{yunze2020p4contrast} combines sensor fusion with self-supervised pre-training.
 Their approach is to augment 3D points with features from 2D images and use their combined multi-modal features as the output of their model.
-They require however that their model always have access to 2D images and cannot operate on solely colorized point clouds at runtime.
+Their approach is to fuse both modalities during pre-training and runtime and so always require access to 2D images.
+This means they cannot operate on solely colorized point clouds at runtime.
+Our method seeks to only take learned 2D representations and transfer them to a 3D model during pre-training.
 
@@
 
+This section will layout a theoretical framework for contrastive learning that will be used throughout the rest of this work.
 The goal of self-supervised transfer learning is to learn from unlabeled data using a self-supervised objective which in turn initializes a model's backbone.
@@
 
-\AJ{Add figure of what your feature extractor looks like}
-
-\AJ{Need to review how I do this in the code, I think I use a projector left over from BYOL}
-
 \subsection{Feature projection}
@@
 
-\AJ{Do I need to give the equation for Bi-linear interpolation or is this also considered basic knowledge}
-
 \subsection{Feature Transfer}
@@
 
-\AJ{Need to look at not using a projection head in 3D model.}
-
-\clearpage
 \section{Experiments}
@@
     \label{fig:features2dvis}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=7cm]{images/experiments/image-to-point-vis1.png}
-    \includegraphics[width=7cm]{images/experiments/image-to-point-vis2.png}
-    \caption{\textbf{Left} Visualization of transferred features on the 3D point cloud. \textbf{Right}  Visualization of features from the pre-trained 2D backbone.}
-    \label{fig:features2d-3dvis}
 \end{figure}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{images/plots/s3dis_semantic.pdf}
+    \includegraphics[width=8cm]{images/plots/scannet_0.05_semantic.pdf}
+    % \includegraphics[width=8cm]{images/plots/scannet_semantic.pdf}
+    \caption{Comparison of validation performance over training steps for models pre-trained on Scannet and fine-tuned for semantic segmentation on: \textbf{Top:} S3DIS, \textbf{Bottom:} Scannet with only 5\% labeled data}
+    \label{fig:validationComparison}
+\end{figure}
+
 \section{Conclusion}
@@
 It can also be used to improve development iterations as it achieves close to final performance in a fraction of the training steps.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=8cm]{images/plots/s3dis_semantic.pdf}
-    \includegraphics[width=8cm]{images/plots/scannet_0.05_semantic.pdf}
-    \includegraphics[width=8cm]{images/plots/scannet_semantic.pdf}
-    \caption{Comparison of validation performance over training steps for models pre-trained on Scannet and fine-tuned for semantic segmentation on: \textbf{Top:} S3DIS, \textbf{Middle:} Scannet with only 5\% labeled data, \textbf{Bottom:} Scannet}
-    \label{fig:validationComparison}
-\end{figure}
 

### 2022-05-14 doc/paper.tex
adds overview diagram

@@
 %%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
-\usepackage[review]{cvpr}      % To produce the REVIEW version
+\usepackage{cvpr}      % To produce the REVIEW version
 % \usepackage{cvpr}              % To produce the CAMERA-READY version
@@
 In this work, we adopt the instance discrimination task as in \cite{} whereby each input individually is taken to belong to its own class.
+
+\begin{figure}
+    \centering
+    %angle=-90,origin=c
+    \includegraphics[width=8.5cm]{images/overview.pdf}
+    \caption{Overview of }
+    \label{fig:overview}
+\end{figure}
 
@@
 
+\section{Notes from Edwin}
+
+\begin{itemize}
+    \item Augmentations are random
+    \item Explaining that it's a two step process, and thats how its different from prior work
+    \item Goal is to see how 2D self-supervised features transferred to a 3D model work
+    \item Mention that it uses a single image
+\end{itemize}
+
+\textbf{description:} Three steps.
+First step is pre-training, random combinations of a single image pixel contrast.
+Second stage is try to get 2D to 3D features match.
+Third is fine-tuning on 3D data.
+
 \end{document}

### 2022-05-14 doc/paper.tex
adds caption to overview diagram

@@
 
+Labeling effort: SemanticKITTI had 518 tiles of 100m x 100m and took 4.5 hours per residential tile and 1.5 hours per highway. Total of 1700 hours. Scannet took 20 people 1 month just to capture data.
+
 It is common practice in image tasks to use a model that was pre-trained in a supervised way on a larger but different dataset.
@@
     \centering
-    %angle=-90,origin=c
     \includegraphics[width=8.5cm]{images/overview.pdf}
-    \caption{Overview of }
+    \caption{
+        \textbf{Overview of our pre-training pipeline.} \textbf{Stage 1:} A 2D model is pre-trained by applying a contrastive infoNCE loss on two augmentations of a single image at pixel level.
+\textbf{Stage 2:} The same loss is applied on the outputs of the pre-trained 2D model and raw 3D model by mapping the point features into the image.
+Above, black represents the sample query point, while green represents the positive point and red represents negative points to be contrasted against.
+    }
     \label{fig:overview}

### 2022-05-14 doc/paper.tex
Reworks abstract

@@
 University of Toronto Institute for Aerospace Studies\\
-Institution1 address\\
 {\tt\small andrej.janda@robotics.utias.utoronto.ca}
@@
 University of Toronto Institute for Aerospace Studies\\
-First line of institution2 address\\
 {\tt\small jonathan.kelly@robotics.utias.utoronto.ca}
@@
 \begin{abstract}
-    We show that previous work \cite{xie2020pointcontrast} on self-supervised contrastive learning of 3D point clouds relies on extrapolating points in the scene and not from learning noise as was claimed in the original paper.
-We then investigate if using 2D pre-trained features transfered to 3D can do as well as fully supervised learning.
+    Self-supervised pre-training on large unlabeled datasets has proven to be incredibly successful in 2D vision tasks.
+Reducing the amount of annotations allows models to be applied in more areas where, due to resource constraints, it was not possible before.
+This is especially true for 3D vision tasks as the datasets are much smaller owing mainly to the difficulty of annotating 3D scenes.
+Previous works have shown that pre-training on point clouds alone can improve downstream performance.
+In this work, we investigate how well 2D self-supervised features transfer to 3D tasks, as they offer much higher resolutions.
+They can also leverage information learned on a large array of labeled 2D datasets and unlike some 3D methods, do not require point cloud registration.
+We compare how these features perform against other 3D pre-training algorithms.
+We find that the downstream performance improvement was more consistent across datasets than other point only based methods.
+However, all methods were beneficial when access to labels was severely limited.
 \end{abstract}
@@
 
-% Write topic sentences and go from there
 Segmentation of indoor 3D scenes has the potential to allow robots to navigate and interact with complicated industrial environments, such as warehouses and production facilities.
@@
 
-In contrast to 2D images which are bountiful, easy to collect and annotate, 3D scenes are hard to collect, not very abundant and notoriously hard to annotate.
-Collecting 3D scenes requires a user to physically scan different environments with complicated sensors (\eg stereo cameras, Lidar) and estimate the mapping between measurements to generate a completed scene.
-Since not everyone has access to these sensors, they cannot be easily scrapped off of the web.
-Still, the hardest component is annotation which requires users to manipulate 3D scenes and select only those points that belong to a particular object.
-This motivates us to look for ways of avoiding the annotation process and leverage un-labeled 3D data instead.
-
 \begin{figure}
@@
     \includegraphics[width=7cm]{images/experiments/image-to-point-vis2.png}
-    \caption{\textbf{Left} Visualization of transferred features on the 3D point cloud. \textbf{Right}  Visualization of features from the pre-trained 2D backbone.}
+    \caption{\textbf{Visualization of the impact transfering pixel features onto pointclouds.} \textbf{Left} Visualization of transferred features on the 3D point cloud.
+\textbf{Right}  Visualization of features from the pre-trained 2D backbone.}
     \label{fig:features2d-3dvis}
 \end{figure}
+
+In contrast to 2D images which are bountiful, easy to collect and annotate, 3D scenes are hard to collect, not very abundant and notoriously hard to annotate.
+Collecting 3D scenes requires a user to physically scan different environments with complicated sensors (\eg stereo cameras, Lidar) and estimate the mapping between measurements to generate a completed scene.
+Since not everyone has access to these sensors, they cannot be easily scrapped off of the web.
+Even so, the hardest component is annotation which requires users to manipulate 3D scenes and select only those points that belong to a particular object.
+This motivates us to look for ways of avoiding the annotation process and leverage un-labeled 3D data instead.
 
@@
 This section goes through the types of backbones used as well as how features are extracted from 2D images and transferred to a 3D model.
-
-\AJ{Need a picture depicting the pipeline}
+For an overview of our pipeline, please refer to figure \ref{fig:overview}.
 
@@
 Corresponding pixels are taken as positive samples while all others, including from other images in a batch are taken as negatives.
+
+\AJ{Add Kd-tree mapping to get the correct pixels after transformation.}
 

### 2022-05-14 doc/paper.tex
reworks the intro

@@
 
-Segmentation of indoor 3D scenes has the potential to allow robots to navigate and interact with complicated industrial environments, such as warehouses and production facilities.
-Although robots already play a limited role in these facilities, understanding their surroundings would give them much greater autonomy.
-For instance, through segmentation, an accurate up-to-date digital twin of a physical area can be maintained for the purposes of inventory management, item retrieval, and item routing.
-Advances in segmentation of 3D point clouds \cite{choy20194d} has made this closer to reality.
-
-Alternatively, segmentation of 2D images has come a long way.
-But it does not give us information about the 3D world and so we are left with projecting these predictions into 3D.
-This was the approach taken by \cite{} but has recently been overtaken by methods operating purely in 3D.
+Our world exists in three spatial dimensions.
+By processing data that also exists in three dimensions, we are able to get a more cohesive understanding of our environment which cannot be generated from images alone.
+This unlocks applications particularly in robotics, by allowing machines to interact with the physical world in a meaningful way.
+Through point-cloud segmentation, robots can navigate and interact within complicated settings, such as streets, warehouses, and production facilities.
+Recent advances in 3D vision have made this closer to reality, however, often require large datasets to function adequately.
 
@@
     \includegraphics[width=7cm]{images/experiments/image-to-point-vis2.png}
-    \caption{\textbf{Visualization of the impact transfering pixel features onto pointclouds.} \textbf{Left} Visualization of transferred features on the 3D point cloud.
+    \caption{\textbf{Visualization of the how pixel features have transferred into the 3D model.} \textbf{Left} Visualization of transferred features on the 3D point cloud.
 \textbf{Right}  Visualization of features from the pre-trained 2D backbone.}
@@
 Collecting 3D scenes requires a user to physically scan different environments with complicated sensors (\eg stereo cameras, Lidar) and estimate the mapping between measurements to generate a completed scene.
-Since not everyone has access to these sensors, they cannot be easily scrapped off of the web.
-Even so, the hardest component is annotation which requires users to manipulate 3D scenes and select only those points that belong to a particular object.
+Since not everyone has access to these sensors, they cannot be easily scrapped off of the web like with images.
+The hardest part, however, is annotation, which requires users to manipulate 3D scenes and select only those points that belong to a particular object.
+
+This has been a challenge faced by the existing 3D datasets.
+The semanticKITTI dataset \cite{behley2019semantic} that has 518 square tiles of 100m each took 1700 hours to label.
+Scannet took 20 individuals 1 month just to capture all the scenes.
 This motivates us to look for ways of avoiding the annotation process and leverage un-labeled 3D data instead.
 
-Labeling effort: SemanticKITTI had 518 tiles of 100m x 100m and took 4.5 hours per residential tile and 1.5 hours per highway. Total of 1700 hours. Scannet took 20 people 1 month just to capture data.
-
 It is common practice in image tasks to use a model that was pre-trained in a supervised way on a larger but different dataset.
-However, recent works have been exploring the use of self-supervised objectives without labels in 2D image classification and found that they can greatly improve downstream accuracy, even matching fully supervised training \cite{}.
+However, recent works have been exploring the use of self-supervised objectives without labels in 2D image classification and found that they can greatly improve downstream accuracy, even matching fully supervised training \cite{grill_bootstrap_nodate}.
 
@@
 These algorithms work exclusively with point clouds but throw away the corresponding images of the scene which provide a rich and dense source in extra information that can be used in downstream training.
-This paper investigates how this extra data modality can be leveraged.
-Additionally, it investigates the properties of an existing 3D pre-training algorithm.
+This work investigates how this extra data modality can be leveraged and how it compares to methods that use point clouds exclusively.
 
@@
 \begin{itemize}
-    \item Demonstrates that previous work in 3D \cite{xie2020pointcontrast} works by learning to extrapolate 3D scenes.
-When purely learning improved features on fully-overlapping regions the performance gain disappears.
-    \item Provide a 2D feature extractor and pre-training algorithm. Demonstrate that these features can improve 2D segmentation.
-    \item Transfer learned 2D features onto a 3D model
+    \item A method for extracting 2D self-supervised features and transferring them to a 3D model.
+    \item Shows that features transferred from images can be beneficial in many 3D computer vision tasks
+    \item An analysis of how various pre-training methods perform on different datasets and across various tasks in 3D.
 \end{itemize}

### 2022-10-18 scripts/experiments/pretrain/old_loss.sh
loads of changes

@@ -2,10 +2,11 @@ python src/main.py \
     dataset_dir=$dataset_dir \
     dataset=scannet \
     model=minkowski \
+    model.name=pointcontrast \
     dataset.name=scannetv2_pretrain_new \
     tasks=["pretrain"] \
-    hydra.run.dir=$output_dir/pretrain/old-loss \
-    gpus=[1] \
+    hydra.run.dir=$output_dir/pretrain/old-loss-scan-only \
+    gpus=[0] \
     max_time="01:00:00:00" \
     dataset.pretrain.batch_size=16 \
     dataset.pretrain.accumulate_grad_batches=4 \
@@ -13,5 +14,5 @@ python src/main.py \
     check_val_every_n_epoch=-1 \
     model.optimizer.type=SGD \
     model.optimizer.lr=0.1 \
-    val_check_interval=1000 \
-    pretrain_checkpoint=\"epoch=5-step=6364-val_loss=6.14.ckpt\"
+    val_check_interval=1000 
+    # pretrain_checkpoint=\"epoch=5-step=6364-val_loss=6.14.ckpt\"

### 2022-10-18 scripts/experiments/scannet/minkowski_material_properties.sh

@@ -1,0 +1,17 @@
+python src/main.py \
+    model=minkowski \
+    hydra.run.dir=$output_dir/scannetv2/minkowski_material_properties \
+    model.name=minkowski \
+    tasks=["train","eval"] \
+    gpus=[1] \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    dataset.batch_size=32 \
+    dataset.input_channels=4 \
+    model.net.conv1_kernel_size=4 \
+    dataset.accumulate_grad_batches=2 \
+    dataset.voxel_size=0.05 \
+    max_time="00:12:00:00" \
+    max_epochs=2000 \
+    check_val_every_n_epoch=20 \
+    model.train.train_workers=10

### 2022-10-18 scripts/experiments/scannet/minkowski_vanilla.sh

@@ -1,0 +1,15 @@
+python src/main.py \
+    model=minkowski \
+    hydra.run.dir=$output_dir/scannetv2/minkowski_material_properties \
+    model.name=minkowski \
+    tasks=["train","eval"] \
+    gpus=[1] \
+    dataset_dir=$dataset_dir \
+    dataset=scannet \
+    dataset.batch_size=32 \
+    dataset.accumulate_grad_batches=2 \
+    dataset.voxel_size=0.05 \
+    max_time="00:12:00:00" \
+    max_epochs=2000 \
+    check_val_every_n_epoch=20 \
+    model.train.train_workers=10

### 2022-10-18 scripts/preprocess/scannet.sh

@@ -1,5 +1,5 @@
 export SCANNET_DIR=/media/starslab/datasets/scannet
-export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_2cm_frames_with_image_labels
+export OUTPUT_DIR=/media/starslab/datasets/scannet_preprocessed_vanilla
 export HYDRA_FULL_ERROR=1
 
 preprocess() {
@@ -10,11 +10,11 @@ preprocess() {
         output_dir=$OUTPUT_DIR \
         dataset=scannet \
         dataset.voxel_size=0.02 \
-        dataset.name=scannetv2_pretrain_new \
+        dataset.name=scannetv2 \
         sens_file=$filename
 }
 
 export -f preprocess
 
 # preprocess
-parallel -j 8 --linebuffer time preprocess ::: `find $SCANNET_DIR/scans/ -mindepth 1 -maxdepth 1 -type d`
+parallel -j 12 --linebuffer time preprocess ::: `find $SCANNET_DIR/scans/ -mindepth 1 -maxdepth 1 -type d`

### 2022-10-18 scripts/pretrain.sh

@@ -1,5 +1,5 @@
 # Initalize environment
-dataset_dir=/media/starslab/datasets/scannet_preprocessed_2cm_frames_with_labels
+dataset_dir=/media/starslab/datasets/scannet_preprocessed_2cm_frames_and_scans
 source .venv/bin/activate
 
 # Set output folder

### 2022-10-18 scripts/run_scannet.sh

@@ -1,6 +1,6 @@
 
 # Initalize environment
-dataset_dir=/media/starslab/datasets/scannet_preprocessed_2cm_frames_and_scans
+dataset_dir=/media/starslab/datasets/scannet_material_properties
 source .venv/bin/activate
 
 # Set output folder

### 2022-10-18 src/main.py

@@ -285,7 +285,7 @@ class Trainer:
         dataset_type = self.model_factory.get_backbone_dataset_type()
         dataset = dataset_type(self.data_interface.pretrain_val_data, self.cfg)
         collate_fn = dataset.collate
-        scene = collate_fn([dataset[2100]])
+        scene = collate_fn([dataset[2350]])
 
         # Generate feature embeddings
         z1 = backbonewraper.model(scene.images2).detach().cpu().numpy()
@@ -321,7 +321,7 @@ class Trainer:
         dataset_type = self.model_factory.get_backbone_dataset_type()
         dataset = dataset_type(self.data_interface.pretrain_val_data, self.cfg)
         collate_fn = dataset.collate
-        scene = collate_fn([dataset[2100]])
+        scene = collate_fn([dataset[3430]])
 
         # Generate feature embeddings
         input_3d_1 = ME.SparseTensor(scene.features1, scene.points1)
@@ -345,7 +345,13 @@ class Trainer:
         )
 
         # o3d.io.write_triangle_mesh("vis_3d.obj", vis_pcd, write_vertex_normals=False)
-        o3d.visualization.draw_geometries([vis_pcd])
+        o3d.visualization.draw_geometries(
+            [vis_pcd],
+            front=[-0.93761197130076546, -0.17869957213484086, 0.29824529198684352],
+            lookat=[0.5, -3.5, 4.0],
+            up=[0.2765394863307537, 0.13663222521433616, 0.95124000522102836],
+            zoom=0.69999999999999996,
+        )
 
         waithere = 1
 

### 2022-10-18 src/models/factory.py

@@ -1,5 +1,6 @@
 from pathlib import Path
 from omegaconf import DictConfig
+import torch
 
 from models.minkowski.trainer import (
     MinkowskiTrainer,
@@ -84,11 +85,17 @@ class ModelFactory:
 
     def load_from_checkpoint(self, checkpoint_path: Path):
         if minkowski_name in self.model_name:
-            return MinkowskiTrainer.load_from_checkpoint(
-                cfg=self.cfg,
-                data_interface=self.data_interface,
-                checkpoint_path=checkpoint_path,
-            )
+            try:
+                return MinkowskiTrainer.load_from_checkpoint(
+                    cfg=self.cfg,
+                    data_interface=self.data_interface,
+                    checkpoint_path=checkpoint_path,
+                )
+            except:
+                state_dict = torch.load(checkpoint_path)
+                trainer = MinkowskiTrainer(self.cfg, self.data_interface)
+                trainer.model.load_state_dict(state_dict["state_dict"], strict=False)
+                return trainer
 
         else:
             raise RuntimeError(self.error_msg)

### 2022-10-18 src/models/minkowski/modules/res16unet.py

@@ -37,14 +37,14 @@ class Res16UNetBase(ResNetBase):
         def space_n_time_m(n, m):
             return n if D == 3 else [n, n, n, m]
 
-        if D == 4:
-            self.OUT_PIXEL_DIST = space_n_time_m(self.OUT_PIXEL_DIST, 1)
+        # if D == 4:
+        #     self.OUT_PIXEL_DIST = space_n_time_m(self.OUT_PIXEL_DIST, 1)
 
         self.inplanes = self.INIT_DIM
         self.conv0p1s1 = conv(
             in_channels,
             self.inplanes,
-            kernel_size=space_n_time_m(config.net.conv1_kernel_size, 1),
+            kernel_size=config.net.conv1_kernel_size,
             stride=1,
             dilation=1,
             conv_type=self.NON_BLOCK_CONV_TYPE,

### 2022-10-18 src/models/minkowski/modules/resnet.py

@@ -38,8 +38,8 @@ class ResNetBase(Model):
         def space_n_time_m(n, m):
             return n if D == 3 else [n, n, n, m]
 
-        if D == 4:
-            self.OUT_PIXEL_DIST = space_n_time_m(self.OUT_PIXEL_DIST, 1)
+        # if D == 4:
+        #     self.OUT_PIXEL_DIST = space_n_time_m(self.OUT_PIXEL_DIST, 1)
 
         dilations = config.dilations
         bn_momentum = config.opt.bn_momentum
@@ -47,7 +47,7 @@ class ResNetBase(Model):
         self.conv1 = conv(
             in_channels,
             self.inplanes,
-            kernel_size=space_n_time_m(config.conv1_kernel_size, 1),
+            kernel_size=config.conv1_kernel_size,
             stride=1,
             D=D,
         )
